---
title: "MAPSPAM_tonnes_production.Rmd"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Summary

# Data sources

 - GAEZ: https://www.nature.com/articles/s41597-021-01115-2
 - Halpern et al. 2022: https://knb.ecoinformatics.org/view/doi%3A10.5063%2FF1FN14NK

# Methods

## Determine total crop grown in each region
This should be run if there are changes to the MAPSPAM production data.

Calculates total crop production using the MAPSPAM data adjusted to 2021 production. These data include all growing systems (intensive, irrigated, etc.) and are in units of metric tonnes.

This is used in subsequent script to estimate the proportion of each crop going to animal (salmon) feed in each country.

```{r}
feed_folder <- file.path(here("prep/02_feed"))

library(here)
library(tidyverse)
library(countrycode)
library(strex)
library(terra)

source("src/directories.R")

focus_year = 2021

```

Extract by region the crop production from the raster maps.

```{r}

regions <- read_csv(here("data/spatial/output/food_rgns.csv")) # these are regions originally derived from GADM36 datasets, but I poached them from https://knb.ecoinformatics.org/view/doi%3A10.5063%2FF1FN14NK
 
crop_tonnes <- list.files(file.path(gaez_dir, "/scaled_prod_maps_2021"), pattern = "_Total_scaled.tif", full=TRUE)

crop_stack <- terra::rast(crop_tonnes)

names(crop_stack) <- str_after_last(sources(crop_stack), "/")

zones <- terra::rast(file.path(rdsi_raw_data_dir, "food-systems-project/land_eez_rgns.tif")) # taken from https://knb.ecoinformatics.org/view/doi%3A10.5063%2FF1FN14NK

crop_production <- terra::zonal(crop_stack, zones, fun="sum", na.rm=TRUE, progress="text")

crop_production_df <- data.frame(crop_production) %>%
  gather("GAEZ_category", "tonnes_producing_crop", -1) %>%
  rename("ID_0" = land_eez_rgns) %>%
  left_join(regions, by="ID_0") %>%
  mutate(GAEZ_category = gsub("crop_", "", GAEZ_category)) %>%
  mutate(GAEZ_category = gsub("_Total_scaled.tif", "", GAEZ_category)) %>%
  dplyr::select(iso3c_producing=iso3c, GAEZ_category, tonnes_producing_crop) %>%
  mutate(tonnes_producing_crop = ifelse(is.na(tonnes_producing_crop), 0, tonnes_producing_crop)) # replace NAs with 0

```

Check and save: should be about equal to 2021 FAO production

```{r}

check <- crop_production_df %>%
  group_by(GAEZ_category) %>% 
  summarize(our_tonnes = sum(tonnes_producing_crop, na.rm=TRUE)) %>%
  data.frame()
  
fao_translate <- vroom::vroom(here("prep/01_crops_prep/data/GAEZ_names.csv")) %>%
  dplyr::select(GAEZ_category, item_code) %>%
  unique()

fao <- read_csv(here("prep/01_crops_prep/data/FAOSTAT_crop_production_2015_2021.csv")) %>% 
  filter(year %in% c(2019, 2020, 2021)) %>%
  filter(area != "China",
         area_code < 5000) %>%
 # dplyr::select(area, item_code, value, year) %>%
  left_join(fao_translate, by="item_code") %>% 
  group_by(GAEZ_category, year) %>%
  summarise(total = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(GAEZ_category) %>%
  summarize(total = mean(total, na.rm=TRUE)) %>%
  ungroup() %>%
  filter(!is.na(GAEZ_category))
  
compare <- check %>% left_join(fao, by="GAEZ_category") %>%
  mutate(prop = our_tonnes/total)
compare
sum(compare$our_tonnes, na.rm=TRUE)/sum(compare$total, na.rm = TRUE) # oilpalmfruit, millet, and stimulants are a little problematic (stimulants is the worst ~50%...)

write_csv(crop_production_df, here("prep/02_feed/data/GAEZcrop_production.csv"))

```
