---
title: "Deconstructing trade"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Summary

This script should be updated if any of the following files have been updated:
* FAO Detailed Trade Matrix
* FAO Food Balance Sheets
* Crop production data 

This summarizes FAO trade data by GAEZ crop category (vs. FAO crop category) to get a dataset that describes for each country the country of origin for each crop. This is presented as a proportion so for each country/crop the total will sum to 1.  

Output: feed_consumption/data/FAO_GAEZcrop_trade_data.csv"

Here, we apply a filter that for any country that produces <1000 tonnes of a raw material, we assume that if they are trading it, this is low likelihood of being domestically produced. This is the assumption that to be attractive and competitive as an exporter globally, you need economies of scale. 

## Data sources: 

**Reference**: Grogan, D., Frolking, S., Wisser, D. et al. Global gridded crop harvested area, production, yield, and monthly physical area data circa 2015. Sci Data 9, 15 (2022). https://doi.org/10.1038/s41597-021-01115-2

**Downloaded**: 07/14/2023

**Description**: "Here we provide an update to global gridded annual and monthly crop datasets. This new dataset uses the crop categories established by the Global Agro-Ecological Zones (GAEZ) Version 3 model, which is based on the Food and Agricultural Organization of the United Nations (FAO) crop production data. We used publicly available data from the FAOSTAT database as well as GAEZ Version 4 global gridded dataset to generate circa 2015 annual crop harvested area, production, and yields by crop production system (irrigated and rainfed) for 26 crops and crop categories globally at 5-minute resolution. We additionally used available data on crop rotations, cropping intensity, and planting and harvest dates to generate monthly gridded cropland data for physical areas for the 26 crops by production system. These data are in standard georeferenced gridded format, and can be used by any global hydrology, land surface, or other earth system model that requires gridded annual or monthly crop data inputs."

**Download Link**: https://gaez.fao.org/pages/data-access-download

**Reference**: FAOSTAT (2023) Food Balances (2010-). Available at: http://www.fao.org/faostat/en/#data/FBS (Accessed: 16 March 2023).

**Downloaded**: 03/16/2023


**Reference**: FAOSTAT (2023) Detailed Trade Matrix. FAO. Available at: http://www.fao.org/faostat/en/#data/TM (Accessed: 16 March 2023).

**Downloaded**: 03/16/2023

```{r}

library(here)
library(tidyverse)
library(countrycode)
library(janitor)


source(here("src/directories.R"))


```

Get tonnes production from GAEZ data adjusted to 2021 production.  These data include all growing systems (intensive, irrigated, etc.) and are in units of metric tonnes. This is used to adjust some of the trade values and for gapfilling when trade data are not available. 

`GAEZcrop_production.csv` created in 1.1.

```{r}

crop_production_df <- read_csv( here("prep/02_feed/data/GAEZcrop_production.csv")) # made in the last script

crop_production_gf <- crop_production_df %>%
  group_by(GAEZ_category) %>%
  mutate(global_tonnes = sum(tonnes_producing_crop, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(prop_global_tonnes = ifelse(tonnes_producing_crop==0 & global_tonnes==0, 0, tonnes_producing_crop/global_tonnes))

summary(crop_production_gf)

crop_production_gf <- crop_production_gf %>%
  dplyr::select(iso3c_producing, GAEZ_category, imports_per = prop_global_tonnes, tonnes_producing_crop) 

```

# Production vs. import

Consolidate data to determine the proportional amount of local production and total imports for each crop and country.

Here we take the file `01_crops_prep/data/GAEZ_names.csv` and edit it to only include salmon feed primary and intermediate product items that we include in our diets (e.g., include both Wheat and Gluten, wheat). We save this to `02_feed/data/GAEZ_names_FEED.csv`

```{r}
mapspam_feed_products <- read_csv(here("prep/02_feed/data/MapSPAM_to_FAO_v2.csv")) %>%
  rename(FAONAMES_mapspam = FAONAMES) # read in old mapspam categories

feed_products <- read_csv(here("prep/02_feed/data/GAEZ_names_FEED.csv")) %>%
  left_join(mapspam_feed_products, by = c("item_code" = "FAO_item_code")) %>% # match to GAEZ categories
  filter(!is.na(FAO_item_code_NFB)) %>%
    filter(is.na(non_food)) %>%
  filter(!(FAONAMES %in% c("Mustard_seed", "Lupins", "Vetches"))) %>% # filter out duplicates
  dplyr::select(GAEZ_category, FAO_Item_Code_NFB = FAO_item_code_NFB) %>%
  unique()


feed_products[duplicated(feed_products$FAO_Item_Code_NFB), ] # mostly FAO_item_code is unique, but be aware of duplicates: occurs when MAPSPAM reports crops with more specificity (e.g. millet = small millet and pearl millet in MapSPAM)
summary(feed_products)

# FAO_Item_Code_NFB           item
# 1               2575 Cottonseed Oil
# 10              2582 Maize Germ Oil

```

Here we use the food balance sheets to determine the local production vs. imported production of GAEZ crop categories (i.e., soybean, wheat, pulses, etc.)

```{r}

item_codes <- read.csv(file.path(rdsi_raw_data_dir, "fao/FAO_FoodBalanceSheets/d2023/FoodBalanceSheets_E_ItemCodes.csv")) %>%
  clean_names() %>%
  dplyr::select(item_code, item)

fbs <- read.csv(file.path(rdsi_raw_data_dir, "fao/FAO_FoodBalanceSheets/d2023/FoodBalanceSheets_E_All_Data_(Normalized).csv"), fileEncoding = "ISO-8859-1") %>%
  filter(Year == 2020) %>% # have to use 2020 here because fbs doesn't have 2021 yet... 
  clean_names() %>%
  dplyr::select(consuming_area_code = area_code, consuming_country = area, element, FAO_Item_Code_NFB = item_code, item, value) %>%
  filter(element %in% c("Domestic supply quantity", "Export Quantity", "Feed", "Import Quantity", "Production", "Stock Variation"))
# %>% # filter for relevant element
# left_join(item_codes, by = c("FAO_Item_Code_NFB" = "item_code"))

## provides China average that we want to delete
dim(fbs)
fbs <- fbs %>%
  filter(!(consuming_country %in% "China")) 
dim(fbs)

# food balance sheet contains some derived oil products, these result in double counting because they are already counted in the primary product
discards <- read_csv(here("prep/02_feed/data/FAOSTAT_FBS_Items_discard_30-9-2024.csv")) %>% # made by hand
  filter(cut=="yes") %>%
  pull(`Item Code`)

check <- filter(fbs, FAO_Item_Code_NFB %in% discards)
table(check$item) # this should be non-crop items and Oils (counted in main product already)


fbs <- fbs %>% 
  filter(!(FAO_Item_Code_NFB %in% discards),
         consuming_area_code < 5000)
table(fbs$item)

# make sure that all relevant product codes match up
setdiff(feed_products$FAO_Item_Code_NFB, fbs$FAO_Item_Code_NFB) #
missing <- setdiff(fbs$FAO_Item_Code_NFB, feed_products$FAO_Item_Code_NFB)
missing2 <- filter(fbs, FAO_Item_Code_NFB %in% missing) %>% 
  dplyr::select(FAO_Item_Code_NFB, item) %>%
  unique() %>% 
  data.frame()
missing2 # this is fine, because we don't include these feed items in our diet scenarios

# convert products to GAEZ crop and combine
fbs <- fbs %>%
  filter(FAO_Item_Code_NFB %in% feed_products$FAO_Item_Code_NFB) %>%
  left_join(feed_products, by = "FAO_Item_Code_NFB") %>%
  group_by(consuming_area_code, consuming_country, GAEZ_category, element) %>%
  summarize(value = sum(value, na.rm=TRUE)) %>%
  ungroup() 

# some cleaning
local_prod <- fbs %>%
  spread(element, value) %>%
  rename(supply = `Domestic supply quantity`, imports=`Import Quantity`, exports=`Export Quantity`, production=Production) %>%
  replace(is.na(.), 0) %>%
 # filter(production > 1000) %>%
  mutate(local_crop_prod_tonnes = production - exports) %>%            #local crop production for use within country 
  mutate(supply2 = production+imports-exports) %>%     # a version of supply without stock variation variable
  mutate(local_crop_prod_tonnes = ifelse(local_crop_prod_tonnes < 0, 0, local_crop_prod_tonnes)) %>%
  mutate(local_crop_prod_per = ifelse(local_crop_prod_tonnes == 0, 0, local_crop_prod_tonnes/supply2)) %>%
  mutate(external_crop_prod_per = ifelse(supply2 == 0, 0, imports/supply2)) %>%
  mutate(external_crop_prod_per = ifelse(external_crop_prod_per < 0, 0, external_crop_prod_per)) %>%
      mutate(external_crop_prod_per = ifelse(external_crop_prod_per > 1, 1, external_crop_prod_per)) %>%
  dplyr::select(consuming_area_code, consuming_country, GAEZ_category, local_crop_prod_per, external_crop_prod_per)

summary(local_prod)
filter(local_prod, consuming_country == "Sudan" & GAEZ_category == "Maize")
filter(local_prod, consuming_country == "Myanmar")

```


Get iso3c codes

```{r}

local_prod_iso <- local_prod %>%
    mutate( # consuming_country = iconv(consuming_country, "UTF-8", "UTF-8",sub=''),
         consuming_country = ifelse(consuming_country == "Eswatini", "Swaziland", consuming_country),
         consuming_country = ifelse(consuming_country == "Netherlands Antilles (former)", "Bonaire, Sint Eustatius and Saba", consuming_country),
         consuming_country = ifelse(consuming_country == "French Guyana", "French Guiana", consuming_country),
         consuming_country = ifelse(consuming_country == "China, mainland", "China", consuming_country))

## add standardized country names
local_prod_iso$iso3c_consuming <- countrycode(as.character(local_prod_iso$consuming_country), origin="country.name", destination = "iso3c")

## make sure all countries have an iso3c code
check <- filter(local_prod_iso, is.na(iso3c_consuming)) 
unique(check$consuming_country) # Turkey... 

local_prod_iso <- local_prod_iso %>%
  mutate(iso3c_producing = iso3c_consuming)

summary(local_prod_iso)

```


Join GAEZ crop production data with fbs data.  If there is less than 1000 tonnes of local production according to GAEZ,
local production will be zero and external production will be one.  And, if both local_crop_prod_per and external_crop_prod_per are 0, assuming that external_crop_prod_per == 1.

```{r}

local_prod_gf <- left_join(local_prod_iso, crop_production_df, by=c("iso3c_producing", "GAEZ_category")) %>%
  mutate(local_crop_prod_per = ifelse(tonnes_producing_crop < 1000, 0, local_crop_prod_per)) %>%
  mutate(external_crop_prod_per = ifelse(tonnes_producing_crop < 1000, 1, external_crop_prod_per)) %>%
  mutate(external_crop_prod_per = ifelse(local_crop_prod_per ==0 & external_crop_prod_per==0, 1, external_crop_prod_per)) %>%
  dplyr::select(iso3c_consuming, GAEZ_category, local_crop_prod_per, external_crop_prod_per)


summary(local_prod_gf)
filter(local_prod_gf, iso3c_consuming=="SDN")
filter(local_prod_gf, iso3c_consuming=="IRQ") %>% data.frame()
filter(local_prod_gf, iso3c_consuming == "ETH" & GAEZ_category == "Rapeseed")
filter(local_prod_gf, iso3c_consuming == "MMR")

tmp <- local_prod_gf %>%
  rowwise() %>%
  mutate(total = local_crop_prod_per + external_crop_prod_per)
summary(tmp)

```


# Import data
FAO provides a trade matrix showing tonnes of each crop imported into each country.

FAO codes:
```{r}

feed_products_full <- read_csv(here("prep/02_feed/data/GAEZ_names_FEED.csv")) %>%
  left_join(mapspam_feed_products, by = c("item_code" = "FAO_item_code")) %>% 
  filter(!is.na(GAEZ_category))


feed_products <- feed_products_full %>% # match to GAEZ categories
    filter(is.na(non_food)) %>%
  dplyr::select(GAEZ_category, FAO_Item_Code = item_code) %>%
  dplyr::filter(!is.na(FAO_Item_Code)) %>%
  unique() 

```

Get FAO import data and convert to GAEZ crops:

```{r}

import_data <- read_csv(file.path(fao_dir, "FAO_detailed_trade/d2023/Trade_DetailedTradeMatrix_E_All_Data_(Normalized).csv")) # takes awhile to load.. this is a massive csv (~7gb)

# cut "Unspecified Area" for exporting countries.  This has the effect of spreading the imports among the countries with existing data
imports <- import_data %>%
  filter(Year == 2020) %>%
  filter(!(`Partner Countries` %in% c("Unspecified Area"))) %>%
    filter(!(`Partner Countries` %in% c("China"))) %>%                 # combination of all china regions
  dplyr::filter(Element == "Import Quantity") %>%
  dplyr::select(consuming_area_code = `Reporter Country Code`,
                consuming_country = `Reporter Countries`,
                producing_area_code = `Partner Country Code`, 
                producing_country = `Partner Countries`,
                FAO_Item_Code = `Item Code`, Element, Value, Item)

test_imps <- imports %>% filter(consuming_country == "Norway", str_detect(Item, "Wheat|wheat"))
# sum(test_imps$Value)
# [1] 293578.5

test <- imports %>% distinct(Item)
## ok so the trade matrix only has
# Wheat
# Bran of wheat
# Wheat and meslin flour
# Buckwheat
# Germ of wheat
# Flour of buckwheat
# and the fbs only has "Wheat and products" so that is what we'll have to work with. 

# make sure that all relevant product codes match up
extra_products <- setdiff(feed_products$FAO_Item_Code, imports$FAO_Item_Code) 
extra_products_description <- feed_products_full %>%
  dplyr::filter(item_code %in% extra_products)
unique(extra_products_description$product_description_FAOSTAT)
# in the master list, but not in the trade data.. i guess there is nothing we can do about this unfortunately. 


extra_imports <- setdiff(imports$FAO_Item_Code, feed_products_full$item_code)

# These are in the trade data, but not our master list, these are all products that are not considered feed items in our scenarios, so this is ok. 

test <- import_data %>%
  dplyr::select(Item, `Item Code`) %>%
  filter(`Item Code` %in% extra_imports) %>%
  unique() %>% data.frame()

# convert products to MAPSPAM categories 
imports <- imports %>%
  filter(FAO_Item_Code %in% feed_products$FAO_Item_Code) %>%
  left_join(feed_products, by = "FAO_Item_Code") %>%
  dplyr::group_by(consuming_area_code, consuming_country, producing_area_code, producing_country, GAEZ_category) %>%
  summarize(Value = sum(Value)) %>%
  ungroup()

```

Assign iso3c region ids:

```{r}

# not catching Eswatini
imports <- imports %>%
  mutate(producing_country = ifelse(producing_country == "Eswatini", "Swaziland", producing_country),
                producing_country = ifelse(grepl("Ivoire", imports$producing_country), "Ivory Coast", producing_country),
         consuming_country = ifelse(consuming_country == "Eswatini", "Swaziland", consuming_country),
                consuming_country = ifelse(grepl("Ivoire", imports$consuming_country), "Ivory Coast", consuming_country)) %>%
  mutate_if(is.character, utf8::utf8_encode)



## add standardized country names
imports$iso3c_producing <- countrycode(as.character(imports$producing_country), origin="country.name", destination = "iso3c")
imports$iso3c_consuming <- countrycode(as.character(imports$consuming_country), origin="country.name", destination = "iso3c")


imports <- imports %>%
  mutate(iso3c_producing = ifelse(imports$producing_country == "Netherlands Antilles (former)", "BES", iso3c_producing)) %>%
    mutate(iso3c_consuming = ifelse(imports$consuming_country == "Netherlands Antilles (former)", "BES", iso3c_consuming)) %>%
    mutate(iso3c_producing = ifelse(is.na(iso3c_producing), "TUR", iso3c_producing)) %>%
    mutate(iso3c_consuming = ifelse(is.na(iso3c_consuming), "TUR", iso3c_consuming)) %>%
  mutate(producing_country = ifelse(is.na(producing_country), "Turkey", producing_country),
         consuming_country = ifelse(is.na(consuming_country), "Turkey", consuming_country))

## make sure all countries have an iso3c code
filter(imports, is.na(iso3c_consuming))
filter(imports, is.na(iso3c_producing))

## make sure no duplicated regions (should all be one)
tmp <- imports %>%
  mutate(check = paste(consuming_area_code, iso3c_consuming, sep="_")) %>%
  group_by(iso3c_consuming) %>%
  summarize(check_n = length(unique(check)))
data.frame(tmp)
sum(tmp$check_n>1)

```

Clean the imports data
```{r}

imports <- imports %>% 
   dplyr::select(iso3c_consuming, iso3c_producing, GAEZ_category, imports_tonnes = Value) %>%   
  filter(iso3c_producing != iso3c_consuming) # just to be sure

filter(imports, iso3c_consuming=="NOR" & GAEZ_category=="Wheat")

```

If there is <1000 tonnes of production for a country/crop according to GAEZ, then there will be no exports for that country/crop.

```{r}

imports <- left_join(imports, crop_production_df, by=c("iso3c_producing", "GAEZ_category")) %>%
  mutate(tonnes_producing_crop = ifelse(is.na(tonnes_producing_crop), 0, tonnes_producing_crop)) %>%
  mutate(correction = ifelse(tonnes_producing_crop < 1000, 0, 1)) %>%
  mutate(imports_tonnes = imports_tonnes*correction) %>%
  dplyr::select(iso3c_consuming, iso3c_producing, GAEZ_category, imports_tonnes)

filter(imports, iso3c_consuming=="AFG" & GAEZ_category=="Pulses")
```


Determine which fbs regions/crops should be gapfilled because there is no fbs data but there is trade matrix data.
In these cases, we use the proportion of imported to production to determine fbs proportional values.

```{r}

imports_rgn_crop_list <- imports %>%
  group_by(iso3c_consuming, GAEZ_category) %>%
  summarize(total_import_tonnes = sum(imports_tonnes)) %>%
  ungroup() %>%
  mutate(major_group_imports = paste(iso3c_consuming, GAEZ_category))

local_prod_rgn_crop_list <- local_prod_gf %>%
  mutate(major_group_prod = paste(iso3c_consuming, GAEZ_category))

missing_local_prod <- setdiff(imports_rgn_crop_list$major_group_imports, local_prod_rgn_crop_list$major_group_prod)


crop_production_tonnes <- crop_production_df %>%
  rename(iso3c_consuming = iso3c_producing)

more_local_gf <- filter(imports_rgn_crop_list, major_group_imports %in% missing_local_prod) %>%
  left_join(crop_production_tonnes, by = c("iso3c_consuming", "GAEZ_category")) %>%
  mutate(local_crop_prod_per = tonnes_producing_crop/(tonnes_producing_crop + total_import_tonnes)) %>%
  mutate(external_crop_prod_per = total_import_tonnes/(tonnes_producing_crop + total_import_tonnes)) %>%
  mutate(local_crop_prod_per = ifelse(is.na(local_crop_prod_per), 0, local_crop_prod_per)) %>%
  mutate(external_crop_prod_per = ifelse(is.na(external_crop_prod_per), 1, external_crop_prod_per)) %>%
  dplyr::select(iso3c_consuming, GAEZ_category, local_crop_prod_per, external_crop_prod_per)

tmp <- more_local_gf %>% mutate(total=local_crop_prod_per + external_crop_prod_per) 
summary(tmp)

local_prod_gf <- rbind(local_prod_gf, more_local_gf)


```


Calculate the proportion of crop coming from each country

```{r}

imports_prop <- imports %>%
  group_by(iso3c_consuming, GAEZ_category) %>%
  mutate(imports_all_tonnes = sum(imports_tonnes)) %>%
  ungroup() %>%
  mutate(imports_per = ifelse((imports_tonnes + imports_all_tonnes) == 0, 0, imports_tonnes/imports_all_tonnes)) %>%
  dplyr::select(-imports_tonnes, -imports_all_tonnes)


filter(imports_prop, iso3c_consuming == "LVA" & GAEZ_category == "Pulses")

```


Determine which trade regions/crops should be gapfilled because there is no trade data but there is fbs data.
In these cases, we use every country's proportion of global production as the trade data.

```{r}

imports_rgn_crop_list <- imports_prop %>%
  mutate(major_group_imports = paste(iso3c_consuming, GAEZ_category)) %>%
  dplyr::select(major_group_imports) %>% unique()

local_prod_rgn_crop_list <- local_prod_gf %>%
  mutate(major_group_prod = paste(iso3c_consuming, GAEZ_category))

missing_trade <- setdiff(local_prod_rgn_crop_list$major_group_prod, imports_rgn_crop_list$major_group_imports)

imports_new_data <- filter(local_prod_rgn_crop_list, major_group_prod %in% missing_trade) %>%
  dplyr::select(iso3c_consuming, GAEZ_category)

imports_new_data_expand <- left_join(imports_new_data, crop_production_gf) %>%
  dplyr::filter(iso3c_consuming != iso3c_producing) %>%
  dplyr::select(-tonnes_producing_crop)

imports_prop_gf <- rbind(imports_prop, imports_new_data_expand)


```

Figure out which countries crops have neither trade or fbs data.  We will just use the global proportions in this case.  

```{r}

mstr_imports <- imports_prop_gf %>%
  mutate(rgn_crop_imports = paste(iso3c_consuming, GAEZ_category)) %>%
  dplyr::select(rgn_crop_imports) %>%
  unique()

# full region list
mast_rgns_crops <- crop_production_gf %>%
  mutate(rgn_crops = paste(iso3c_producing, GAEZ_category)) %>%
  dplyr::select(rgn_crops) %>%
  unique()
  
missing_rgn_crops <- data.frame(rgn_crops_missing = setdiff(mast_rgns_crops$rgn_crops, mstr_imports$rgn_crop_imports)) %>%
  separate(rgn_crops_missing, c("iso3c_consuming", "GAEZ_category"))

imports_new_data_expand <- left_join(missing_rgn_crops, crop_production_gf %>% dplyr::select(-tonnes_producing_crop)) 

imports_prop_gf <- rbind(imports_prop_gf, imports_new_data_expand)

```



Join local and import proportions.

For now this is just temporary in order to identify and then 
fix cases where there is no import data, but external_crop_production >0, in these cases we will use global proportions. 

```{r}

import_export_test <- imports_prop_gf %>% 
  full_join(local_prod_gf, by = c("iso3c_consuming", "GAEZ_category")) 
  
mismatch <- import_export_test %>%
  group_by(iso3c_consuming, GAEZ_category) %>%
  mutate(total_imports = sum(imports_per, na.rm=TRUE)) %>%
  filter(total_imports == 0 & external_crop_prod_per >0)

# fix this one which has a different pattern:
filter(mismatch, local_crop_prod_per > 0)

## repeat after removing the strange one
import_export_test <- imports_prop_gf %>% 
  full_join(local_prod_gf, by = c("iso3c_consuming", "GAEZ_category"))
  
mismatch <- import_export_test %>%
  group_by(iso3c_consuming, GAEZ_category) %>%
  mutate(total_imports = sum(imports_per, na.rm=TRUE)) %>%
  filter(total_imports == 0 & external_crop_prod_per >0) %>%
  ungroup()

filter(mismatch, local_crop_prod_per > 0)
summary(mismatch)

## now get the gapfilled data
mismatch <- mismatch %>%
  mutate(rgn_crops = paste(iso3c_consuming, GAEZ_category)) %>%
  dplyr::select(rgn_crops) %>%
  unique()
  
mismatch_rgn_crops <- mismatch %>%
  separate(rgn_crops, c("iso3c_consuming", "GAEZ_category"))

imports_new_data_expand <- left_join(mismatch_rgn_crops, crop_production_gf) %>%
  filter(iso3c_consuming != iso3c_producing) %>%
  dplyr::select(-tonnes_producing_crop)

## remove old values from data
imports_prop_gf <- imports_prop_gf %>%
  mutate(rgn_crop = paste(iso3c_consuming, GAEZ_category)) %>%
  filter(!(rgn_crop %in% mismatch$rgn_crops)) %>%
  dplyr::select(-rgn_crop)

imports_prop_gf <- rbind(imports_prop_gf, imports_new_data_expand)

```

Combine for real now the trade data and fbs proportion of imports data.

```{r}

import_export <- imports_prop_gf %>% 
  full_join(local_prod_gf, by = c("iso3c_consuming", "GAEZ_category")) 

  summary(import_export)

import_export_gf <- import_export %>%
  mutate(local_crop_prod_per = ifelse(is.na(local_crop_prod_per), 0, local_crop_prod_per)) %>%
  mutate(external_crop_prod_per = ifelse(is.na(external_crop_prod_per), 1, external_crop_prod_per))

 
 summary(import_export_gf)
```


Multiply country imports by proportion of imports (these all sum to one regardless of what proportion of supply are from imports) to reflect the total proportion of imports into the country.

```{r}

import_export_gf <- import_export_gf %>%
  mutate(prop_of_supply = imports_per * external_crop_prod_per) %>%
  dplyr::select(iso3c_consuming, iso3c_producing, GAEZ_category, prop_of_supply)

data.frame(filter(import_export_gf, iso3c_consuming== "USA")) %>%
  arrange(GAEZ_category)

tmp <- filter(import_export_gf, iso3c_consuming== "USA") %>%
  filter(GAEZ_category=="Cotton")
sum(tmp$prop_of_supply)  
```


Add in the within country proportion

```{r}

# the producing and consuming country are the same in this case:
local_prod <- local_prod_gf %>%
  mutate(iso3c_producing = iso3c_consuming, 
         prop_of_supply = local_crop_prod_per) %>%
         dplyr::select(iso3c_consuming, iso3c_producing, GAEZ_category, prop_of_supply)

filter(local_prod, iso3c_consuming=="BLZ") %>%data.frame

trade_data <- rbind(import_export_gf, local_prod) %>% 
  filter(GAEZ_category %in% c(unique(imports$GAEZ_category)))

filter(trade_data, iso3c_consuming == "FJI" & iso3c_producing == "FJI") %>% data.frame()
filter(trade_data, iso3c_consuming == "ARM" & iso3c_producing == "ARM") %>% data.frame()
filter(trade_data, iso3c_consuming == "CHN" & iso3c_producing == "CHN") %>% data.frame()

```

## Final data checks and save


```{r}
# These were gapfilled earlier and should be empty.
regions <- read_csv(here("data/spatial/output/region_ids.csv"))
setdiff(regions$iso3c, trade_data$iso3c_consuming)
setdiff(trade_data$iso3c_consuming, regions$iso3c)

## make sure we got all the values for each countries within country production
within_country <- trade_data %>%
  filter(iso3c_consuming==iso3c_producing) %>%
  dplyr::select(iso3c_consuming) %>%
  unique()

missing_country <- setdiff(regions$iso3c, within_country$iso3c_consuming)
missing_country

# make sure each consuming/providing/crop is unique
check <- trade_data %>%
  mutate(unique=paste(iso3c_consuming, iso3c_producing, GAEZ_category))

sum(table(check$unique)>1) # should be zero...if not, figure out why

# Check that sums within a country crop sum to 1
check <- trade_data %>%
  group_by(iso3c_consuming, GAEZ_category) %>%
  summarize(total = round(sum(prop_of_supply), 2))


data.frame(dplyr::filter(check, total!=1)) # good 0

write_csv(trade_data, here("prep/02_feed/data/FAO_GAEZcrop_trade_data_10-02-2024.csv"))

test2 <- read.csv(here("prep/02_feed/data/FAO_GAEZcrop_trade_data_10-02-2024.csv")) 

## most everything looks the same... 

test <- test2 %>% 
  filter(iso3c_consuming == "NOR" | iso3c_producing == "NOR")


```

