---
title: "Forage fish and trimmings marine footprint disturbance (km2)"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Overall approach

Using the marine footprint approach outlined in Cashion et al https://onlinelibrary.wiley.com/doi/full/10.1111/faf.12222 based off of the primary production required for catch in a given cell.

We need to calculate this separately for fishmeal and oil as we are interested in the allocation of pressures to ingredients
  
```{r, setup}

library(tidyverse)
library(here)
library(terra)
library(sf)
library(data.table)
library(dtplyr)
library(future)
library(furrr)
#library(doParallel)
library(rfishbase)

source(here("src/directories.R"))
source(here("src/spatial.R"))

select <- dplyr::select

```


Equal area projection resources
```{r}

gall_peters <- "+proj=cea +lon_0=0 +x_0=0 +y_0=0 +lat_ts=45 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
#raster template 
base_rast <- rast(res=0.5)
ext(base_rast) <- c(-180, 180, -90, 90)

eq_area_rast <- project(base_rast, gall_peters) 
res(eq_area_rast) <- 10000

values(eq_area_rast) <- 1:ncell(eq_area_rast)

```


MARINE FOOTPRINT


Create a cell area xyz file for 0.5 res wgs84 raster

```{r}

cell_area <- terra::as.data.frame(cellSize(rast(res=0.5), unit="km"), xy=TRUE) |> rename(LonCentre = x, LatCentre = y)

```



Import net primary productivity data
```{r}


#make npp into an xy file

mean_npp <- rast(here("../npp_synthesis/data/npp_2013_2022_mean.tif"))
median_npp <- rast(here("../npp_synthesis/data/npp_2013_2022_median.tif"))
min_npp <- rast(here("../npp_synthesis/data/npp_2013_2022_min.tif"))
max_npp <- rast(here("../npp_synthesis/data/npp_2013_2022_max.tif"))

# mean_npp <- rast(here("prep/02_feed/data/npp_old/annual_mean_npp_2019_gf_wgs.tiff"))
# median_npp <- rast(here("prep/02_feed/data/npp_old/annual_median_npp_2019_gf_wgs.tiff"))
# min_npp <- rast(here("prep/02_feed/data/npp_old/annual_min_npp_2019_gf_wgs.tiff"))
# max_npp <- rast(here("prep/02_feed/data/npp_old/annual_max_npp_2019_gf_wgs.tiff"))


#mean 
mean_npp_xyz <- terra::as.data.frame(mean_npp, xy = TRUE) |> 
  rename(mean_npp_t_km2_yr=mean) |> 
  rename(LonCentre = x, 
         LatCentre = y)

#median
median_npp_xyz <- terra::as.data.frame(median_npp, xy = TRUE) |> 
  rename(median_npp_t_km2_yr=mean) |> 
  rename(LonCentre = x, 
         LatCentre = y)

#min
min_npp_xyz <- terra::as.data.frame(min_npp, xy = TRUE) |> 
  rename(min_npp_t_km2_yr=mean) |> 
  rename(LonCentre = x, 
         LatCentre = y)

#max
max_npp_xyz <- terra::as.data.frame(max_npp, xy = TRUE) |> 
  rename(max_npp_t_km2_yr=mean) |> 
  rename(LonCentre = x, 
         LatCentre = y)

```

Import ecosystem transfer-efficiency

```{r}

ecosystem_te <- read_csv(file.path(rdsi_raw_data_dir, "cashion/ecosystem_te.csv")) |> select(TE, fao_area_code) |> group_by(fao_area_code) |> summarise(te=mean(TE, na.rm = TRUE)) |> mutate(te= te/100)


```


Get forage fish trophic levels - needed for calculating the primary productivity required

```{r}

#Separate the catch records that have Genus and species.
forage_spp <- readRDS(file.path(watson_dir, "v5.0/int/all_catch_foragefish.rds")) |> 
  filter(forage_id == 1) %>%
  mutate(binomial = stringi::stri_count(TaxonName, regex="\\S+")) |> 
  filter(binomial==2) |>  
  pull(TaxonName) |> 
  unique() 

# this_sp <- forage_spp[[1]]

taxa_info <- load_taxa()

binomial_tls <- rfishbase::ecology(species_list =  forage_spp, server = getOption("fishbase")) |>
  select(Species, DietTroph, FoodTroph) |> 
  mutate(DietTroph = if_else(is.na(DietTroph), true = FoodTroph, false = DietTroph)) |> 
  drop_na(DietTroph) |> 
  select(-FoodTroph) |> 
  distinct() %>% #both extracted manually from fishbase as not coming up from app search
  left_join(taxa_info, by = "Species")

# get family means

family_means <- binomial_tls %>%
  group_by(Family) %>%
  summarise(DietTroph = mean(DietTroph, na.rm = TRUE)) %>%
  ungroup()

genus_means <- binomial_tls %>%
  group_by(Genus) %>%
  summarise(DietTroph = mean(DietTroph, na.rm = TRUE)) %>%
  ungroup()
  
#Add back the Genus only or family records and match to any species with the same Genus

(forage_higherorder <- readRDS(file.path(watson_dir, "v5.0/int/all_catch_foragefish.rds")) |> filter(forage_id == 1) %>% mutate(binomial = stringi::stri_count(TaxonName, regex="\\S+")) |> filter(binomial==1) |>  pull(TaxonName) |> unique()
)

forage_nonbinomial_tls <- tibble(TaxonName = forage_higherorder) |> 
  left_join(family_means, by = c("TaxonName" = "Family")) %>% 
  left_join(genus_means, by = c("TaxonName" = "Genus")) %>% 
  mutate(DietTroph = ifelse(is.na(DietTroph.x), DietTroph.y, DietTroph.x)) %>%
  dplyr::select(TaxonName, DietTroph) %>% 
  mutate(DietTroph = case_when(TaxonName %in% 
                                 c("Engraulidae", "Clupeiformes", "Stolephorus") ~ binomial_tls |> 
                                 filter(grepl("Engraulis",Species)) |> 
                                 pull(DietTroph) |> 
                                 mean(),
                                str_detect(TaxonName, "Scomb") ~ binomial_tls |> 
                                 filter(grepl("Scomber", Species)) |> 
                                 pull(DietTroph) |> 
                                 mean(),
                               TaxonName %in% c("Sardinella") ~ binomial_tls |> 
                                 filter(grepl("Sardinella", Species)) |> 
                                 pull(DietTroph) |> 
                                 mean(),
                               TaxonName %in% c("Rastrelliger") ~ binomial_tls |> 
                                 filter(grepl("Rastrelliger", Species)) |> 
                                 pull(DietTroph) |> 
                                 mean(),
                               TaxonName %in% c("Trachurus") ~ binomial_tls |> 
                                 filter(grepl("Trachurus", Species)) |> 
                                 pull(DietTroph) |> 
                                 mean(),
                               TaxonName %in% c("Harengula", "Clupeidae", "Clupeoids") ~ binomial_tls |> 
                                 filter(grepl("Clupea", Species)) |> 
                                 pull(DietTroph),
                               TaxonName %in% c("Atherinidae") ~ binomial_tls |> 
                                 pull(DietTroph) |> 
                                 mean(),
         TRUE ~ DietTroph)) |> 
  add_row(TaxonName = "Euphausia superba", DietTroph = 2.25) |> 
    add_row(TaxonName = "Clupea pallasii pallasii", DietTroph = 3.16) 
                               
                          
     
                               
forage_tls <- 
  bind_rows(binomial_tls |> rename(TaxonName = Species),
          forage_nonbinomial_tls)


test <- readRDS(file.path(watson_dir, "v5.0/int/all_catch_foragefish.rds")) %>% 
  filter(forage_id == 1)

missing_spp <- setdiff(test$TaxonName, forage_tls$TaxonName)

# missing 8 spp...


missing_tls <- tibble(TaxonName = missing_spp, DietTroph = NA)

# lets fill in the species with NA DietTroph

ga_troph <- mean(forage_tls$DietTroph, na.rm = TRUE)

forage_tls_gf <- bind_rows(forage_tls, missing_tls) %>%
 # filter(is.na(DietTroph)) %>%
  mutate(DietTroph = case_when(
    TaxonName == "Holocentridae" ~ 3.1, # https://www.fishbase.se/summary/1061
    TaxonName == "Scorpaenidae" ~ binomial_tls |> 
                                 filter(Order == "Perciformes/Scorpaenoidei") |> 
                                 pull(DietTroph) |> 
                                 mean(), 
    TaxonName == "Ostraciidae" ~ binomial_tls |> 
                                 filter(Order == "Tetraodontiformes") |> # https://en.wikipedia.org/wiki/Ostraciidae
                                 pull(DietTroph) |> 
                                 mean(), 
    TaxonName == "Pomacanthidae" ~ binomial_tls |> 
                                 filter(Order == "Perciformes/Notothenioidei") |> # https://en.wikipedia.org/wiki/Pomacanthidae
                                 pull(DietTroph) |> 
                                 mean(), 
    TaxonName == "Moridae" ~ binomial_tls |> 
                                 filter(Order == "Gadiformes") |> # https://en.wikipedia.org/wiki/Moridae
                                 pull(DietTroph) |> 
                                 mean(), 
    TaxonName == "Tetraodontidae" ~ binomial_tls |> 
                                 filter(Order == "Tetraodontiformes") |> 
                                 pull(DietTroph) |> 
                                 mean(), 
     TaxonName == "Exocoetidae" ~ binomial_tls |> 
                                 filter(Order == "Beloniformes") |> # https://en.wikipedia.org/wiki/Flying_fish
                                 pull(DietTroph) |> 
                                 mean(), 
     TaxonName == "Batrachoididae" ~ ga_troph,
    TaxonName == "Gobiidae" ~ 3.2, # https://www.fishbase.se/summary/Gobiosoma-robustum
    TaxonName == "Cynoglossidae" ~ binomial_tls |> 
                                 filter(Order == "Pleuronectiformes") |> # https://en.wikipedia.org/wiki/Tonguefish
                                 pull(DietTroph) |> 
                                 mean(), 
    TaxonName == "Leiognathidae" ~ binomial_tls %>%
      filter(str_detect(Order, "Perciformes")) %>%
      pull(DietTroph) %>%
      mean(), 
    TaxonName == "Caesionidae" ~ binomial_tls %>%
      filter(str_detect(Order, "Perciformes")) %>% # https://en.wikipedia.org/wiki/Caesionidae
      pull(DietTroph) %>%
      mean(), 
    TaxonName == "Chlorophthalmidae" ~ binomial_tls %>%
      filter(str_detect(Order, "Aulopiformes")) %>% # https://en.wikipedia.org/wiki/Greeneye
      pull(DietTroph) %>%
      mean(), 
    TaxonName == "Trachinidae" ~ ga_troph,
    TaxonName == "Trachichthyidae" ~ ga_troph,
    TaxonName == "Cottidae" ~ ga_troph, 
    TaxonName == "Apogonidae" ~ ga_troph,
    TaxonName == "Oblada melanura" ~ binomial_tls %>%
      filter(str_detect(Family, "Sparidae")) %>% # https://en.wikipedia.org/wiki/Greeneye
      pull(DietTroph) %>%
      mean(),
    TaxonName == "Liza saliens" ~ ga_troph,
    TaxonName == "Ammodytes personatus" ~ 3.1, # https://www.fishbase.se/summary/Ammodytes-personatus.html
    TaxonName == "Clupanodon thrissa" ~ 3.1, # https://www.fishbase.se/summary/Clupanodon-thrissa
    TaxonName == "Liza klunzingeri" ~ 2.6, # https://www.fishbase.se/summary/46376
    TaxonName == "Pellona ditchela" ~ 4.0, # https://www.fishbase.se/summary/1641
    TaxonName == "Engraulis capensis" ~ 3.0, # https://www.fishbase.se/summary/Engraulis-capensis.html
    TaxonName == "Lepidoperca pulchella" ~ 3.6, #https://www.fishbase.se/summary/Lepidoperca-pulchella.html
    TRUE ~ DietTroph
  ))

setdiff(test$TaxonName, forage_tls_gf$TaxonName) # character(0) - perfect

write.csv(forage_tls_gf, here("prep/02_feed/data/forage_fish_tls.csv"), row.names = FALSE)
```


Get trimmings species trophic levels  - needed for calculating the primary productivity required


```{r}
trim_catch <-  readRDS(file.path(watson_dir, "v5.0/int/all_catch_trimmingsfish.rds"))

#Separate the catch records that have Genus and species.
trim_spp <- trim_catch |> 
  filter(trimmingsfish == 1) %>%
   mutate(binomial = stringi::stri_count(TaxonName, regex="\\S+")) |> 
   filter(binomial==2) |> 
   pull(TaxonName) |> 
   unique() 



trimmings_binomial_tls <- rfishbase::ecology(species_list =  trim_spp, server = getOption("fishbase")) |> 
  select(Species, DietTroph, FoodTroph) |> 
  mutate(DietTroph = if_else(is.na(DietTroph), true = FoodTroph, false = DietTroph)) |>
  select(-FoodTroph) |> 
  mutate(DietTroph = case_when(Species == "Engraulis capensis" ~ mean(c(3.14,3.12,2.51)), #mean of otehr Engraulis values
                               TRUE ~ DietTroph)) |> 
  distinct() |> 
    drop_na() |> 
    add_row(Species = "Theragra chalcogramma", DietTroph = 3.57) #add Alaska pollack that fell out of the species filter

  
#Add back the Genus only or family records and match to any species with the same Genus

trim_higherorder <- trim_catch |> 
  filter(trimmingsfish == 1) |> 
  mutate(binomial = stringi::stri_count(TaxonName, regex="\\S+")) |> 
  filter(binomial==1) |>  
  pull(TaxonName) |> 
  unique()


trim_nonbinomial_tls <- tibble(TaxonName = trim_higherorder, DietTroph = NA) |> 
  mutate(DietTroph = case_when(TaxonName %in% 
                                 c("Engraulidae") ~ trimmings_binomial_tls |>
                                 filter(grepl("Engraulis",Species)) |> 
                                 pull(DietTroph) |> 
                                 mean(),
                               TaxonName %in% c("Sardinella") ~ trimmings_binomial_tls |> 
                                 filter(grepl("Sardinella", Species)) |> 
                                 pull(DietTroph) |> 
                                 mean(),
                               TaxonName %in% c("Clupeoids") ~ trimmings_binomial_tls %>% 
                                 filter(grepl("Engraulis ringens", Species)) %>% 
                                          pull(DietTroph) %>% 
                                          mean(),
                               TaxonName %in% c("Thunnus") ~ trimmings_binomial_tls |> 
                                 filter(grepl("Thunnus", Species)) |> 
                                 pull(DietTroph) |> 
                                 mean(),
                               TaxonName %in% c("Stolephorus") ~ trimmings_binomial_tls |>
                                 filter(grepl("Engraulis",Species)) |> 
                                 pull(DietTroph) |> 
                                 mean()))
                               
                               
                               
trim_tls <- 
  bind_rows(trimmings_binomial_tls |> rename(TaxonName = Species),
          trim_nonbinomial_tls)


test <- trim_catch %>% 
  filter(trimmingsfish == 1)

missing_spp <- setdiff(test$TaxonName, trim_tls$TaxonName) # character(0)


sort(unique(test$TaxonName))

sort(unique(trim_tls$TaxonName))

write.csv(trim_tls, here("prep/02_feed/data/trimmings_fish_tls.csv"), row.names = FALSE)

```



Attach NPP, TL, data to forage and trimmings catch data and calculate the primary productivity required for the catch, the proportion of the annual PPR that is and tmultiply this proportion by the size of the cell.  


```{r}

# read in tls data
forage_tls_gf <- read.csv(here("prep/02_feed/data/forage_fish_tls.csv"))

#bring in the FAO regions shape file

fao_areas_shp <- read_sf("https://www.fao.org/fishery/geoserver/fifao/ows?service=WFS&request=GetFeature&version=1.0.0&typeName=fifao:FAO_AREAS_CWP_NOCOASTLINE&outputFormat=json") |> filter(F_LEVEL == "MAJOR") |> 
  mutate(F_AREA = as.numeric(F_AREA)) |> 
  group_by(F_AREA) |> 
  group_split()


#get xyz file for the FAO areas in 0.5 resolution (catch data resolution)

fao_xyz <- 
  
  map_df(.x = fao_areas_shp, .f = \(this_area){
  
  this_area_shp <- this_area
  
  this_area_raster <- terra::rasterize(x = vect(this_area_shp), y = rast(res = 0.5), field = this_area_shp$F_AREA)
  
  this_area_xyz <- terra::as.data.frame(x = this_area_raster, xy = TRUE)  |> mutate(fao_area_name = this_area_shp$NAME_EN)
  
  return(this_area_xyz)
  
}) |> rename(fao_area_code = layer,
             LonCentre = x,
             LatCentre = y)


foragefish <- readRDS(file.path(watson_dir, "v5.0/int/all_catch_foragefish.rds"))

FOFM_props <- readRDS(here("prep/02_feed/data/embodied_FMFO_forage_rgns.rds"))

watson_cells <- read.csv(file.path(watson_dir, "v5.0/Codes_cells.csv"))

carbon_conversion <- 9

forage_npp <- foragefish %>% 
  left_join(FOFM_props, by = c("iso3c", "TaxonName")) %>%
  filter(!is.na(ingredient)) %>%
  mutate(prop = ifelse(is.na(prop), 0, prop)) %>% # just incase
  mutate(fmfo_embodied_fish = prop*FOFM_catch) %>%
    left_join(fao_xyz) %>%
    left_join(ecosystem_te, by="fao_area_code") |> 
    left_join(forage_tls_gf, by = "TaxonName") |>
    mutate(te = case_when(is.na(te) ~ ecosystem_te |> pull(te) |> mean(),
                          TRUE ~ te)) |> 
    left_join(mean_npp_xyz) |> 
    left_join(median_npp_xyz) |> 
    left_join(min_npp_xyz) |> 
    left_join(max_npp_xyz) |> 
    left_join(cell_area) |> 
    mutate(mean_npp_t_C_yr = mean_npp_t_km2_yr*area) |> 
    mutate(median_npp_t_C_yr = median_npp_t_km2_yr*area) |> 
    mutate(min_npp_t_C_yr = min_npp_t_km2_yr*area) |> 
    mutate(max_npp_t_C_yr = max_npp_t_km2_yr*area) |>
    mutate(ppr_fmfo = (fmfo_embodied_fish/carbon_conversion)*(1/te)^(DietTroph-1)) %>% #T C yr-1
    mutate(ppr_prop_fmfo = ppr_fmfo/mean_npp_t_C_yr) |> 
    mutate(area_fmfo = ppr_prop_fmfo*area) %>%
  group_by(Cell, diet, allocation_type, ingredient, water_col_position) %>%
  summarise(area_fmfo = sum(area_fmfo, na.rm = TRUE),
            fmfo_embodied_fish = sum(fmfo_embodied_fish, na.rm = TRUE)) %>%
  ungroup() %>% 
  left_join(watson_cells) %>%
  dplyr::select(LonCentre, LatCentre, diet, allocation_type, ingredient, water_col_position, fmfo_embodied_fish, area_fmfo)
  
 

##### Now do trimmings

trim_tls <- read.csv(here("prep/02_feed/data/trimmings_fish_tls.csv"))

trimmings_fish <- readRDS(file.path(watson_dir, "v5.0/int/all_catch_trimmingsfish.rds"))

trim_props <- readRDS(here("prep/02_feed/data/embodied_FMFO_trimmings_rgns.rds"))

trim_npp <- 
 trimmings_fish %>% 
  left_join(trim_props, by = c("iso3c", "TaxonName")) %>%
  filter(!is.na(ingredient)) %>%
  mutate(prop = ifelse(is.na(prop), 0, prop)) %>% # just incase
  mutate(fmfo_embodied_fish = prop*trimmings_catch) %>% ## ADD IN TRIMMINGS WEIGHTS HERE ?
  left_join(fao_xyz) %>%
  left_join(ecosystem_te, by="fao_area_code") |> 
  left_join(trim_tls, by = "TaxonName") |>
  mutate(te = case_when(is.na(te) ~ ecosystem_te |> pull(te) |> mean(),
                        TRUE ~ te)) |> 
  left_join(mean_npp_xyz) |> 
    left_join(median_npp_xyz) |> 
    left_join(min_npp_xyz) |> 
    left_join(max_npp_xyz) |> 
    left_join(cell_area) |> 
    mutate(mean_npp_t_C_yr = mean_npp_t_km2_yr*area) |> 
    mutate(median_npp_t_C_yr = median_npp_t_km2_yr*area) |> 
    mutate(min_npp_t_C_yr = min_npp_t_km2_yr*area) |> 
    mutate(max_npp_t_C_yr = max_npp_t_km2_yr*area) |>
   mutate(ppr_fmfo = (fmfo_embodied_fish/carbon_conversion)*(1/te)^(DietTroph-1)) %>% #T C yr-1
    mutate(ppr_prop_fmfo = ppr_fmfo/mean_npp_t_C_yr) |> 
    mutate(area_fmfo = ppr_prop_fmfo*area)  %>%
  group_by(Cell, diet, allocation_type, ingredient, water_col_position) %>%
  summarise(area_fmfo = sum(area_fmfo, na.rm = TRUE),
            fmfo_embodied_fish = sum(fmfo_embodied_fish, na.rm = TRUE)) %>%
  ungroup() %>% 
  left_join(watson_cells) %>%
  dplyr::select(LonCentre, LatCentre, diet, allocation_type, ingredient, water_col_position, fmfo_embodied_fish, area_fmfo)
    


#Bind the forage and trimmings catch npp per cell

all_fish_npp <- 
  bind_rows(forage_npp |> select(diet, allocation_type, ingredient, water_col_position, LonCentre, LatCentre, area_fmfo, fmfo_embodied_fish),
            trim_npp |> select(diet, allocation_type, ingredient, water_col_position, LonCentre, LatCentre, area_fmfo, fmfo_embodied_fish)) |> 
  group_by(diet, allocation_type, ingredient, water_col_position, LonCentre, LatCentre) |> 
  summarise(area_fmfo = sum(area_fmfo, na.rm = TRUE),
            fmfo_embodied_fish = sum(fmfo_embodied_fish, na.rm = TRUE)) |> 
  ungroup()



## save as rasters 

diets <- unique(all_fish_npp$diet)
ingredients <- unique(all_fish_npp$ingredient)
allocations <- unique(all_fish_npp$allocation_type)
depth_positions <- unique(all_fish_npp$water_col_position)

for(diet_type in diets){
  for(ingredient_type in ingredients){
    for(allocation in allocations){
      for(pos in depth_positions){
      
      # diet_type = "plant-dominant"
      # ingredient_type = "fish meal, cut offs"
      # allocation = "mass"
      #  pos = "benthic"
      
      rast_xyz_df <- all_fish_npp %>%
        filter(diet == diet_type,
               ingredient == ingredient_type,
               allocation_type == allocation,
               water_col_position == pos) %>%
        dplyr::select(LonCentre, LatCentre, area_fmfo)
      
  if(str_detect(ingredient_type, "cut offs")){
        fish_type = "trimmings fish"
  }else{
        fish_type = "forage fish"
      }
        
  
      rast_xyz <- rast(rast_xyz_df, type = "xyz", crs = crs(base_rast)) %>%
        project(base_rast)
      
      writeRaster(rast_xyz, sprintf(here("prep/02_feed/output/%s/%s_%s_%s_%s_A.tif"), diet_type, fish_type, ingredient_type, allocation, pos), overwrite = TRUE)
      
      
      }
    }
  }
}

test <- rast(here("prep/02_feed/output/plant-dominant/archive/trimmings fish_fish meal, cut offs_mass_A.tif"))
plot(log(test+1))

global(test, "sum", na.rm = TRUE) # 2464.607 km2


test_ben <- rast(here("prep/02_feed/output/plant-dominant/trimmings fish_fish meal, cut offs_mass_benthic_A.tif"))
test_pel <- rast(here("prep/02_feed/output/plant-dominant/trimmings fish_fish meal, cut offs_mass_pelagic_A.tif"))

global(test_ben, "sum", na.rm = TRUE) + global(test_pel, "sum", na.rm = TRUE)  # 2464.607 km2

## PERFECT
```  

Data checking:

```{r}

test <- forage_npp %>%
  filter(diet == "plant-dominant",
         allocation_type == "mass",
         ingredient == "fish meal"
         )

test2 <- rast(test %>% dplyr::select(LonCentre, LatCentre, area_fmfo), type = "xyz", crs = crs(base_rast)) %>%
  project(base_rast)
plot(test2)
plot(log(test2+1))


sum(test$fmfo_embodied_fish) # 1156102 tonnage
sum(test$area_fmfo, na.rm = TRUE) # 6419.171 km2



test <- trim_npp %>%
  filter(diet == "plant-dominant",
         allocation_type == "economic",
         ingredient %in% c("fish meal, cut offs"))

test2 <- rast(test %>% dplyr::select(LonCentre, LatCentre, area_fmfo), type = "xyz", crs = crs(base_rast)) %>%
  project(base_rast)
 plot(test2)
plot(log(test2+1))

global(test2, "sum", na.rm = TRUE) # 211.5595 km2

sum(test$fmfo_embodied_fish) # 23076.47 tonnage
sum(test$area_fmfo, na.rm = TRUE) # 211.6933 km2

trim_catch <- rast(here("prep/02_feed/output/plant-dominant/archive/trimmings fish_fish meal, cut offs_mass_P.tif"))
plot(trim_catch)
plot(log(trim_catch + 1))

test <- all_fish_npp %>%
  filter(diet == "plant-dominant",
         allocation_type == "mass",
         str_detect(ingredient, "fish meal")
         ) %>%
  dplyr::select(LonCentre, LatCentre, area_fmfo, fmfo_embodied_fish) 


test2 <- rast(test %>% dplyr::select(LonCentre, LatCentre, area_fmfo), type = "xyz")
 plot(test2)
plot(log(test2+1))

sum(test$fmfo_embodied_fish) # 1371036 tonnage
sum(test$area_fmfo, na.rm = TRUE) # 8883.769

1371036/8883.769 # 154.3304


```

Check against Kuempel et al; will be different due to different methods

```{r}
kuempel_fofm_dist <- read.csv(here("prep/02_feed/test/sum_pressures_country.csv")) %>%
  filter(source == "feedfofm",
         animal_system == "salmon",
         type == "aquaculture",
         stressor == "disturbance"
         )

sum(kuempel_fofm_dist$value) # 80902.86... makes sense as they likely over estimate by a lot
```


