---
title: "Animal_consumption"
output: html_document
editor_options: 
  chunk_output_type: console
---

Get proportion of each MAPSPAM crop that is consumed by salmon aquaculture in each country.

Outputs: a raster for each crop describing the proportion of each crop consumed in each country by salmon aquaculture systems. This is a final output.

Be sure to run previous scripts if changes to data.

```{r}

library(here)
library(tidyverse)
library(countrycode)
library(tidytext)
library(knitr)

focus_year = 2021

source(here("src/fxns.R"))
source(here("src/directories.R"))

```


# STEP 1: determine amount of crop produced in each country for each food system.
Trade proportion data:
```{r}

trade_data <- read_csv(here("prep/02_feed/data/FAO_MAPSPAMcrop_trade_data.csv"))

```


Total consumption by animals per country in MAPSPAM crops
```{r}

consumption <- read_csv(here("prep/02_feed/data/demand/system_country_mapspam_tonnes_consumption.csv")) %>%
  rename(iso3c_consuming = iso3c, consuming_crop_consumed_tonnes =tonnes_product)

```


Join consumption by country and crop to trade data to get likely country of origin regardless of country of consumption.

```{r}
# check that all feed products match up: should be nothing showing up here
setdiff(consumption$SPAM_super, trade_data$SPAM_super)

production_location <- left_join(trade_data, consumption, by=c("iso3c_consuming", "SPAM_super")) %>%
  filter(!is.na(animal_system)) %>%  #crops not fed to animal system in given country
  mutate(consuming_crop_consumed_tonnes = ifelse(is.na(consuming_crop_consumed_tonnes), 0, consuming_crop_consumed_tonnes)) %>%
  mutate(producing_crop_consumed_tonnes = consuming_crop_consumed_tonnes * prop_of_supply)


```


Summarize by producing country to get the total production of feed for salmon aquaculture across all countries and diet scenarios


```{r}

feed_source <- production_location %>%
  group_by(iso3c_producing, animal_system, SPAM_super, product) %>%
  summarize(consumed_tonnes = sum(producing_crop_consumed_tonnes)) %>% 
  ungroup()



tmp <- filter(feed_source, animal_system == "salmon_aquaculture_meat") %>%
    group_by(product) %>%
  slice_max(order_by = consumed_tonnes, n = 15) %>%
  ungroup()


ggplot(data = tmp, aes(x = consumed_tonnes, y = reorder_within(iso3c_producing, consumed_tonnes, SPAM_super))) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~product, scales = "free") + 
  labs(x = "Tonnes of raw ingredient produced", y = "Country", title = "Plant diet") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5)) 

ggsave(here("prep/02_feed/plots/total_crop_feed_origins.png"), units = "in", height = 11, width = 11)
```

# STEP 2 Calculate proportion of production for animal system for each country

Crop production according to MAPSPAM data adjusted to 2021 production. These data include all growing systems (intensive, irrigated, etc.) and are in units of metric tonnes.

```{r}

crop_production_df <- read_csv(here("prep/02_feed/data/MAPSPAMcrop_production.csv"))
  
crop_production_global <- crop_production_df %>%
  group_by(SPAM_super) %>%
  summarize(tonnes_global = sum(tonnes_producing_crop, na.rm=TRUE))

```

Divide tonnes of consumption for each country/food system by total production to get the proportion of production produced for each food system:

```{r}

# if the data were perfect, we could leave it at this...but there are some corrections to make
# when consumption overshoots production
proportion <- left_join(feed_source, crop_production_df, by=c("iso3c_producing", "SPAM_super")) %>%
  mutate(tonnes_producing_crop = ifelse(is.na(tonnes_producing_crop), 0, tonnes_producing_crop)) %>%
  mutate(prop_produced_for_system = ifelse(consumed_tonnes == 0, 0, consumed_tonnes/tonnes_producing_crop))

```

# STEP 3: Distribute excess, and adjust proportion produced for system values

NOTE: if our consumption and production data were perfect these complicated steps would be unnecessary and the above calculation for prop_produced_for_system would be good enough. But due to mismatches, we spread excess consumption based on relative global production.

Identify the countries/crops with total consumption rates (across all systems) that exceed production

```{r}
## step 1 ID regions where consumption exceeds production

consumption_exceed <- proportion %>%
  group_by(iso3c_producing, SPAM_super, product) %>%
  mutate(tonnes_consumed_country = sum(consumed_tonnes)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(extra_crop_tonnes = tonnes_producing_crop - tonnes_consumed_country) %>%
  mutate(percent_exceed = ifelse(tonnes_producing_crop == 0, 0, tonnes_consumed_country/tonnes_producing_crop)) %>%
    data.frame()

```


Adjust proportion of consumption to not exceed production by more than 5%

```{r}

corrected_prop <- consumption_exceed %>%
  mutate(prop_of_consumption_within_iso = ifelse(tonnes_consumed_country == 0, 0, consumed_tonnes/tonnes_consumed_country)) %>%  # calculate proportion of crop each system eats within a country
  mutate(prop_produced_for_system_adjust = ifelse(percent_exceed >1 , 
                                        ifelse(percent_exceed > 1.05, prop_of_consumption_within_iso*1.05, prop_of_consumption_within_iso*percent_exceed),
                                        prop_produced_for_system))

summary(corrected_prop)
filter(corrected_prop, iso3c_producing=="BEL", SPAM_super=="soyb")

corrected_prop %>%
  group_by(iso3c_producing, SPAM_super, product) %>%
  summarize(total_adj_prop = sum(prop_produced_for_system_adjust),
            total_prop = sum(prop_produced_for_system))




```

For each animal system determine tonnes that was included and excluded.  The excluded will  be disbursed.

```{r}

tonnes_included_system <- corrected_prop %>%
  mutate(included_tonnes = prop_produced_for_system_adjust * tonnes_producing_crop) %>%
  mutate(lost_crop_tonnes = consumed_tonnes - included_tonnes) %>%
  mutate(included_prop = ifelse(tonnes_producing_crop == 0 , 0, included_tonnes/tonnes_producing_crop))
summary(tonnes_included_system)

# check, sum of included and excluded should equal sum of all consumption
tmp <- tonnes_included_system %>%
  group_by(SPAM_super) %>%
  summarize(included_crop = sum(included_tonnes, na.rm=TRUE),
            excluded_crop = sum(lost_crop_tonnes, na.rm=TRUE)) %>% 
  mutate(prop_redistributed = excluded_crop/(excluded_crop + included_crop)) %>%
           data.frame() 

sum(tmp$included_crop)/(sum(tmp$included_crop) + sum(tmp$excluded_crop)) # 97% included and does not need to be distributed

tonnes_lost_system <- tonnes_included_system %>%
  group_by(SPAM_super, animal_system, product) %>%
  summarize(lost_crop_tonnes = sum(lost_crop_tonnes, na.rm=TRUE))


#calculate extra crop tonnes in each country, and determine what proportion of extra each country accounts for
tonnes_extra_crop <- corrected_prop %>%
  dplyr::select(iso3c_producing, SPAM_super, extra_crop_tonnes, tonnes_producing_crop, product) %>%
  unique() %>%
  mutate(extra_crop_tonnes = ifelse(extra_crop_tonnes<0, 0, extra_crop_tonnes)) %>%
  group_by(SPAM_super) %>%
  mutate(excess_crop_global = sum(extra_crop_tonnes)) %>%
  rowwise() %>%
  mutate(prop_excess_crop = ifelse(excess_crop_global == 0, 0, extra_crop_tonnes/excess_crop_global))

# distribute the tonnes lost for each crop/system based on each countries extra production
excess_prop <- merge(tonnes_extra_crop, tonnes_lost_system) %>%
  mutate(tonnes_dispersed = prop_excess_crop * lost_crop_tonnes) %>%
  mutate(prop_dispersed = ifelse(tonnes_producing_crop == 0, 0, tonnes_dispersed/tonnes_producing_crop))


### NOTE: you can get higher total props for countries with very small proportions. This happens only for soybeans and for small production areas.
## this explores this:

excess_excess_prop <- excess_prop %>%
  mutate(extra_dispersed_tonnes = tonnes_dispersed - extra_crop_tonnes) %>%
  mutate(extra_dispersed_tonnes = ifelse(prop_dispersed > 1, extra_dispersed_tonnes, 0)) %>%
  left_join(crop_production_global, by = "SPAM_super") %>%
  mutate(prop_global_crop_production = tonnes_producing_crop/tonnes_global) %>%
  group_by(SPAM_super, product, animal_system) %>%
  mutate(extra_dispersed_tonnes_by_system_crop = sum(extra_dispersed_tonnes))
  

filter(excess_excess_prop, prop_dispersed>1) %>% data.frame()

```


Combine the included and lost proportions for each system and country for plotting below

```{r}

assigned_prop <- tonnes_included_system  %>%
  dplyr::select(iso3c_producing, product, animal_system, SPAM_super, prop_produced_for_system = included_prop)
excess_prop <- excess_prop %>%
  dplyr::select(iso3c_producing, product, animal_system, SPAM_super, prop_extra_for_system = prop_dispersed)

total_prop <- left_join(assigned_prop, excess_prop, by=c("iso3c_producing", "animal_system", "SPAM_super", "product")) %>%
  rowwise() %>%
  mutate(total_prop = prop_produced_for_system + prop_extra_for_system)
summary(total_prop)
filter(total_prop, iso3c_producing=="AFG" & SPAM_super=="soyb") %>% data.frame()

```

Checking
```{r}

summary(total_prop) # ok for total prop to be a bit above 1

filter(total_prop, prop_produced_for_system>1)

## check: should match original consumption:
consumption_global <- feed_source %>%
  group_by(animal_system, product, SPAM_super) %>%
  summarize(obs_tonnes = sum(consumed_tonnes, na.rm=TRUE))

check <- left_join(total_prop, crop_production_df) %>%
  mutate(total_tonnes = total_prop*tonnes_producing_crop) %>%
  group_by(animal_system, product, SPAM_super) %>%
  summarize(est_tonnes = sum(total_tonnes, na.rm=TRUE)) %>%
  left_join(consumption_global) %>%
  rowwise() %>%
  mutate(difference = obs_tonnes - est_tonnes )

# these should be the same
check %>%
  group_by(SPAM_super) %>%
  summarize(est_tonnes =sum(est_tonnes),
            obs_tonnes = sum(obs_tonnes)) %>% data.frame()

plot(log(check$est_tonnes+1), log(check$obs_tonnes+1))
abline(0,1, col="red") # these should look about the same!

```


```{r}

write_csv(total_prop, here("prep/02_feed/output/plant_diet/proportion_feed_per_country_system.csv"))

```



For each crop determine the proportion of each countries production dedicated to the animal system.

```{r}

library(raster)

## organize region data
rgns <- read_csv(file.path(food_systems_data_dir, "food_rgns_xy.csv"))

####################
## start loop here
#################3

todo <- unique(dplyr::select(total_prop, SPAM_super, product, animal_system))


for(i in 1:dim(todo)[1]){
#i=4
crop <- todo$SPAM_super[i]
system <- todo$animal_system[i]
product <- todo$product[i]
diet = "plant_diet"


crop_spam <- str_replace(crop, "x", "o")


# total country feed production
feed_crop <- filter(total_prop, SPAM_super==crop, animal_system==system) %>%
  dplyr::select(iso3c = iso3c_producing, prop_feed = total_prop)

# proportion crop grown for feed per country
prop_feed_crop <- left_join(rgns, feed_crop, by="iso3c") %>%
  dplyr::select(x, y, prop_feed)

prop_feed_crop_raster <- rasterFromXYZ(prop_feed_crop)

# plot(prop_feed_crop_raster)

crs(prop_feed_crop_raster) <- "+proj=longlat +datum=WGS84"


tonnes_crop_raster <- raster(sprintf(file.path(mapspam_dir, "scaled_maps_2021/crop_%s_A_scaled.tif"), crop_spam))

tonnes_crop_feed_raster <- tonnes_crop_raster*prop_feed_crop_raster

writeRaster(tonnes_crop_feed_raster, filename= sprintf(here("prep/02_feed/output/%s/%s_%s_%s.tif"), diet, system, crop, product), overwrite=TRUE)

}


```
 

