---
title: "Animal_consumption"
output: html_document
editor_options: 
  chunk_output_type: console
---

Get proportion of each MAPSPAM crop that is consumed by salmon aquaculture in each country.

Outputs: a raster for each crop describing the proportion of each crop consumed in each country by salmon aquaculture systems. This is a final output.

Be sure to run previous scripts if changes to data.

```{r}

library(here)
library(tidyverse)
library(countrycode)
library(tidytext)
library(knitr)
library(strex)
library(sjmisc)

focus_year = 2021

source(here("src/fxns.R"))
source(here("src/directories.R"))

```


# STEP 1: determine amount of crop produced in each country for each food system.
Trade proportion data:
```{r}

trade_data <- read_csv(here("prep/02_feed/data/FAO_MAPSPAMcrop_trade_data.csv")) 

```


Total consumption by animals per country in MAPSPAM crops
```{r}

consumption <- read_csv(here("prep/02_feed/data/demand/total_aquaculture_ingredient_demand_corrected.csv")) %>%
  mutate(animal_system = paste(animal, system, product, sep="_")) %>%
  rename(source_ingredient = ingredient) %>%
  select(iso3c_consuming = iso3c, animal_system, diet, source_ingredient, consuming_ingredient_consumed_tonnes = ingredient_demand) 

## need to match `consumption` ingredients to equivalent crop.. need lookup table 

feed_cats <- read_csv(here("prep/02_feed/data/ingredient_spam_fao_codes.csv")) %>% ## this table was created manually. It matches our ingredient names from Aas et al. 2022 to mapspam names. Must also update product_to_fao_spam.csv. 
  select(-notes)  %>%
  filter(included_source == "mapspam")


consumption_spam <- consumption %>%
  left_join(feed_cats) %>%
  filter(included_source == "mapspam") 

```


Join consumption by country and crop to trade data to get likely country of origin regardless of country of consumption.

```{r}
# check that all feed products match up: should be nothing showing up here
setdiff(consumption_spam$SPAM_super, trade_data$SPAM_super)
# [1] "opul" "ooil" "rape" "sunf"... this should be empty. I need to go back and fix this in 1.1_MAPSPAM_production.Rmd

production_location <- left_join(trade_data, consumption_spam, by=c("iso3c_consuming", "SPAM_super")) %>%
  filter(!is.na(animal_system)) %>%  #crops not fed to animal system in given country
  mutate(consuming_ingredient_consumed_tonnes = ifelse(is.na(consuming_ingredient_consumed_tonnes), 0, consuming_ingredient_consumed_tonnes)) %>%
  mutate(producing_ingredient_consumed_tonnes = consuming_ingredient_consumed_tonnes * prop_of_supply)


```


Take a look at the data 

```{r}

feed_source <- production_location %>%
  group_by(iso3c_producing, diet, animal_system, SPAM_super, source_ingredient) %>%
  summarize(consumed_ingredient_tonnes = sum(producing_ingredient_consumed_tonnes)) %>% 
  ungroup()



tmp <- filter(feed_source, animal_system == "salmon_aquaculture_meat") %>%
    group_by(source_ingredient, diet) %>%
  slice_max(order_by = consumed_ingredient_tonnes, n = 15) %>%
  ungroup() %>%
  filter(consumed_ingredient_tonnes > 0)


ggplot(data = tmp %>%
         filter(diet == "plant-dominant"), aes(x = consumed_ingredient_tonnes, y = reorder_within(iso3c_producing, consumed_ingredient_tonnes, SPAM_super))) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~source_ingredient, scales = "free") + 
  labs(x = "Tonnes of raw ingredient produced", y = "Country") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5)) 

# ggsave(here("prep/02_feed/plots/total_crop_feed_origins.png"), units = "in", height = 11, width = 11)
```


Ok now we need to convert the ingredient to crop raw material using the conversion factors that Rich has put together 


```{r}
## read in raw material extraction rates (yields) that Rich compiled 

#get yield data from main ingredients
cf_data <- read_csv(file.path("/mnt/rdsi/raw_data/allocation/top_crop_producers_conversions.csv")) |> 
  mutate(ingredient = if_else(ingredient == "canola/camelina oil", true = "canola oil", false = ingredient)) %>% 
  dplyr::select(1:6)
# get coproduct yield data

cf_coproduct_data <- read_csv(file.path("/mnt/rdsi/raw_data/allocation/coproduct_conversions.csv"))

unique(cf_data$ingredient)

#  [1] "faba beans"              "tapioca starch"          "linseed oil"             "lupins"                  "corn starch"            
#  [6] "corn gluten meal"        "pea starch"              "pea protein concentrate" "guar meal"               "canola oil"             
# [11] "soy oil"                 "soy protein concentrate" "soybean meal"            "sunflower meal"          "wheat"                  
# [16] "wheat gluten" 

unique(production_location$source_ingredient)

setdiff(unique(cf_data$ingredient), unique(production_location$source_ingredient))

# [1] "tapioca starch"          "lupins"                  "corn starch"             "corn gluten meal"        "pea starch"             
#  [6] "pea protein concentrate" "guar meal"               "canola oil"              "soy oil"                 "sunflower meal"  

setdiff(unique(production_location$source_ingredient), unique(cf_data$ingredient))

# [1] "guar protein"              "pea flour"                 "pea protein"               "corn gluten"               "soybean oil"              
# [6] "coconut oil"               "sunflower"                 "rapeseed and camelina oil"


## ok so... in our data I need to change: 
# rapeseed and camelina oil ---> canola oil
# soybean oil ---> soy oil
# pea flour ---> pea starch 
# guar protein ---> guar meal ? 
# corn gluten ---> corn gluten meal 
# sunflower ---> sunflower meal 
# pea protein ---> pea protein concentrate 
# missing coconut oil in Rich's data 

```









# STEP 2 Calculate proportion of production for animal system for each country

Crop production according to MAPSPAM data adjusted to 2021 production. These data include all growing systems (intensive, irrigated, etc.) and are in units of metric tonnes.

```{r}

crop_production_df <- read_csv(here("prep/02_feed/data/MAPSPAMcrop_production.csv"))
  
crop_production_global <- crop_production_df %>%
  group_by(SPAM_super) %>%
  summarize(tonnes_global = sum(tonnes_producing_crop, na.rm=TRUE))

```

Divide tonnes of consumption for each country/food system by total production to get the proportion of production produced for each food system:

```{r}

# if the data were perfect, we could leave it at this...but there are some corrections to make
# when consumption overshoots production
proportion <- left_join(feed_source, crop_production_df, by=c("iso3c_producing", "SPAM_super")) %>%
  mutate(tonnes_producing_crop = ifelse(is.na(tonnes_producing_crop), 0, tonnes_producing_crop)) %>%
  mutate(prop_produced_for_system = ifelse(consumed_tonnes == 0, 0, consumed_tonnes/tonnes_producing_crop))

```

# STEP 3: Distribute excess, and adjust proportion produced for system values

NOTE: if our consumption and production data were perfect these complicated steps would be unnecessary and the above calculation for prop_produced_for_system would be good enough. But due to mismatches, we spread excess consumption based on relative global production.

Identify the countries/crops with total consumption rates (across all systems) that exceed production

```{r}
## step 1 ID regions where consumption exceeds production

consumption_exceed <- proportion %>%
  group_by(iso3c_producing, diet, SPAM_super, product) %>%
  mutate(tonnes_consumed_country = sum(consumed_tonnes)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(extra_crop_tonnes = tonnes_producing_crop - tonnes_consumed_country) %>%
  mutate(percent_exceed = ifelse(tonnes_producing_crop == 0, 0, tonnes_consumed_country/tonnes_producing_crop)) %>%
    data.frame()

```


Adjust proportion of consumption to not exceed production by more than 5%

```{r}

corrected_prop <- consumption_exceed %>%
  mutate(prop_of_consumption_within_iso = ifelse(tonnes_consumed_country == 0, 0, consumed_tonnes/tonnes_consumed_country)) %>%  # calculate proportion of crop each system eats within a country
  mutate(prop_produced_for_system_adjust = ifelse(percent_exceed >1 , 
                                        ifelse(percent_exceed > 1.05, prop_of_consumption_within_iso*1.05, prop_of_consumption_within_iso*percent_exceed),
                                        prop_produced_for_system))

summary(corrected_prop)
filter(corrected_prop, iso3c_producing=="BEL", SPAM_super=="soyb")

corrected_prop %>%
  group_by(iso3c_producing, diet, SPAM_super, product) %>%
  summarize(total_adj_prop = sum(prop_produced_for_system_adjust),
            total_prop = sum(prop_produced_for_system))




```

For each animal system determine tonnes that was included and excluded.  The excluded will  be disbursed.

```{r}

tonnes_included_system <- corrected_prop %>%
  mutate(included_tonnes = prop_produced_for_system_adjust * tonnes_producing_crop) %>%
  mutate(lost_crop_tonnes = consumed_tonnes - included_tonnes) %>%
  mutate(included_prop = ifelse(tonnes_producing_crop == 0 , 0, included_tonnes/tonnes_producing_crop))
summary(tonnes_included_system)

# check, sum of included and excluded should equal sum of all consumption
tmp <- tonnes_included_system %>%
  group_by(SPAM_super, diet) %>%
  summarize(included_crop = sum(included_tonnes, na.rm=TRUE),
            excluded_crop = sum(lost_crop_tonnes, na.rm=TRUE)) %>% 
  mutate(prop_redistributed = excluded_crop/(excluded_crop + included_crop)) %>%
           data.frame() 

sum(tmp$included_crop)/(sum(tmp$included_crop) + sum(tmp$excluded_crop)) # 98% included and does not need to be distributed

tonnes_lost_system <- tonnes_included_system %>%
  group_by(SPAM_super, diet, animal_system, product) %>%
  summarize(lost_crop_tonnes = sum(lost_crop_tonnes, na.rm=TRUE))


#calculate extra crop tonnes in each country, and determine what proportion of extra each country accounts for
tonnes_extra_crop <- corrected_prop %>%
  dplyr::select(iso3c_producing, diet, SPAM_super, extra_crop_tonnes, tonnes_producing_crop, product) %>%
  unique() %>%
  mutate(extra_crop_tonnes = ifelse(extra_crop_tonnes<0, 0, extra_crop_tonnes)) %>%
  group_by(SPAM_super) %>%
  mutate(excess_crop_global = sum(extra_crop_tonnes)) %>%
  rowwise() %>%
  mutate(prop_excess_crop = ifelse(excess_crop_global == 0, 0, extra_crop_tonnes/excess_crop_global))

# distribute the tonnes lost for each crop/system based on each countries extra production
excess_prop <- merge(tonnes_extra_crop, tonnes_lost_system) %>%
  mutate(tonnes_dispersed = prop_excess_crop * lost_crop_tonnes) %>%
  mutate(prop_dispersed = ifelse(tonnes_producing_crop == 0, 0, tonnes_dispersed/tonnes_producing_crop))


### NOTE: you can get higher total props for countries with very small proportions. This happens only for soybeans and for small production areas.
## this explores this:

excess_excess_prop <- excess_prop %>%
  mutate(extra_dispersed_tonnes = tonnes_dispersed - extra_crop_tonnes) %>%
  mutate(extra_dispersed_tonnes = ifelse(prop_dispersed > 1, extra_dispersed_tonnes, 0)) %>%
  left_join(crop_production_global, by = "SPAM_super") %>%
  mutate(prop_global_crop_production = tonnes_producing_crop/tonnes_global) %>%
  group_by(SPAM_super, diet, product, animal_system) %>%
  mutate(extra_dispersed_tonnes_by_system_crop = sum(extra_dispersed_tonnes))
  

filter(excess_excess_prop, prop_dispersed>1) %>% data.frame()

```


Combine the included and lost proportions for each system and country for plotting below

```{r}

assigned_prop <- tonnes_included_system  %>%
  dplyr::select(iso3c_producing, diet, product, animal_system, SPAM_super, prop_produced_for_system = included_prop)
excess_prop <- excess_prop %>%
  dplyr::select(iso3c_producing, diet, product, animal_system, SPAM_super, prop_extra_for_system = prop_dispersed)

total_prop <- left_join(assigned_prop, excess_prop, by=c("iso3c_producing", "diet", "animal_system", "SPAM_super", "product")) %>%
  rowwise() %>%
  mutate(total_prop = prop_produced_for_system + prop_extra_for_system)
summary(total_prop)
filter(total_prop, iso3c_producing=="AFG" & SPAM_super=="soyb") %>% data.frame()

```

Checking
```{r}

summary(total_prop) # ok for total prop to be a bit above 1

filter(total_prop, prop_produced_for_system>1)

## check: should match original consumption:
consumption_global <- feed_source %>%
  group_by(animal_system, diet, product, SPAM_super) %>%
  summarize(obs_tonnes = sum(consumed_tonnes, na.rm=TRUE))

check <- left_join(total_prop, crop_production_df) %>%
  mutate(total_tonnes = total_prop*tonnes_producing_crop) %>%
  group_by(animal_system, diet, product, SPAM_super) %>%
  summarize(est_tonnes = sum(total_tonnes, na.rm=TRUE)) %>%
  left_join(consumption_global) %>%
  rowwise() %>%
  mutate(difference = obs_tonnes - est_tonnes )

# these should be the same
check %>%
  group_by(SPAM_super) %>%
  summarize(est_tonnes =sum(est_tonnes),
            obs_tonnes = sum(obs_tonnes)) %>% data.frame()

plot(log(check$est_tonnes+1), log(check$obs_tonnes+1))
abline(0,1, col="red") # these should look about the same!


plot(check$est_tonnes, check$obs_tonnes+1)
abline(0,1, col="red") # these should look about the same!

```


```{r}

write_csv(total_prop, here("prep/02_feed/output/proportion_feed_per_country_system.csv"))

```


