---
title: "GAEZ_scaling_2015_to_2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Script Description

Rescale GAEZ data to year 2021.

# Data information

We do this using FAOSTAT crop production data trends between year 2015 and 2021. Using 3 year mean. I.e. 2014, 2015, 2016 for early years (this is what GAEZ actually represents when they say "2015"); 2019, 2020, 2021 for later years (to deal with shocks to system, like 2020 covid shock and 2021 recovery).

# Setup
```{r setup, include = FALSE}
# Raster templates
library(here)
library(janitor)
library(terra)
library(vroom)

source(here("src/spatial.R"))
source(here("src/fxns.R"))
source(here("src/directories.R"))


food_rgns <- read.csv(here("data/spatial/output/food_rgns.csv"))

```

# Methods 

## Import and wrangle FAO data for 2015 and 2021

```{r}
crops_gaez <- read.csv(here("prep/01_crops_prep/data/GAEZ_names.csv")) %>%
  dplyr::select(GAEZ_category, item_code)

scaling_coef_tidy_gaez <- 
  read.csv(here("prep/01_crops_prep/data/FAOSTAT_crop_production_2015_2021.csv")) %>% 
  clean_names() %>% 
  dplyr::select(area_code, item_code, year, value) %>% 
  rename(tonnes = value) %>% 
  na.omit() %>%
  filter(item_code %in% crops_gaez$item_code) %>%
  mutate(era = ifelse(year %in% c(2014, 2015, 2016), "early", "late"))

test <- scaling_coef_tidy_gaez %>%
  filter(area_code %in% c(276, 277)) # just check sudan and south sudan are split...

```

# Wrangle FAO dataset 

```{r}
scaling_coef_tidy_gaez <- 
  scaling_coef_tidy_gaez %>% 
  filter(!area_code %in% 351) %>% # Remove China
  left_join(read_csv(here("data/spatial/output/FAO_rgn_codes.csv")), 
            by = "area_code") %>% # Add iso3c
  left_join(crops_gaez) %>% # Add SPAM_short_name
  group_by(year, iso3c, GAEZ_category, era) %>%
  summarise(tonnes = sum(tonnes, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::select(year, tonnes, iso3c, GAEZ_category, era) %>% 
  na.omit() # Remove NA created by milled rice.

# Distinct area_codes
n_distinct(scaling_coef_tidy_gaez$iso3c) # 200

missing_rgns <- setdiff(food_rgns$iso3c, scaling_coef_tidy_gaez$iso3c)
missing_rgns # Nothing unusual, all small islands.
```

# Calculate scaling from 2010 to 2021 for all crops and countries using 3 year means.
```{r}
scaling_coef <- 
  scaling_coef_tidy_gaez %>% 
  group_by(GAEZ_category, era, iso3c) %>%
  summarize(tonnes = mean(tonnes, na.rm = TRUE)) %>% # Calculate mean tonnes for each crop, year and country.
  ungroup() %>% 
  spread(., era, tonnes) %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  group_by(GAEZ_category, iso3c, `late`, `early`) %>%
  summarize(coef = `late` / `early`, .groups = "keep") %>% 
  ungroup()

# Make sure that countries with more than 0 production in 2015 but
# no data in 2021 have 0 value for 2021.
scaling_coef$coef[scaling_coef$`early` > 0 & scaling_coef$`late` == 0] <- 0

# Some countries will have data in 2017, according to the FAO, but none in 2010. We need to make sure that we manually insert these values in the scaled maps.

inf_values <- 
  scaling_coef %>% 
  filter(`early` == 0 & `late` > 0) %>% 
  dplyr::select(GAEZ_category, iso3c, tonnes = `late`)
inf_values

scaling_coef$coef[scaling_coef$`early` == 0 & scaling_coef$`late` > 0] <- 1

# All infinite numbers assigned a coef of 1.
scaling_coef$coef[which(!is.finite(scaling_coef$coef))] <- 1


try <- scaling_coef %>% filter(GAEZ_category == "Pulses")
sum(try$early) #78030351 - ok that's perfect. 
```

# Save
```{r}
vroom_write(inf_values, here("prep/01_crops_prep/data/inf_values.csv"))
write_csv(scaling_coef, here("prep/01_crops_prep/data/scaling_coef.csv"))

```

-------------------------------------------------------------------------------
