---
title: "Make maps"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Setup

Load libraries

```{r setup, include=FALSE}

library(tidyverse)
library(here)
library(data.table)
library(terra)
library(raster)
library(rnaturalearth)
library(strex)
library(janitor)
library(sf)
library(tools)
library(khroma)
library(scales)
library(fields)
library(patchwork)
library(cowplot)
library(glue)
library(viridis)
library(RColorBrewer)
library(ggpubr)
library(smoothr)

source(here("src/directories.R"))

source(here("src/spatial.R"))

select <- dplyr::select
setwd(dirname(rstudioapi::getSourceEditorContext()$path)) # set working directory to where this script is located
this_dir <- getwd()

biodiv_dir <- file.path(rdsi_dir, "biodiversity_impacts")

impact_maps_dir <- file.path(biodiv_dir, "output")

globe_bbox <- rbind(c(-180, -90), c(-180, 90), 
                      c(180, 90), c(180, -90), c(-180, -90)) 

globe_border <- st_polygon(list(globe_bbox)) %>%
  st_sfc(crs = 4326) %>%
  st_sf(data.frame(rgn = 'globe', geom = .)) %>%
  smoothr::densify(max_distance = 0.5) %>%
  st_transform(crs = crs(moll_template))


```

```{r establish color palettes}

## establish color palette from food footprint project
d_brown <- "#515256"
m_brown <- "#B27B54"
l_brown <- "#BC995F"
green <- "#A8B072"
yellow<- "#EFCE71"
l_yellow <- "#F7F6C1"

jv_pal <- c(l_yellow, yellow, green, l_brown, m_brown, d_brown)

continuous_pal <-  colorRampPalette(jv_pal, space="Lab", bias = 3.5)(10000)
image(volcano, asp=1, col=continuous_pal)

light_gray <- "#F8F9FA"
final_palette <- c(light_gray, continuous_pal)
image(volcano, asp=1, col=final_palette)

palette_red <- c(final_palette, "#B90000")
image(volcano, asp=1, col=palette_red)

light_gray <- "#F8F9FA"
final_palette <- c(light_gray, continuous_pal)
scales::show_col(final_palette)

countries_shp <- ne_countries(scale = 110, returnclass = "sf") |> 
  st_transform(crs(moll_template)) # read in countries shapefile

theme_ohara <- function(base_size = 9) {
  theme(axis.ticks = element_blank(),
        text             = element_text(family = 'Helvetica',
                                        color = 'gray30', size = base_size),
        plot.title       = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
        panel.background = element_blank(),
        legend.position  = 'right',
        panel.border     = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(colour = 'grey90', size = .25),
        legend.key       = element_rect(colour = NA, fill = NA),
        axis.line        = element_blank())
}


```

Read in marine and terrestrial dfs and combine

```{r}
marine_overlap <- readRDS(here("prep/03_prep_spp_habitats/output/aoh_marine_impact.rds")) %>%
  group_by(sciname, spp_type, diet, fcr_type, allocation, ingredient) %>%
  summarise(vulnerability = mean(vulnerability, na.rm = TRUE), 
            impact_total = sum(impact_total, na.rm = TRUE)) %>%
  ungroup() %>%
    dplyr::select(sciname, spp_type, diet, fcr_type, allocation, ingredient, vulnerability, impact_total)


terrestrial_overlap <- readRDS(here("prep/03_prep_spp_habitats/output/aoh_terrestrial_mammals_birds_impact.rds")) %>% 
  group_by(sciname, diet, fcr_type, allocation, ingredient, spp_type) %>%
  summarise(vulnerability = 1 - mean(suitability, na.rm = TRUE),
            impact_total = sum(impact_total, na.rm = TRUE)) %>%
  ungroup() %>%
    dplyr::select(sciname, spp_type, diet, fcr_type, allocation, ingredient, vulnerability, impact_total)

all_overlap <- rbind(marine_overlap, terrestrial_overlap)

spp_impacted <- all_overlap %>%
  group_by(spp_type, diet, allocation, fcr_type) %>%
  summarise(spp_count_impacted = n_distinct(sciname)) %>%
  ungroup()


aquamaps_spp <- readRDS(file.path(rdsi_dir, "biodiversity_impacts/int/spp_vuln_depth_info.rds")) %>%
  distinct(spp_type = taxon, sciname = species)
  
eyres_spp <- readRDS(here("prep/03_prep_spp_habitats/data/iucn/eyres_iucn_spp.rds")) %>%
  dplyr::select(spp_type, sciname = scientific_name) %>%
  mutate(sciname = tolower(sciname)) %>%
  mutate(spp_type = case_when(
    spp_type == "mammals" ~ "Terrestrial mammal",
    spp_type == "amphibians" ~ "Amphibians", 
    spp_type == "reptiles" ~ "Reptiles", 
    spp_type == "birds" ~ "Bird"
  ))

spp_assessed <- rbind(aquamaps_spp, eyres_spp) %>%
  group_by(spp_type) %>%
  summarise(spp_count_assessed = n_distinct(sciname)) %>%
  ungroup()
```

Maps!

Global maps across ingredients and taxon; Delta plots as well

```{r}


## only need to save RDS when impact maps info has changed
allocations <- c("ge", "mass")

for(allocation_type in allocations){

 # allocation_type = "economic"

global_maps_list <- list.files(file.path(impact_maps_dir, "impact_maps_across_taxon_ingredient"), pattern = allocation_type, recursive = TRUE, full.names = TRUE)

global_maps_mean <- rast(global_maps_list)
names(global_maps_mean) <- file_path_sans_ext(str_after_nth(global_maps_list, "\\/", 6))


all_df <- as.data.frame(global_maps_mean, xy = TRUE) %>%
  pivot_longer(cols = c(3:ncol(.)), values_to = "value", names_to = "type") %>%
  separate_wider_delim(type, delim = "/", names = c("diet", "fcr_type", "calc_type")) %>%
  mutate(value_log =  ifelse(str_detect(calc_type, "mean"), log(value*1000000 + 1), value)) %>%
  rename(long = x, lat = y)

for(i in unique(all_df$calc_type)){

  # i = "economic_mean"

  loop_df <- all_df %>%
    filter(calc_type == i) 

data.table::fwrite(loop_df, glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/{i}.rds.gz")))

}
}



## make plots showing impacts for each scenario 

allocation_types <- c("mass", "economic", "ge")  
calc_types <- c("mean", "nspp", "sd")
types_df <- expand.grid(allocation_types, calc_types) %>%
  unite(new_var, Var1:Var2, sep = "_") %>%
  pull(new_var) %>%
  unique()

diets <- c("fish-dominant", "plant-dominant")
fcrs <- c("regular", "efficienct")

for(type in types_df){
  
  # type = "economic_mean"
  
 filepath <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/{type}.rds.gz"))

loop_df <-  fread(filepath) 
  
## create maps in separate plots, force common scale between them
maps <- map(.x = diets, 
            .f = function(x, y) 
  ggplot() + 
  geom_tile(data = loop_df  %>% 
              filter(diet == x), aes(x= long, y = lat, fill = value_log)) +  # change to regular value if you don't want log transformed...
      # scale_fill_gradientn(colors = YlOrBr(9), na.value = "white") +  # Set the color for missing values (NA) to white
   scale_fill_gradientn(colors = final_palette, na.value = "white") +
   #  scale_fill_viridis(direction = -1) + 
              geom_sf(data = countries_shp, fill = NA, colour = "grey") + 
    facet_wrap(~fcr_type) + 
  theme_minimal() +
      theme(
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank())
            )

ggsave(glue(here(this_dir, "output/plots/{type}_plot.png")), plot = plot_grid(plotlist = maps, labels = glue("{diets}:{type}"), label_size = 10, nrow = 2), width = 10, height = 8, bg = "white")


}

test <- loop_df %>% filter(calc_type == "economic_mean")


# create delta (change) plots between each diet scenario

for(type in types_df){
  
  # type = "economic_mean"
 
  # if(file.exists(glue(here(this_dir, "output/plots/delta/{type}_plot.png")))){
  #   cat("exists")
  #   next()
  # }

 filepath <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/{type}.rds.gz"))

delta_df <-  fread(filepath) %>%
        dplyr::select(-value_log)

  delta_df <- delta_df %>%
      pivot_wider(names_from = c(diet), values_from = value) %>%
    mutate(`fish-dominant` = ifelse(is.na(`fish-dominant`), 0, `fish-dominant`),
           `plant-dominant` = ifelse(is.na(`plant-dominant`), 0, `plant-dominant`)) %>%
  mutate(delta = (`fish-dominant` - `plant-dominant`)) %>%
    mutate(delta_log = log(delta + 1)) %>%
    mutate(delta_rescale = delta*1000000) %>%
    mutate(delta_cat = case_when(
      delta > 0 ~ "Fish-dominant",
      delta < 0 ~ "Plant-dominant", 
      delta == 0 ~ "No difference"
    ))

  
data.table::fwrite(delta_df, glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/delta/{type}.rds.gz")))


  
if(str_detect(type, "nspp")){
  limit <- c(min(delta_df$delta), max(delta_df$delta))

}else{
limit <- c(quantile(delta_df$delta, 0.05), quantile(delta_df$delta, 0.9))

}

## create maps in separate plots, force common scale between them
p <- ggplot() +
  geom_tile(data = delta_df %>% filter(delta != 0) , aes(x= long, y = lat, fill = delta)) +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, limit = limit, na.value = "white", oob=scales::squish) + # this will make sure anything above the scales limit will be a solid color
  geom_sf(data = countries_shp, fill = NA, colour = "grey") +
    facet_wrap(~fcr_type) +
  theme_minimal() +
      theme(
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank()) +
    labs(caption = str_replace(glue("{type}"), "_", "; "))

p <- ggplot() +
  geom_raster(data = delta_df , aes(x= long, y = lat, fill = delta)) +
  geom_sf(data = countries_shp, fill = NA, colour = "grey") +
        geom_sf(data = globe_border,
            fill = NA, color = 'grey70', 
            size = .1) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_void() +
    facet_wrap(~fcr_type) +
    labs(caption = str_replace(glue("{type}"), "_", "; "), 
         fill = 'Δ Extinction Risk') + 
        scale_fill_gradient2(high = '#d01c8b', mid = '#ffffdf', low = '#4dac26', limit = limit, oob=scales::squish) 
      # scale_fill_manual(values = c("#d01c8b", "#ffffdf", "#4dac26")) 
       


  ggsave(plot = p, glue(here(this_dir, "output/plots/delta/{type}_plot_test.png")), width = 6.5, height = 3, dpi = 300)

}

test <- rast("/mnt/rdsi/biodiversity_impacts/output/impact_maps_across_taxon_ingredient/plant-dominant/regular/economic_mean.tif")
test2 <- rast("/mnt/rdsi/biodiversity_impacts/output/impact_maps_across_taxon_ingredient/fish-dominant/regular/economic_mean.tif")

test3 <- test2 - test

test3[test3 == 0] <- NA

plot(test3*1000000)
# seems to match


## zoom in on south america 

delta_df <- data.table::fread((file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/delta/economic_mean.rds.gz")))


limit <- c(quantile(delta_df$delta, 0.05), quantile(delta_df$delta, 0.90))

test <- delta_df %>% filter(delta > 0) %>% filter(fcr_type == "regular") %>% dplyr::select(long, lat, delta) %>%
  rast(., type = "xyz", crs = moll_template)
plot(test, col = "red")
  

test_rasterize <- countries_shp %>% filter(continent == "South America") %>%
  rasterize(., moll_template) %>%
  as.data.frame(., xy = TRUE)

## create maps in separate plots, force common scale between them
  ggplot() +
  geom_tile(data = delta_df %>% filter(long %in% c(unique(test_rasterize$x)),
                                       lat %in% c(unique(test_rasterize$y))) , aes(x= long, y = lat, fill = delta)) +
   scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, limit = limit, oob = scales::squish) +
    facet_wrap(~fcr_type) +
                  geom_sf(data = countries_shp %>% filter(continent == "South America"), fill = NA, colour = "grey") +
  theme_minimal() +
      theme(
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank()) +
    labs(caption = "economic_mean")
  
    ggsave(here(this_dir, "output/plots/delta/south_america_compare.png"), width = 10, height = 8, bg = "white")

```

Global maps across ingredients per taxon - there is an if statement within each loop that will skip if the files exists. If you want to rerun all of the files you will need to comment that out.

```{r}

allocations <- c("economic", "ge", "mass")
calc_types <- c("mean", "sd", "nspp")
taxon <- all_overlap %>% 
  filter(spp_type != "polychaetes") %>% 
  pull(spp_type) %>% unique()

## this part only needs to be run if the source maps changed! 
for(allocation_type in allocations){
    for(taxa_type in taxon){
      for(type in calc_types){

# allocation_type = "economic"
#  taxa_type = "corals"

      out_fstem <- glue(file.path(impact_maps_dir, "csvs/impact_maps_by_taxon_across_ingredient/{allocation_type}_{taxa_type}_{type}.rds.gz"))

      # if(file.exists(out_fstem)){
      # 
      #   message(glue("skipping {allocation_type}_{taxa_type}_{type}"))
      #   next()
      # 
      # }

global_maps_list <- list.files(file.path(impact_maps_dir, "impact_maps_by_taxon_across_ingredient"), pattern = glue("{allocation_type}_{taxa_type}"), recursive = TRUE, full.names = TRUE)

stack_rast <- rast()

for(file in global_maps_list){
  # file <- global_maps_list[[1]]

  rast_1 <- rast(file) %>%
    project(., moll_template, method = "near")

  stack_rast <- c(stack_rast, rast_1)
}

names(stack_rast) <- file_path_sans_ext(str_after_nth(global_maps_list, "\\/", 6))


all_df <- as.data.frame(stack_rast, xy = TRUE) %>%
  pivot_longer(cols = c(3:ncol(.)), values_to = "value", names_to = "type") %>%
  filter(!is.na(value)) %>%
  separate_wider_delim(type, delim = "/", names = c("diet", "fcr_type", "taxa_calc_type")) %>%
  separate_wider_delim(taxa_calc_type, delim = "_", names = c("allocation", "taxa", "calc_type")) %>%
  mutate(value_log =  ifelse(str_detect(calc_type, "mean"), log(value*1000000 + 1), value)) %>%
  rename(long = x, lat = y)

for(i in unique(all_df$calc_type)){

  # calc_type = "mean"

  # if(glue(file.path(impact_maps_dir, "csvs/impact_maps_by_taxon_across_ingredient/{allocation_type}_{taxa_type}_{i}.rds.gz")) %in% c(unique(done$f))){
  # 
  #   cat("file exists")
  #   next()
  # }

  loop_df <- all_df %>%
    filter(calc_type == i) %>%
    filter(!is.na(value))

  data.table::fwrite(loop_df, glue(file.path(impact_maps_dir, "csvs/impact_maps_by_taxon_across_ingredient/{allocation_type}_{taxa_type}_{i}.rds.gz")))

      }
    }
  }
}
     
  
# make and save plots across ingredients per taxon  

for(allocation_type in allocations){
    for(taxa_type in taxon){
      for(type in calc_types){

        # allocation_type = "economic"
        # taxa_type = "Bird"
        # type = "mean"
        
        all_df <- fread(glue(file.path(impact_maps_dir, "csvs/impact_maps_by_taxon_across_ingredient/{allocation_type}_{taxa_type}_{type}.rds.gz"))) %>%
          filter(!is.na(value))
        
diets <- unique(all_df$diet)
  
## create maps in separate plots, force common scale between them
maps <- map(.x = diets, 
            .f = function(x, y) 
  ggplot() + 
  geom_tile(data = all_df  %>% filter(diet == x), aes(x= long, y = lat, fill = value_log)) + 
      # scale_fill_gradientn(colors = YlOrBr(9), na.value = "white") +  # Set the color for missing values (NA) to white
   scale_fill_gradientn(colors = final_palette, na.value = "white") +
   #  scale_fill_viridis(direction = -1) + 
              geom_sf(data = countries_shp, fill = NA, colour = "grey") + 
    facet_wrap(~fcr_type) +
  theme_minimal() +
      theme(
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank())
            )


ggsave(glue(here(this_dir, "output/plots/by_taxa/{allocation_type}_{taxa_type}_{type}_plot.png")), plot = plot_grid(plotlist = maps, labels = glue("{diets}:{allocation_type}:{taxa_type}:{type}"), label_size = 10, nrow = 2), width = 10, height = 8, bg = "white")

    }
  }
}



# save delta plots across ingredients per taxon 

for(allocation_type in allocations){
    for(taxa_type in taxon){
      for(type in calc_types){
  
        # type = "mean"
        # allocation_type = "economic"
        # taxa_type = "arthropods"
  
  # if(file.exists(glue(here(this_dir, "output/plots/delta/by_taxa/{allocation_type}_{taxa_type}_{type}_plot.png")))){
  #   cat("exists")
  #   next()
  # }
  # 
        all_df <- fread(glue(file.path(impact_maps_dir, "csvs/impact_maps_by_taxon_across_ingredient/{allocation_type}_{taxa_type}_{type}.rds.gz"))) %>%
          filter(!is.na(value))

if(str_detect(type, "mean")){
  
  delta_df <- all_df %>%
      pivot_wider(names_from = c(diet, taxa), values_from = value) %>%
    rename_with(
      ~ case_when(
        str_detect(., "fish-dominant") ~ "fish-dominant",
        str_detect(., "plant-dominant") ~ "plant-dominant",
        TRUE ~ .)
      ) %>%
    mutate(`fish-dominant` = ifelse(is.na(`fish-dominant`), 0, `fish-dominant`),
           `plant-dominant` = ifelse(is.na(`plant-dominant`), 0, `plant-dominant`)) %>%
  mutate(delta = (`fish-dominant` - `plant-dominant`)) %>%
    mutate(delta_log = log(delta + 1)) %>%
    mutate(delta_rescale = delta*1000000)
  
}else{
    delta_df <- all_df %>%
      pivot_wider(names_from = c(diet, taxa), values_from = value) %>%
    rename_with(
      ~ case_when(
        str_detect(., "fish-dominant") ~ "fish-dominant",
        str_detect(., "plant-dominant") ~ "plant-dominant",
        TRUE ~ .)
      ) %>%
          mutate(`fish-dominant` = ifelse(is.na(`fish-dominant`), 0, `fish-dominant`),
           `plant-dominant` = ifelse(is.na(`plant-dominant`), 0, `plant-dominant`)) %>%
  mutate(delta = (`fish-dominant` - `plant-dominant`)) %>%
    mutate(delta_log = log(delta + 1)) %>%
    mutate(delta_rescale = delta*1000000)
}
  
data.table::fwrite(delta_df, glue(file.path(impact_maps_dir, "csvs/impact_maps_by_taxon_across_ingredient/delta/{allocation_type}_{taxa_type}_{type}.rds.gz")))

if(str_detect(type, "nspp")){
  limit <- c(min(delta_df$delta), max(delta_df$delta))

}else{
limit <- c(quantile(delta_df$delta, 0.05), quantile(delta_df$delta, 0.9))
}

## create maps in separate plots, force common scale between them
  ggplot() +
  geom_tile(data = delta_df %>% filter(delta != 0) , aes(x= long, y = lat, fill = delta)) +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,  limit = limit, na.value = "white", oob=scales::squish) + # squish makes sure anything above the scales limit will be a solid red or blue... better to visualize this way since the numbers are so small
              geom_sf(data = countries_shp, fill = NA, colour = "grey") +
    facet_wrap(~fcr_type) +
  theme_minimal() +
      theme(
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank()) +
    labs(caption = glue("{allocation_type}; {taxa_type}; {type}"))


ggsave(glue(here(this_dir, "output/plots/delta/by_taxa/{allocation_type}_{taxa_type}_{type}_plot.png")), width = 10, height = 8, bg = "white")

    }
  }
}
```

Global maps by ingredients across taxon

```{r}

allocations <- c("economic", "ge", "mass")
calc_types <- c("mean", "sd", "nspp")
ingredients <- all_overlap %>% 
  pull(ingredient) %>% unique()

## commented out code only needs to be run if the source impact maps are updated! 
for(allocation_type in allocations){
  for(ingredient_type in ingredients){

      # ingredient_type = "Soybean_soybean meal"
      # allocation_type = "economic"
      # type = "mean"

global_maps_list <- list.files(file.path(impact_maps_dir, "impact_maps_across_taxon_by_ingredient"), pattern = glue("{allocation_type}_{ingredient_type}"), recursive = TRUE, full.names = TRUE)

stack_rast <- rast()

for(file in global_maps_list){
  # file <- global_maps_list[[1]]

  rast_1 <- rast(file) %>%
    project(., moll_template, method = "near")

  stack_rast <- c(stack_rast, rast_1)
}

names(stack_rast) <- file_path_sans_ext(str_after_nth(global_maps_list, "\\/", 6))


all_df <- as.data.frame(stack_rast, xy = TRUE) %>%
  pivot_longer(cols = c(3:ncol(.)), values_to = "value", names_to = "type") %>%
  separate_wider_delim(type, delim = "/", names = c("diet", "fcr_type", "ingredient_calc_type")) %>%
    separate_wider_delim(ingredient_calc_type, delim = "_", names = c("allocation", "raw_material", "ingredient", "calc_type")) %>%
  mutate(value_log =  ifelse(str_detect(calc_type, "mean"), log(value*1000000 + 1), value)) %>%
  rename(long = x, lat = y) %>%
  filter(!is.na(value)) %>%
  unite("crop_ingredient", c(raw_material:ingredient), sep = "_", remove = FALSE)

for(i in unique(all_df$calc_type)){
  # i = "mean"

  loop_df <- all_df %>%
    filter(calc_type == i) %>%
    filter(!is.na(value))

  data.table::fwrite(loop_df, glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_by_ingredient/{allocation_type}_{ingredient_type}_{i}.rds.gz")))

   }
  }
}





for(allocation_type in allocations){
    for(ingredient_type in ingredients){
      for(type in calc_types){
        
      # ingredient_type = "Soybean_soybean meal"
      # allocation_type = "economic"
      # type = "mean"
        
                all_df <- fread(glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_by_ingredient/{allocation_type}_{ingredient_type}_{type}.rds.gz"))) 
 
diets <- unique(all_df$diet)
  
## create maps in separate plots, force common scale between them
maps <- map(.x = diets, 
            .f = function(x, y) 
  ggplot() + 
  geom_tile(data = all_df  %>% filter(diet == x), aes(x= long, y = lat, fill = value_log)) + 
   scale_fill_gradientn(colors = final_palette, na.value = "white") +
              geom_sf(data = countries_shp, fill = NA, colour = "grey") + 
    facet_wrap(~fcr_type) + 
  theme_minimal() +
      theme(
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank())
            )

ggsave(glue(here(this_dir, "output/plots/by_ingredient/{allocation_type}_{ingredient_type}_{type}_plot.png")), plot = plot_grid(plotlist = maps, labels = glue("{diets}:{allocation_type}:{ingredient_type}:{type}"), label_size = 10, nrow = 2), width = 10, height = 8, bg = "white")

    
    }
  }
}


# make delta plots! 

for(allocation_type in allocations){
    for(ingredient_type in ingredients){
      for(type in calc_types){
  
        # type = "mean"
        # allocation_type = "economic"
        # ingredient_type = "CropsNES_coconut oil"
  
  # if(file.exists(glue(here(this_dir, "output/plots/delta/by_ingredient/{allocation_type}_{ingredient_type}_{type}_plot.png")))){
  #   cat("exists")
  #   next()
  # }
  
        all_df <- fread(glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_by_ingredient/{allocation_type}_{ingredient_type}_{type}.rds.gz"))) %>%
          filter(!is.na(value)) %>%
          dplyr::select(long, lat, fcr_type, allocation, crop_ingredient, calc_type, diet, value) %>%
          filter(crop_ingredient == ingredient_type)

if(str_detect(type, "mean")){
  
  delta_df <- all_df %>%
      pivot_wider(names_from = c(diet, crop_ingredient), values_from = value) %>%
    rename_with(
      ~ case_when(
        str_detect(., "fish-dominant") ~ "fish-dominant",
        str_detect(., "plant-dominant") ~ "plant-dominant",
        TRUE ~ .)
      ) 
  
  if(!("fish-dominant" %in% c(colnames(delta_df)))){
    
    delta_df <- delta_df %>%
      mutate(`fish-dominant` = 0)
  }else if(!("plant-dominant" %in% c(colnames(delta_df)))){
        delta_df <- delta_df %>%
      mutate(`plant-dominant` = 0)
  }
  
  delta_df <- delta_df %>% 
  mutate(`fish-dominant` = coalesce(`fish-dominant`, 0),
         `plant-dominant` = coalesce(`plant-dominant`, 0)) %>%
  mutate(delta = (`fish-dominant` - `plant-dominant`)) %>%
    mutate(delta_log = log(delta + 1)) %>%
    mutate(delta_rescale = delta*1000000)
  
}else{
    delta_df <- all_df %>%
      pivot_wider(names_from = c(diet, crop_ingredient), values_from = value) %>%
    rename_with(
      ~ case_when(
        str_detect(., "fish-dominant") ~ "fish-dominant",
        str_detect(., "plant-dominant") ~ "plant-dominant",
        TRUE ~ .)
      ) 
  
  if(!("fish-dominant" %in% c(colnames(delta_df)))){
    
    delta_df <- delta_df %>%
      mutate(`fish-dominant` = 0)
  }else if(!("plant-dominant" %in% c(colnames(delta_df)))){
        delta_df <- delta_df %>%
      mutate(`plant-dominant` = 0)
  }
  
  delta_df <- delta_df %>% 
  mutate(`fish-dominant` = coalesce(`fish-dominant`, 0),
         `plant-dominant` = coalesce(`plant-dominant`, 0)) %>%
  mutate(delta = (`fish-dominant` - `plant-dominant`)) %>%
    mutate(delta_log = log(delta + 1)) %>%
    mutate(delta_rescale = delta*1000000)
}
  
  
  data.table::fwrite(loop_df, glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_by_ingredient/{allocation_type}_{ingredient_type}_{type}.rds.gz")))

        
if(str_detect(type, "nspp")){
  limit <- c(min(delta_df$delta), max(delta_df$delta))

}else{
limit <- c(quantile(delta_df$delta, 0.05), quantile(delta_df$delta, 0.9))
}

## create maps in separate plots, force common scale between them
  ggplot() +
  geom_tile(data = delta_df %>% filter(delta != 0) , aes(x= long, y = lat, fill = delta)) +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,  limit = limit, na.value = "white", oob=scales::squish) +
              geom_sf(data = countries_shp, fill = NA, colour = "grey") +
    facet_wrap(~fcr_type) +
  theme_minimal() +
      theme(
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank()) +
    labs(caption = glue("{allocation_type}; {ingredient_type}; {type}"))


ggsave(glue(here(this_dir, "output/plots/delta/by_ingredient/{allocation_type}_{ingredient_type}_{type}_plot.png")), width = 10, height = 8, bg = "white")

    }
  }
}

```

Global maps by raw material across taxon

```{r}

allocations <- c("economic", "ge", "mass")
calc_types <- c("mean", "sd", "nspp")
ingredients <- all_overlap %>% pull(ingredient) %>% unique()
raw_mats <- all_overlap %>% mutate(raw_material = str_before_first(ingredient, "_")) %>% 
  # filter(!(raw_material %in% c("forage fish", "trimmings fish"))) %>% 
  pull(raw_material) %>% unique()

## commented out code only needs to be run if the source impact maps are updated! 
for(allocation_type in allocations){
  for(raw_mat in raw_mats){
    for(type in calc_types){

    # ingredient_type = "Soybean_soybean meal"
    # allocation_type = "economic"
    # type = "mean"
    #  raw_mat = "CropsNES"

    all_files <- list.files(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_by_ingredient"), pattern = glue("{allocation_type}_{raw_mat}_.*._{type}"), full.names = TRUE)


all_df <- map_dfr(all_files, ~fread(.x), header = TRUE)

all_df_fin <- all_df %>% 
  mutate(value_log = ifelse(calc_type == "mean", (exp(value) - 1) / 1000000, value)) %>%
  group_by(long, lat, diet, fcr_type, allocation, raw_material, calc_type) %>% 
  summarise(value = sum(value, na.rm = TRUE),
            value_log = sum(value_log, na.rm = TRUE)) %>%
  ungroup()

# test <- all_df_fin %>% group_by(diet, fcr_type, allocation) %>% summarise(sum = sum(value, na.rm = TRUE))


data.table::fwrite(all_df_fin, glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_by_raw_material/{allocation_type}_{raw_mat}_{type}.rds.gz")))


delta_df <- all_df_fin %>%
        pivot_wider(names_from = c(diet), values_from = value) 


  if(!("fish-dominant" %in% c(colnames(delta_df)))){
    
    delta_df <- delta_df %>%
      mutate(`fish-dominant` = 0)
  }else if(!("plant-dominant" %in% c(colnames(delta_df)))){
        delta_df <- delta_df %>%
      mutate(`plant-dominant` = 0)
  }

delta_df <- delta_df %>% 
  mutate(delta = (`fish-dominant` - `plant-dominant`)) 

data.table::fwrite(delta_df, glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_by_raw_material/delta/{allocation_type}_{raw_mat}_{type}.rds.gz")))

    }
  }
}



for(allocation_type in allocations){
    for(raw_mat in raw_mats){
      for(type in calc_types){
        
      # raw_mat = "forage fish"
      # allocation_type = "economic"
      # type = "nspp"
      
                all_df <- fread(glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_by_raw_material/{allocation_type}_{raw_mat}_{type}.rds.gz"))) 
 
diets <- unique(all_df$diet)
  
## create maps in separate plots, force common scale between them
maps <- map(.x = diets, 
            .f = function(x, y) 
  ggplot() + 
  geom_tile(data = all_df  %>% filter(diet == x), aes(x= long, y = lat, fill = value_log)) + 
   scale_fill_gradientn(colors = final_palette, na.value = "white") +
              geom_sf(data = countries_shp, fill = NA, colour = "grey") + 
    facet_wrap(~fcr_type) + 
  theme_minimal() +
      theme(
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank())
            )

ggsave(glue(here(this_dir, "output/plots/by_material/{allocation_type}_{raw_mat}_{type}_plot.png")), plot = plot_grid(plotlist = maps, labels = glue("{diets}:{allocation_type}:{raw_mat}:{type}"), label_size = 10, nrow = 2), width = 10, height = 8, bg = "white")

    
    }
  }
}


# make delta plots! 

for(allocation_type in allocations){
    for(raw_mat in raw_mats){
      for(type in calc_types){
  
        # type = "mean"
        # allocation_type = "economic"
        # raw_mat = "Soybean"
  
  # if(file.exists(glue(here(this_dir, "output/plots/delta/by_ingredient/{allocation_type}_{ingredient_type}_{type}_plot.png")))){
  #   cat("exists")
  #   next()
  # }
  
        delta_df <- fread(glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_by_raw_material/delta/{allocation_type}_{raw_mat}_{type}.rds.gz"))) %>%
          filter(!is.na(delta)) %>%
          dplyr::select(long, lat, fcr_type, allocation, raw_material, calc_type, delta)
  
if(str_detect(type, "nspp")){
  limit <- c(min(delta_df$delta), max(delta_df$delta))
  
}else{
limit <- c(quantile(delta_df$delta, 0.05), quantile(delta_df$delta, 0.9))
}

## create maps in separate plots, force common scale between them
  ggplot() + 
  geom_tile(data = delta_df %>% filter(delta != 0) , aes(x= long, y = lat, fill = delta)) +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,  limit = limit, na.value = "white", oob=scales::squish) +
              geom_sf(data = countries_shp, fill = NA, colour = "grey") + 
    facet_wrap(~fcr_type) + 
  theme_minimal() +
      theme(
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank()) + 
    labs(caption = glue("{allocation_type}; {raw_mat}; {type}"))
            

ggsave(glue(here(this_dir, "output/plots/delta/by_material/{allocation_type}_{raw_mat}_{type}_plot.png")), width = 10, height = 8, bg = "white")

    }
  }
}

```
