---
title: "Summary plots: maps"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Setup

Load libraries

```{r setup, include=FALSE}

library(tidyverse)
library(here)
library(data.table)
library(terra)
library(rnaturalearth)
library(strex)
library(janitor)
library(sf)
library(tools)
library(patchwork)
library(cowplot)
library(glue)
library(RColorBrewer)
library(ggpubr)
library(smoothr)
library(ggtext)
library(ggnewscale)

source(here("src/directories.R"))

source(here("src/spatial.R"))

setwd(dirname(rstudioapi::getSourceEditorContext()$path)) # set working directory to where this script is located
this_dir <- getwd()

biodiv_dir <- file.path(rdsi_dir, "biodiversity_impacts")

impact_maps_dir <- file.path(biodiv_dir, "output")

globe_bbox <- rbind(c(-180, -90), c(-180, 90), 
                      c(180, 90), c(180, -90), c(-180, -90)) 

globe_border <- st_polygon(list(globe_bbox)) %>%
  st_sfc(crs = 4326) %>%
  st_sf(data.frame(rgn = 'globe', geom = .)) %>%
  smoothr::densify(max_distance = 0.5) %>%
  st_transform(crs = crs(moll_template))


```

```{r establish color palettes and themes}

## establish color palette from food footprint project
d_brown <- "#515256"
m_brown <- "#B27B54"
l_brown <- "#BC995F"
green <- "#A8B072"
yellow<- "#EFCE71"
l_yellow <- "#F7F6C1"

jv_pal <- c(l_yellow, yellow, green, l_brown, m_brown, d_brown)

continuous_pal <-  colorRampPalette(jv_pal, space="Lab", bias = 3.5)(10000)
image(volcano, asp=1, col=continuous_pal)

light_gray <- "#F8F9FA"
final_palette <- c(light_gray, continuous_pal)
image(volcano, asp=1, col=final_palette)

palette_red <- c(final_palette, "#B90000")
image(volcano, asp=1, col=palette_red)


countries_shp <- ne_countries(scale = 110, returnclass = "sf") |> 
  st_transform(crs(moll_template)) # read in countries shapefile

theme_ohara <- function(base_size = 9) {
  theme(axis.ticks = element_blank(),
        text             = element_text(family = 'Helvetica',
                                        color = 'gray30', size = base_size),
        plot.title       = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
        panel.background = element_blank(),
        legend.position  = 'right',
        panel.border     = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(colour = 'grey90', size = .25),
        legend.key       = element_rect(colour = NA, fill = NA),
        axis.line        = element_blank())
}


```

Read in marine and terrestrial dfs and combine

```{r}
marine_overlap <- readRDS(here("prep/03_prep_spp_habitats/output/aoh_marine_impact.rds")) %>%
    dplyr::select(sciname, spp_type, diet, fcr_type, allocation, ingredient, vulnerability, impact_total, total_hab_area)

terrestrial_overlap <- readRDS(here("prep/03_prep_spp_habitats/output/aoh_terrestrial_impact.rds")) %>% 
    dplyr::select(sciname, spp_type, diet, fcr_type, allocation, ingredient, vulnerability, impact_total, total_hab_area)

all_overlap <- rbind(marine_overlap, terrestrial_overlap)

```


Make map split by quantiles in each cell
 - regular quantiles up to 95th 
 - Then 95-99.9th
 - then >= 99.9th
 
Additionally, make a delta map, and color either fish dominant or plant dominant. Have the legend colors be the same as the green and blue we use for the boxplot, but split by quantile of delta (same as regular impact quantiles above).
 
 
```{r}
# make publication ready mean extinction risk plots with all feed scenarios, economic allocation
 filepath <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/economic_mean_prop.rds.gz"))

quantile_pal <- c(
    "0-24" = "#D4F7F0", 
    "25-49" = "#538AC6", 
    "50-74" = "#1E438F", 
    "75-94" = "#F8D422", 
    "95-99.9" = "#E95D08", 
    "≥99.9" = "#781211"
)

all_df <-  fread(filepath) %>%
  filter(fcr_type == "regular") %>%
    filter(!is.na(value)) %>%
    mutate(diet_cap = str_to_sentence(diet))


quantiles <- all_df %>%
  group_by(diet) %>%
  summarise(q25 = quantile(value, 0.25),
            q50 = quantile(value, 0.50),
            q75 = quantile(value, 0.75),
            q95 = quantile(value, 0.95),
            q999 = quantile(value, 0.999)) %>%
  ungroup()

all_df_quant <- all_df %>%
  left_join(quantiles, by = "diet") %>%
    mutate(category = case_when(
    value < q25 ~ "0-24",
    value >= q25 & value < q50 ~ "25-49",
    value >= q50 & value < q75 ~ "50-74",
    value >= q75 & value < q95 ~ "75-94",
    value >= q95 & value < q999 ~ "95-99.9",
    value >= q999 ~ "≥99.9"
  ))

all_df_quant$category <- factor(all_df_quant$category, levels = c("0-24", "25-49", "50-74", "75-94", "95-99.9", "≥99.9"))

  quantile_plot <-  ggplot() + 
      geom_tile(data = all_df_quant, aes(x = long, y = lat, fill = category, color = category)) +
                 geom_sf(data = countries_shp, fill = NA, colour = "grey", alpha = 0.2)  +
                 geom_sf(data = globe_border, fill = NA, colour = "grey") + 
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_minimal() + 
      theme(plot.title = element_text(hjust = 0.5), 
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.margin = margin(r = 10), # add a margin to legend text so it isn't cut off when saving 
          legend.position = "bottom",
          text = element_text(family = "Arial", colour = "black"),
          legend.title = element_text(size = 8),
          strip.text.x = element_text(size = 8, family = "Arial", colour = "black"),
          legend.key.size = unit(4, "mm")) + # change legend box sizes
  labs(fill = "Impact Quantile") + 
     guides(color = "none",
             fill = guide_legend(title.position = "bottom", title.hjust = 0.5, nrow = 1)) + 
    facet_wrap(~diet_cap) + 
    scale_fill_manual(values = quantile_pal) +
    scale_colour_manual(values = quantile_pal) 
  
    ggsave(glue(here(this_dir, "output/plots/publication/SI/economic_mean_prop_quantiles.png")), plot = quantile_plot, width = 8, height = 8, bg = "white")
    
    
## Delta plot 
    
 filepath <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/delta/economic_mean_prop.rds.gz"))

delta_df <-  fread(filepath) %>%
  filter(fcr_type == "regular", 
         delta_cat != "No difference")


quantiles_delta_fish <- delta_df %>%
  filter(delta_cat == "Fish-dominant") %>%
  group_by(delta_cat) %>%
  summarise(q25 = quantile(delta, 0.25),
            q50 = quantile(delta, 0.50),
            q75 = quantile(delta, 0.75),
            q95 = quantile(delta, 0.95),
            q999 = quantile(delta, 0.999)) %>%
  ungroup()

quantiles_delta_plant <- delta_df %>%
  filter(delta_cat == "Plant-dominant") %>%
  group_by(delta_cat) %>%
  summarise(q25 = quantile(-1*delta, 0.25),
            q50 = quantile(-1*delta, 0.50),
            q75 = quantile(-1*delta, 0.75),
            q95 = quantile(-1*delta, 0.95),
            q999 = quantile(-1*delta, 0.999)) %>%
  ungroup() %>%
  mutate(across(2:6, ~.x*-1))

quantiles_delta <- rbind(quantiles_delta_plant, quantiles_delta_fish)

delta_df_quant <- delta_df %>%
  left_join(quantiles_delta, by = "delta_cat") %>%
    mutate(category = case_when(
    abs(delta) < abs(q25) ~ "0-24",
    abs(delta) >= abs(q25) & abs(delta) < abs(q50) ~ "25-49",
    abs(delta) >= abs(q50) & abs(delta) < abs(q75) ~ "50-74",
    abs(delta) >= abs(q75) & abs(delta) < abs(q95) ~ "75-94",
    abs(delta) >= abs(q95) & abs(delta) < abs(q999) ~ "95-99.9",
    abs(delta) >= abs(q999) ~ "≥99.9"
  ))


delta_pal <- c(
  "Fish-dominant" = "#0D6EB5",
  "Plant-dominant" = "#7BD67F"
)

plant_pal <-  colorRampPalette(brewer.pal(n=9, name = "Greens"))(7)

plant_pal <- c(
      "0-24" = "#DBF1D5", 
    "25-49" = "#ADDEA7", 
    "50-74" = "#74C476", 
    "75-94" = "#37A055", 
    "95-99.9" = "#0B7734", 
    "≥99.9" = "#00441B"
)

fish_pal <- colorRampPalette(brewer.pal(n=9, name = "Blues"))(7)

fish_pal <- c(
      "0-24" = "#D6E5F4", 
    "25-49" = "#ABCFE5", 
    "50-74" = "#6BAED6", 
    "75-94" = "#3787C0", 
    "95-99.9" = "#105BA4", 
    "≥99.9" = "#08306B"
)

delta_df_quant_fis <- delta_df_quant %>% filter(delta_cat == "Fish-dominant")
delta_df_quant_fis$category <- factor(delta_df_quant_fis$category, levels = c("0-24", "25-49", "50-74", "75-94", "95-99.9", "≥99.9"))


delta_df_quant_plant <- delta_df_quant %>% filter(delta_cat == "Plant-dominant")
delta_df_quant_plant$category <- factor(delta_df_quant_plant$category, levels = c("0-24", "25-49", "50-74", "75-94", "95-99.9", "≥99.9"))


   delta_plot <-  ggplot() + 
      geom_tile(data = delta_df_quant_plant, aes(x = long, y = lat, fill = category, color = category)) +
        scale_fill_manual(values = plant_pal) +
    scale_colour_manual(values = plant_pal) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
     theme_ohara() + 
  labs(fill = "Plant-dominant") +
  guides(color = "none") +
          theme(legend.position = "none", 
                text = element_text(family = "Arial")) + 
     new_scale_color() +
     new_scale_fill() + 
           geom_tile(data = delta_df_quant_fis, aes(x = long, y = lat, fill = category, color = category)) +
             scale_fill_manual(values = fish_pal) +
                 geom_sf(data = countries_shp, fill = NA, colour = "grey", alpha = 0.2)  +
                 geom_sf(data = globe_border, fill = NA, colour = "grey") + 
    scale_colour_manual(values = fish_pal) +
       labs(fill = "Fish-dominant", x = "", y = "") +
  guides(color = "none") + 
     theme(legend.position = "none", 
           axis.text.y = element_blank(), 
           axis.text.x = element_blank(), 
           plot.margin = margin(t = 10, r = 10, b = 50, l = 10, unit = "pt"),
           text = element_text(family = "Arial", colour = "black"))
   
   fish_legend <- get_legend(ggplot() + 
      geom_tile(data = delta_df_quant_fis, aes(x = long, y = lat, fill = category, color = category)) +
             scale_fill_manual(values = fish_pal) +
              scale_colour_manual(values = fish_pal) +
                  labs(fill = "Fish-dominant") +
          guides(color = "none",
            fill = guide_legend(title.position = "bottom", title.hjust = 0.5, nrow = 1)) +
          theme(legend.position = "bottom", 
                  legend.title = element_text(size = 8),
                legend.text = element_text(size = 8), 
      text = element_text(family = "Arial", colour = "black"),
      legend.margin = margin(r = 10),
      legend.key.size = unit(4, "mm")))
   
   plant_legend <- get_legend(ggplot() + 
      geom_tile(data = delta_df_quant_plant, aes(x = long, y = lat, fill = category, color = category)) +
             scale_fill_manual(values = plant_pal) +
              scale_colour_manual(values = plant_pal) +
                  labs(fill = "Plant-dominant") +
          guides(color = "none",
            fill = guide_legend(title.position = "bottom", title.hjust = 0.5, nrow = 1)) +
          theme(legend.position = "bottom",
                legend.title = element_text(size = 8),
                legend.text = element_text(size = 8), 
                   text = element_text(family = "Arial", colour = "black"), 
      legend.margin = margin(r = 10),
      legend.key.size = unit(4, "mm")))
     
   full_plot <- ggdraw() +
                draw_plot(delta_plot) +
                draw_plot(fish_legend, height = 0.15, width = 0.52) +
                draw_plot(plant_legend, height = 0.15, width = 1.54)
     
   
   
        ggsave(glue(here(this_dir, "output/plots/publication/economic_mean_prop_quantiles_delta.png")),
                plot = plot_grid(quantile_plot, full_plot, nrow = 2, labels = c("A", "B"), label_fontfamily = "Arial", label_fontface = "plain", label_size = 8), 
                width = 180, height = 180, bg = "white", units = "mm")
      
         ggsave(glue(here(this_dir, "output/plots/publication/SI/delta_plot.png")),
                plot = full_plot, 
                width = 180, height = 180, bg = "white", units = "mm")
   
    
```

Histogram of impact values by quantile group 

```{r}
hist_plot_fish <- ggplot(data = all_df_quant %>% filter(diet == "fish-dominant")) + 
    geom_histogram(aes(value, fill = category)) +
    facet_wrap(~category, scales = "free", ncol = 2) +
    theme_ohara() +
    labs(x = "Average proportion of habitat impacted", y = "Number of cells", title = "Fish-dominant") +
    scale_fill_manual(values = quantile_pal) +
  theme(legend.position = "none")

hist_plot_plant <- ggplot(data = all_df_quant %>% filter(diet == "plant-dominant")) + 
    geom_histogram(aes(value, fill = category)) +
    facet_wrap(~category, scales = "free", ncol = 2) +
    theme_ohara() +
    labs(x = "Average proportion of habitat impacted", y = "Number of cells", title = "Plant-dominant") +
    scale_fill_manual(values = quantile_pal) + 
    theme(legend.position = "none")

  
  ggsave(glue(here(this_dir, "output/plots/publication/SI/economic_mean_prop_quantiles_hist_plant.png")), plot = plot_grid(hist_plot_plant , nrow = 1), width = 8, height = 8, bg = "white") 
  
    ggsave(glue(here(this_dir, "output/plots/publication/SI/economic_mean_prop_quantiles_hist_fish.png")), plot = plot_grid(hist_plot_fish, nrow = 1), width = 8, height = 8, bg = "white") 
    
```


Make map showing actual values
```{r}
# make publication ready mean extinction risk plots with all feed scenarios, economic allocation
 filepath <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/economic_mean_prop.rds.gz"))

all_df <-  fread(filepath) %>%
  filter(fcr_type == "regular")

 plot_list <- list()
  
 diets <- unique(all_df$diet)
 
 for(diet_type in diets){

      # diet_type = "plant-dominant"
      
  loop_loop_df <- all_df %>%
    filter(diet == diet_type)
  
quantile_value <- quantile(loop_loop_df$value, probs = 0.999, na.rm = TRUE)
  
  loop_loop_df <- loop_loop_df %>% 
    mutate(value_new = ifelse(value > quantile_value, quantile_value, value))
  
  loop_plot <-  ggplot() + 
      geom_tile(data = loop_loop_df, aes(x = long, y = lat, fill = value_new, color = value_new)) +
     geom_tile(fill = ifelse(loop_loop_df$value_new < quantile_value, "#B90000", NA)) + 
            geom_sf(data = countries_shp, fill = NA, colour = "grey") + 
       scale_fill_gradientn(colors = palette_red, na.value = "white") +
            scale_color_gradientn(colors = palette_red, na.value = "white") +
            geom_sf(data = globe_border,
            fill = NA, color = 'grey70', 
            size = .1) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_void() +
      theme(plot.title = element_text(hjust = 0.5), 
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.margin = margin(r = 10)) + # add a margin to legend text so it isn't cut off when saving 
  labs(fill = "Proportion\nhabitat\nimpacted") + 
     guides(color = "none") + 
    facet_wrap(~diet)
  
        
      plot_list[[length(plot_list) + 1]] <- loop_plot

  } 

ggsave(glue(here(this_dir, "output/plots/publication/SI/economic_mean_prop.png")), plot = plot_grid(plotlist = plot_list, nrow = 1), width = 10, height = 4, bg = "white")
```


Map stats!

```{r}

## read in plant-dominant, economic, regulat
 filepath <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/economic_mean_prop.rds.gz"))

all_df <-  fread(filepath)  %>%
  filter(value > 0) %>%
  filter(!is.na(value))

max(all_df$value) # 0.035490633.5

moll_template_land <- readRDS(here("prep/03_prep_spp_habitats/data/spatial/moll_template_land_xy.rds"))
moll_template_ocean <- readRDS(here("prep/03_prep_spp_habitats/data/spatial/moll_template_ocean_xy.rds"))


maximum_location <- all_df %>%
  filter(value > 0.02) %>%
  left_join(land_eez_rgns, by = c("long" = "x", "lat" = "y")) %>%
 left_join(moll_template_land, by = c("long" = "x", "lat" = "y")) # max value is 0.035 on land in Norway

test <- all_df %>%
  group_by(diet, fcr_type) %>%
  summarise(prop_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

scenario_summary <- all_df %>% 
  filter(value > 0) %>%
  group_by(diet, fcr_type) %>%
  summarise(cells_impacted = n()) %>%
  ungroup()

moll_land <- readRDS(here("prep/03_prep_spp_habitats/data/spatial/moll_template_land_xy.rds"))

all_df_land <- all_df %>%
  left_join(moll_land, by = c("long" = "x", "lat" = "y")) %>%
  filter(!is.na(cell_id)) %>%
  filter(value > 0) %>%
    group_by(diet, fcr_type) %>%
  summarise(cells_impacted = n()) %>%
  ungroup() %>%
   mutate(land_cells_total = 2876998) %>%
  mutate(prop = cells_impacted/land_cells_total)

moll_ocean <- readRDS(here("prep/03_prep_spp_habitats/data/spatial/moll_template_ocean_xy.rds"))

all_df_ocean <- all_df %>%
  left_join(moll_ocean, by = c("long" = "x", "lat" = "y")) %>%
  filter(!is.na(cell_id)) %>%
    filter(value > 0) %>%
    group_by(diet, fcr_type) %>%
  summarise(cells_impacted = n()) %>%
  ungroup() %>%
  mutate(ocean_cells_total = 3684240) %>%
  mutate(prop = cells_impacted/ocean_cells_total)

# Comparing the fish-dominant and plant-dominant feed scenarios, under the regular FCR scenario, we see that the plant-dominant diet has 82,162 more pixels impacted (there are ~6.5 million pixels on a 10km by 10 km Mollweide grid). This difference is almost exclusively due to the difference in terrestrial crop impacts, with the plant-dominant diet having 81,990 more terrestrial pixels impacted than that of the fish-dominant diet. Overall, ocean pixels are impacted over a larger area: ~42% of ocean pixels are impacted, regardless of feed scenario. In comparison, 22%-25% of land pixels are impacted, depending on the scenario. 

```




