---
title: "MapSPAM_wrangling"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Summary

Wrangle MapSPAM crop category dataset.

# Data sources

 - FAO crop production

 - MapSPAM
 
 - Halpern et al. 2022


# Setup

```{r setup, include = FALSE}
# getting packages we want
library(tidyverse)
library(here)
library(raster)
library(sf)
library(vroom)
library(janitor)

# Raster templates
source(here("src/spatial.R"))
source(here("src/fxns.R"))
source(here("src/directories.R"))

``` 


# Methods

FAOSTAT coffee #656 is both arabica and robusta coffee, so Millet #79 is pmil and smil

```{r}
# Match with MapSPAM names
MapSPAM_crops <- 
  read.csv(here("prep/01_crops_mapspam/data/MapSPAM_crop_info.csv")) %>% 
  rename(item_code = FAOCODE)

# Filter out non-item_code of interest according to MapSPAMS definitions of crop categories.
MapSPAM_crops$item_code <- gsub(" - ", ":", MapSPAM_crops$item_code, fixed = TRUE)
MapSPAM_crops$item_code <- gsub("-", ":", MapSPAM_crops$item_code, fixed = TRUE)
MapSPAM_crops$item_code[[27]] <- "260:310,312:339"
MapSPAM_crops$item_code <- gsub(" ", "", MapSPAM_crops$item_code, fixed = TRUE)
```

# Splitting data 
```{r}
# split data 
MapSPAM_crops <- 
  MapSPAM_crops %>% 
  mutate(item_code = strsplit(as.character(item_code), ",")) %>%
  unnest(item_code)

# Remove 'rest' crops (677:839) that are nested inside ofib range (773:821).
MapSPAM_crops <- 
  MapSPAM_crops %>% 
  mutate(item_code = if_else(item_code %in% "677:839", "677:772", item_code)) %>%  
  bind_rows(MapSPAM_crops) %>%
  mutate(item_code = if_else(item_code %in% "677:839", "822:839", item_code)) %>%  
  unique() 

MapSPAM_ranges <- MapSPAM_crops %>% 
  filter(str_detect(item_code, pattern = ":")) 
MapSPAM_remainder <- MapSPAM_crops %>% 
  filter(!str_detect(item_code, pattern = ":"))
MapSPAM_remainder$item_code <- as.integer(unlist(MapSPAM_remainder$item_code))

# Replace number ranges with separate numbers.
get_numbers <- function(x, y) {
    item_code <- (x:y)
    return(as.vector(item_code))
}

MapSPAM_ranges <- 
  MapSPAM_ranges %>% 
    filter(str_detect(item_code, pattern = ":")) %>% 
    separate(item_code,
             into = c("x", "y"),
             sep = ":") %>% 
    mutate(x = as.integer(x),
           y = as.integer(y)) %>% 
    mutate(item_code = map2(x, y , get_numbers)) %>% 
    unnest(item_code) %>%
    dplyr::select(-c(x, y)) 

# These are double counted
MapSPAM_ranges <- MapSPAM_ranges %>% filter(!item_code %in% c(267, 270, 289, 292, 328, 826))

# Merge back together
MapSPAM_names <- rbind(MapSPAM_remainder, MapSPAM_ranges)
```

# Check for duplicates
```{r}
MapSPAM_names$item_code[duplicated(MapSPAM_names$item_code)] # This should only say true for `79` (Millet) and `656` (Coffee). 
```

# Save

```{r}
write_csv(MapSPAM_names, here("prep/01_crops_mapspam/data/MapSPAM_names.csv"))
```

# Make FAO crop dataset, with values for 2010 (MapSPAM year) and 2021 (latest FAO reporting year)

```{r}
fao_crop_prod_raw <- readRDS(here("data/tidy_data/production-data/crops_production_tidy.rds"))

fao_crop_prod_2010_2021 <- fao_crop_prod_raw %>%
  dplyr::filter(element_code == 5510,
                year_code %in% c(2009, 2010, 2011, 2019, 2020, 2021)) %>%
  dplyr::select(sector, area_code, area, element_code, element, item_code, item, year_code, year, value)

write.csv(fao_crop_prod_2010_2021, here("prep/01_crops_mapspam/data/FAOSTAT_crop_production_2010_2021.csv"), row.names = FALSE)
```



# Check what crops are in each category.

```{r}
dsdf <- 
  vroom::vroom(here("prep/01_crops_mapspam/data/FAOSTAT_crop_production_2010_2021.csv")) %>% 
  clean_names() %>% 
  dplyr::select(item, item_code) %>% 
  unique() %>% 
  left_join(MapSPAM_names) 

dsfdsf <- vroom::vroom(here("prep/01_crops_mapspam/data/FAO_crop_definitions.csv")) %>% 
  clean_names() %>% 
  dplyr::select(item_code, item, description) %>% 
  right_join(MapSPAM_names)
```

# Add super categories
```{r}
crop_codes <- 
  vroom::vroom(here("prep/01_crops_mapspam/data/MapSPAM_names.csv")) %>% 
  dplyr::select(SPAM_short_name, SPAM_full_name, item_code) %>% 
  left_join(vroom::vroom(here("prep/02_feed/data/MapSPAM_crop_info.csv"), 
                         col_select = c("SPAM_short_name", "SPAM_super"))) %>%
    mutate(SPAM_super = case_when(
    SPAM_short_name %in% c("sunf", "ooil", "rape", "opul") ~ SPAM_short_name, 
    TRUE ~ SPAM_super
  ))

str(crop_codes)

vroom::vroom_write(crop_codes, 
                   here("prep/01_crops_mapspam/data/crop_codes.csv"))
```

