---
title: "MapSPAM_scaling_2010_to_2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Script Description

Rescale MapSPAM data to year 2021.

# Data information

We do this using FAOSTAT crop production data trends between year 2010 and 2021.

# Setup
```{r setup, include = FALSE}
# Raster templates
library(here)
library(janitor)
library(terra)

source(here("src/spatial.R"))
source(here("src/fxns.R"))
source(here("src/directories.R"))


food_rgns <- read.csv(here("data/spatial/output/food_rgns.csv"))

```

# Methods 

## Import and wrangle FAO data for 2010 and 2021

```{r}
# SPAM names
crops <- 
  vroom(here("prep/01_crops_mapspam/data/crop_codes.csv"))  %>% 
  dplyr::select(SPAM_short_name, item_code) 

scaling_coef_tidy <- 
  vroom::vroom(here("prep/01_crops_mapspam/data/FAOSTAT_crop_production_2010_2021.csv")) %>% 
  clean_names() %>% 
  dplyr::select(area_code, item_code, year, value) %>% 
  rename(tonnes = value) %>% 
  na.omit() %>%
  filter(item_code %in% crops$item_code)
```

## Special Sudanese case

Sudan was split into two countries in 2010, so now we need to scale the difference from 2010 to 2021. To do this, I am extracting production data for SPAM's 2010 map as opposed to splitting FAO data.

```{r}
SDN_SSD_2010 <- 
  vroom::vroom(here("prep/01_crops_mapspam/data/prod_crop_rgns_2010.csv")) %>% 
  filter(iso3c %in% c("SSD", "SDN")) %>% 
  group_by(iso3c, SPAM_short_name) %>% 
  summarize(`2010` = sum(production)) %>% 
  ungroup() 

SDN_SSD_2021 <- 
  scaling_coef_tidy %>% 
  filter(area_code %in% c(276, 277)) %>% # Filter Sudan
  left_join(crops) %>%   
  left_join(read_csv(here("data/spatial/output/FAO_rgn_codes.csv")), 
            by = "area_code") %>% # Add iso3c
  group_by(SPAM_short_name, iso3c) %>% 
  summarise(`2021` = sum(tonnes, na.rm = TRUE)) %>% 
  ungroup() %>% 
  na.omit()

sudan_coef <- 
  full_join(SDN_SSD_2010, SDN_SSD_2021) %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  mutate(coef = `2021` / `2010`) %>% 
  dplyr::select(!contains("tonnes")) %>%
  mutate_all(~replace(., is.na(.), 0)) %>% 
  mutate(coef = ifelse(coef == Inf, 0, coef))
```

# Wrangle FAO dataset 
```{r}
scaling_coef_tidy <- 
  scaling_coef_tidy %>% 
  mutate(area_code = if_else(area_code == 206, 277, area_code)) %>% # Re-allocate former Sudan's 2010 values.
  bind_rows(scaling_coef_tidy) %>%
  mutate(area_code = if_else(area_code == 206, 276, area_code)) %>% 
  unique() %>% 
  filter(!area_code %in% 351) %>% # Remove China
  left_join(read_csv(here("data/spatial/output/FAO_rgn_codes.csv")), 
            by = "area_code") %>% # Add iso3c
  left_join(crops) %>% # Add SPAM_short_name
  dplyr::select(year, tonnes, iso3c, SPAM_short_name) %>% 
  na.omit() # Remove NA created by milled rice.

# Distinct area_codes
n_distinct(scaling_coef_tidy$iso3c) # 200

missing_rgns <- setdiff(food_rgns$iso3c, scaling_coef_tidy$iso3c)
missing_rgns
food_rgns %>% filter(iso3c %in% missing_rgns) # Nothing unusual, all small islands.
```

# Calculate scaling from 2010 to 2021 for all crops and countries.
```{r}
scaling_coef <- 
  scaling_coef_tidy %>% 
  group_by(SPAM_short_name, year, iso3c) %>%
  summarize(tonnes = mean(tonnes)) %>% # Calculate mean tonnes for each crop, year and country.
  ungroup() %>% 
  spread(., year, tonnes) %>% 
  filter(!iso3c %in% c("SDN", "SSD")) %>% # Replace Sudan values
  mutate_all(~replace(., is.na(.), 0)) %>% 
  group_by(SPAM_short_name, iso3c, `2021`, `2010`) %>%
  summarize(coef = `2021` / `2010`, .groups = "keep") %>% 
  ungroup() %>% 
  bind_rows(sudan_coef)

# Make sure that countries with more than 0 production in SPAM 2010 but
# no data in 2021 have 0 value for 2021.
scaling_coef$coef[scaling_coef$`2010` > 0 & scaling_coef$`2021` == 0] <- 0

# Some countries will have data in 2017, according to the FAO, but none in 2010. We need to make sure that we manually insert these values in the scaled maps.

inf_values <- 
  scaling_coef %>% 
  filter(`2010` == 0 & `2021` > 0) %>% 
  dplyr::select(SPAM_short_name, iso3c, tonnes = `2021`)
inf_values

scaling_coef$coef[scaling_coef$`2010` == 0 & scaling_coef$`2021` > 0] <- 0

# All infinite numbers assigned a coef of 1.
scaling_coef$coef[which(!is.finite(scaling_coef$coef))] <- 0
```

# Save
```{r}
vroom_write(inf_values, here("prep/01_crops_mapspam/data/inf_values.csv"))
write_csv(scaling_coef, here("prep/01_crops_mapspam/data/scaling_coef.csv"))

test <- read.csv(here("prep/01_crops_mapspam/data/scaling_coef.csv"))
```

-------------------------------------------------------------------------------
