---
title: "Summary plots: maps"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Summary

The code here produces SI Figure 3:  
> SI Figure 3. (A) Map of mean rarity weighted proportion of habitat area impacted across species. Impacts are split by quantile group and aquafeed formulation. (B) Map of differences in mean rarity weighted impact between the feed formulation scenarios. The magnitude of the difference is split into quantile groups, where darker shades indicate larger differences. Related to Figure 1. 

Additionally, we save some supplementary information plots of the values in histograms and maps of the actual impact values (rather than split by quantile groups).

## Setup

Load libraries

```{r setup, include=FALSE}

library(tidyverse)
library(here)
library(rnaturalearth)
library(strex)
library(sf)
library(patchwork)
library(cowplot)
library(glue)
library(RColorBrewer)
library(ggpubr)
library(smoothr)
library(ggtext)
library(ggnewscale)
library(qs)
library(showtext)
showtext_auto()

source(here("src/directories.R"))

source(here("src/spatial.R"))

setwd(dirname(rstudioapi::getSourceEditorContext()$path)) # set working directory to where this script is located
this_dir <- getwd()

biodiv_dir <- file.path(rdsi_dir, "biodiversity_impacts")

impact_maps_dir <- file.path(biodiv_dir, "output")

globe_bbox <- rbind(c(-180, -90), c(-180, 90), 
                      c(180, 90), c(180, -90), c(-180, -90)) 

globe_border <- st_polygon(list(globe_bbox)) %>%
  st_sfc(crs = 4326) %>%
  st_sf(data.frame(rgn = 'globe', geom = .)) %>%
  smoothr::densify(max_distance = 0.5) %>%
  st_transform(crs = crs(moll_template))


```

```{r establish color palettes and themes}

## establish color palette from food footprint project
d_brown <- "#515256"
m_brown <- "#B27B54"
l_brown <- "#BC995F"
green <- "#A8B072"
yellow<- "#EFCE71"
l_yellow <- "#F7F6C1"

jv_pal <- c(l_yellow, yellow, green, l_brown, m_brown, d_brown)

continuous_pal <-  colorRampPalette(jv_pal, space="Lab", bias = 3.5)(10000)

light_gray <- "#F8F9FA"
final_palette <- c(light_gray, continuous_pal)
image(volcano, asp=1, col=final_palette)

palette_red <- c(final_palette, "#B90000")
image(volcano, asp=1, col=palette_red)


countries_shp <- ne_countries(scale = 110, returnclass = "sf") |> 
  st_transform(crs(moll_template)) # read in countries shapefile

theme_ohara <- function(base_size = 9) {
  theme(axis.ticks = element_blank(),
        text             = element_text(family = 'Helvetica',
                                        color = 'gray30', size = base_size),
        plot.title       = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
        panel.background = element_blank(),
        legend.position  = 'right',
        panel.border     = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(colour = 'grey90', size = .25),
        legend.key       = element_rect(colour = NA, fill = NA),
        axis.line        = element_blank())
}


```

Make map split by quantiles in each cell
 - regular quantiles up to 95th 
 - Then 95-99th
 - then >= 99th
 
Additionally, make a delta map, and color either fish dominant or plant dominant. Have the legend colors be the same as the green and blue we use for the boxplot, but split by quantile of delta (same as regular impact quantiles above).
 
 - For main text
 
```{r}

# make publication ready mean extinction risk plots with all feed scenarios, economic allocation
 filepath <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/economic_rarity_mean_prop.qs"))

quantile_pal <- c(
    "0-24" = "#D4F7F0", 
    "25-49" = "#538AC6", 
    "50-74" = "#1E438F", 
    "75-94" = "#F8D422", 
    "95-99" = "#E95D08", 
    "≥99" = "#781211"
)

all_df <-  qread(filepath) %>%
  filter(fcr_type == "regular") %>%
    filter(!is.na(value)) %>%
    mutate(diet_cap = str_to_sentence(diet))

q25 <- quantile(all_df$value, 0.25)
q50 <- quantile(all_df$value, 0.50)
q75 <- quantile(all_df$value, 0.75)
q95 <- quantile(all_df$value, 0.95)
q99 <- quantile(all_df$value, 0.99)


quantile_labels <- formatC(c(q25, q50, q75, q95, q99), format = "e", digits = 0)

formatC(max(all_df$value), digits = 1, format = "e")

all_df_quant <- all_df %>%
    mutate(category = case_when(
    value < q25 ~ "0-24",
    value >= q25 & value < q50 ~ "25-49",
    value >= q50 & value < q75 ~ "50-74",
    value >= q75 & value < q95 ~ "75-94",
    value >= q95 & value < q99 ~ "95-99",
    value >= q99 ~ "≥99"
  )) 

land_eez_rgns <- read.csv(file.path(biodiv_dir, "spatial/rgns_xy_all.csv"))

large_impact_cells <- all_df_quant %>%
  filter(category %in% c("95-99", "≥99")) %>%
  left_join(moll_template_xy, by = c("long" = "x", "lat" = "y")) %>%
  left_join(land_eez_rgns, by = c("long" = "x", "lat" = "y")) %>%
  filter(iso3c %in% c("NOR", "CAN", "ARG", "CHL", "AUS")) %>%
  distinct(cell_id) %>%
  pull() # ok down to 36k cells maybe thats small enough

saveRDS(large_impact_cells, here("analysis/int/large_impact_cells_rarity.rds"))


all_df_quant$category <- factor(all_df_quant$category, levels = c("0-24", "25-49", "50-74", "75-94", "95-99", "≥99"))

quantile_plot <-  ggplot() + 
      geom_tile(data = all_df_quant, aes(x = long, y = lat, fill = category, color = category)) +
                 geom_sf(data = countries_shp, fill = NA, colour = "grey", alpha = 0.2)  +
                 geom_sf(data = globe_border, fill = NA, colour = "grey") + 
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_minimal() + 
  theme(
        axis.title.y = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.margin = margin(b = 0, t = 0), 
        legend.position = "bottom",
        text = element_text(family = "sans", colour = "black"),
        legend.key.spacing.x = unit(-0.1, "pt"),
        strip.text.x = element_text(size = 22, family = "sans", colour = "black"),
        legend.key.width = unit(26, "mm"),
        legend.key.height = unit(0.3, "cm"),
        legend.box = "vertical",  # Stack legends vertically
        legend.box.just = "top",
        legend.spacing.y = unit(-1, "pt"),
        legend.text = element_text(margin = margin(t = 6))
      ) +
  guides(
    color = "none",
    fill = guide_legend(
      nrow = 1, 
      byrow = TRUE,
      label = FALSE
    )
    ) + 
    facet_wrap(~diet_cap) + 
    scale_fill_manual(values = quantile_pal) +
  scale_colour_manual(values = quantile_pal) + 
        labs(color = "",
             fill = "",
             x = "")
  
   quantile_prop_si <- ggdraw(quantile_plot) +
     draw_label("Average rarity weighted proportion of habitat impacted", y = 0.3, x = 0.50, size = 22, fontfamily = "sans") +
     draw_label("Quantiles", y = 0.36, x = 0.50, size = 22, fontfamily = "sans") + 
     draw_label(expression('0–' * 7.6 %*% 10^-18), y = 0.32, x = 0.185, size = 22) +
     draw_label(expression(8 %*% 10^-18 * "–" * 6 %*% 10^-16), y = 0.32, x = 0.31, size = 22) +
     draw_label(expression(6 %*% 10^-16 * "–" * 2 %*% 10^-13), y = 0.32, x = 0.44, size = 22) +
     draw_label(expression(2 %*% 10^-13 * "–" * 5 %*% 10^-11), y = 0.32, x = 0.57, size = 22) +
     draw_label(expression(5 %*% 10^-11 * "–" * 6 %*% 10^-10), y = 0.32, x = 0.7, size = 22) +
     draw_label(expression(6 %*% 10^-10 * "–" * 1 %*% 10^-5), y = 0.32, x = 0.83, size = 22) +
     draw_label("0–24", y = 0.35, x = 0.185, size = 22) +
     draw_label("25–49", y = 0.35, x = 0.31, size = 22) +
     draw_label("50–74", y = 0.35, x = 0.44, size = 22) +
     draw_label("75–94", y = 0.35, x = 0.57, size = 22) +
     draw_label("95–99", y = 0.35, x = 0.7, size = 22) +
     draw_label("≥99", y = 0.35, x = 0.83, size = 22) 
   
  ggsave(glue(here(this_dir, "SI_plots/extra_plots/economic_rarity_mean_prop_quantiles.png")), plot = quantile_prop_si, width = 8, height = 8, bg = "white")
    
    
## Now make the delta plot 
filepath <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/delta/economic_rarity_mean_prop.qs"))

delta_df <-  qread(filepath) %>%
  filter(fcr_type == "regular", 
         delta_cat != "No difference") # there are only a few pixels where there is no difference... so just filter out.

delta_df_quant <- delta_df %>%
  mutate(delta_abs = abs(delta)) %>%
  mutate(q25 = quantile(delta_abs, 0.25),
            q50 = quantile(delta_abs, 0.50),
            q75 = quantile(delta_abs, 0.75),
            q95 = quantile(delta_abs, 0.95),
            q99 = quantile(delta_abs, 0.99)) %>%
  ungroup() %>%
  mutate(category = case_when(
    abs(delta) < abs(q25) ~ "0-24",
    abs(delta) >= abs(q25) & abs(delta) < abs(q50) ~ "25-49",
    abs(delta) >= abs(q50) & abs(delta) < abs(q75) ~ "50-74",
    abs(delta) >= abs(q75) & abs(delta) < abs(q95) ~ "75-94",
    abs(delta) >= abs(q95) & abs(delta) < abs(q99) ~ "95-99",
    abs(delta) >= abs(q99) ~ "≥99"
  ))

# establish palettes for delta map 
plant_pal <- c(
      "0-24" = "#DBF1D5", 
    "25-49" = "#ADDEA7", 
    "50-74" = "#74C476", 
    "75-94" = "#37A055", 
    "95-99" = "#0B7734", 
    "≥99" = "#00441B"
)

fish_pal <- c(
      "0-24" = "#D6E5F4", 
    "25-49" = "#ABCFE5", 
    "50-74" = "#6BAED6", 
    "75-94" = "#3787C0", 
    "95-99" = "#105BA4", 
    "≥99" = "#08306B"
)

delta_df_quant_fis <- delta_df_quant %>% filter(delta_cat == "Fish-dominant")
delta_df_quant_fis$category <- factor(delta_df_quant_fis$category, levels = c("0-24", "25-49", "50-74", "75-94", "95-99", "≥99"))


delta_df_quant_plant <- delta_df_quant %>% filter(delta_cat == "Plant-dominant")
delta_df_quant_plant$category <- factor(delta_df_quant_plant$category, levels = c("0-24", "25-49", "50-74", "75-94", "95-99", "≥99"))


delta_plot <-  ggplot() + 
      geom_tile(data = delta_df_quant_plant, aes(x = long, y = lat, fill = category, color = category)) +
        scale_fill_manual(values = plant_pal) +
    scale_colour_manual(values = plant_pal) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
     theme_ohara() + 
  guides(color = "none") +
  theme(legend.position = "none") + 
     new_scale_color() +
     new_scale_fill() + 
           geom_tile(data = delta_df_quant_fis, aes(x = long, y = lat, fill = category, color = category)) +
             scale_fill_manual(values = fish_pal) +
                 geom_sf(data = countries_shp, fill = NA, colour = "grey", alpha = 0.2)  +
                 geom_sf(data = globe_border, fill = NA, colour = "grey") + 
    scale_colour_manual(values = fish_pal) +
       labs(x = "", y = "") +
  guides(color = "none") + 
     theme(legend.position = "none", 
           axis.text.y = element_blank(), 
           axis.text.x = element_blank(),
           plot.margin = margin(t = 10, r = 10, b = 70, l = 10, unit = "pt"))
   
  fish_legend <- get_legend(ggplot() + 
      geom_tile(data = delta_df_quant_fis, aes(x = long, y = lat, fill = category, color = category)) +
             scale_fill_manual(values = fish_pal) +
              scale_colour_manual(values = fish_pal) +
                  labs(fill = "Fish-dominant") +
 theme(
       legend.margin = margin(b = 0, t = 0), 
        legend.position = "bottom",
        text = element_text(family = "Arial", colour = "black"),
        legend.title = element_text(size = 22, margin = margin(b = 3, t = 3, r = 3)),
        legend.key.width = unit(26, "mm"),
        legend.key.height = unit(0.3, "cm"),
        legend.box = "vertical",  # Stack legends vertically
        legend.box.just = "top",
        legend.spacing.y = unit(0, "pt"),
        legend.text = element_text(margin = margin(t = 3)),
        legend.background = element_rect(fill = "transparent"),
       legend.key.spacing.x = unit(-1, "pt")
      ) +
  guides(
    color = "none",
    fill = guide_legend(
      title.position = "left",  
      title.hjust = 1, 
      title.vjust = -0.2,
      nrow = 1, 
      label.position = "top",  
      label.hjust = 0.5,
      label.vjust = 0.5,
     label.theme = element_text(size = 20),
      byrow = TRUE
    )
    ) 
   )
   
   quant_val_labels <- c(
     unique(delta_df_quant$q25),
     unique(delta_df_quant$q50),
      unique(delta_df_quant$q75),
      unique(delta_df_quant$q95),
      (unique(delta_df_quant$q99)),
            (max(delta_df_quant$delta_abs))
   )
   
   quant_val_labels <- formatC(quant_val_labels, digits = 0, format = "e")

   
plant_legend <- get_legend(ggplot() + 
      geom_tile(data = delta_df_quant_plant, aes(x = long, y = lat, fill = category, color = category)) +
             scale_fill_manual(values = plant_pal) +
              scale_colour_manual(values = plant_pal) +
              labs(fill = "Plant-dominant") +
 theme(
        legend.margin = margin(b = 0, t = 0), 
        legend.position = "bottom",
        text = element_text(family = "Arial", colour = "black"),
        legend.title = element_text(size = 22, margin = margin(b = 3, t = 3, r = 3)),
        legend.key.width = unit(26, "mm"),
        legend.key.height = unit(0.3, "cm"),
        legend.box = "vertical",  # Stack legends vertically
        legend.box.just = "top",
        legend.spacing.y = unit(0, "pt"),
        legend.text = element_text(margin = margin(t = 3)),
        legend.background = element_rect(fill = "transparent"),
        legend.key.spacing.x = unit(-1, "pt")
      ) +
  guides(
    color = "none",
    fill = guide_legend(
      title.position = "left",  
      title.hjust = 1, 
      title.vjust = 1,
      nrow = 1, 
      byrow = TRUE,
     label = FALSE
    )
    ) 
   )
     
   quantile_labels # these are hte ranges for the regular quantiles
   
  quantile_prop_ms <- ggdraw(quantile_plot) +
     draw_label("Average rarity weighted proportion of habitat impacted", y = 0.07, x = 0.50, size = 22, fontfamily = "sans") +
     draw_label("Quantiles", y = 0.21, x = 0.50, size = 22, fontfamily = "sans") + 
     draw_label(expression('0 – ' * 8 %*% 10^-18), y = 0.11, x = 0.17, size = 22) +
     draw_label(expression(8 %*% 10^-18 * " – " * 6 %*% 10^-16), y = 0.11, x = 0.295, size = 22) +
     draw_label(expression(6 %*% 10^-16 * " – " * 2 %*% 10^-13), y = 0.11, x = 0.43, size = 22) +
     draw_label(expression(2 %*% 10^-13 * " – " * 5 %*% 10^-11), y = 0.11, x = 0.57, size = 22) +
     draw_label(expression(5 %*% 10^-11 * " – " * 6 %*% 10^-10), y = 0.11, x = 0.71, size = 22) +
     draw_label(expression(6 %*% 10^-10 * " – " * 1 %*% 10^-5), y = 0.11, x = 0.85, size = 22) +
     draw_label("0–24", y = 0.19, x = 0.175, size = 21) +
     draw_label("25–49", y = 0.19, x = 0.3, size = 21) +
     draw_label("50–74", y = 0.19, x = 0.43, size = 21) +
     draw_label("75–94", y = 0.19, x = 0.57, size = 21) +
     draw_label("95–99", y = 0.19, x = 0.71, size = 21) +
     draw_label("≥99", y = 0.19, x = 0.84, size = 21) 
   
   quant_val_labels # these are the ranges for the delta
   
      full_plot_ms <- ggdraw(delta_plot) +
                draw_plot(plant_legend, height = 0.235, width = 0.86) +
                        draw_plot(fish_legend, height = 0.375, width = 0.868) +
     draw_label("Difference of average rarity weighted proportion of habitat impacted", y = 0.05, x = 0.50, size = 22, fontfamily = "Arial") +
     draw_label("Quantiles", y = 0.22, x = 0.50, size = 22, fontfamily = "Arial") +
     draw_label(expression('0 – ' * 9 %*% 10^-18), y = 0.09, x = 0.17, size = 22) +
     draw_label(expression(9 %*% 10^-18 * " – " * 7 %*% 10^-16), y = 0.09, x = 0.3, size = 22) +
     draw_label(expression(7 %*% 10^-16 * " – " * 3 %*% 10^-13), y = 0.09, x = 0.43, size = 22) +
     draw_label(expression(3 %*% 10^-13 * " – " * 4 %*% 10^-11), y = 0.09, x = 0.57, size = 22) +
     draw_label(expression(4 %*% 10^-11 * " – " *5 %*% 10^-10), y = 0.09, x = 0.705, size = 22) +
     draw_label(expression(5 %*% 10^-10 * " – " * 9 %*% 10^-6), y = 0.09, x = 0.84, size = 22)

  
      
   ## saving two separately because I can
   
        ggsave(glue(here(this_dir, "SI_plots/SI_Fig_3_economic_rarity_mean_prop_quantiles_delta.png")),
                plot = plot_grid(quantile_prop_ms, full_plot_ms, nrow = 2, labels = c("A", "B"), label_fontfamily = "Arial", label_fontface = "bold", label_size = 22), 
                width = 190, height = 180, bg = "white", units = "mm")
          
        
           full_plot_si <- ggdraw(delta_plot) +
                draw_plot(plant_legend, height = 0.3, width = 0.9) +
                        draw_plot(fish_legend, height = 0.375, width = 0.908) +
     draw_label("Difference of average rarity weighted proportion of habitat impacted", y = 0.12, x = 0.50, size = 22, fontfamily = "Arial") +
     draw_label("Quantiles", y = 0.22, x = 0.50, size = 22, fontfamily = "Arial") +
     draw_label(expression('0 – ' * 9 %*% 10^-18), y = 0.14, x = 0.17, size = 22) +
     draw_label(expression(9 %*% 10^-18 * " – " * 7 %*% 10^-16), y = 0.14, x = 0.3, size = 22) +
     draw_label(expression(7 %*% 10^-16 * " – " * 3 %*% 10^-13), y = 0.14, x = 0.43, size = 22) +
     draw_label(expression(3 %*% 10^-13 * " – " * 4 %*% 10^-11), y = 0.14, x = 0.57, size = 22) +
     draw_label(expression(4 %*% 10^-11 * " – " *5 %*% 10^-10), y = 0.14, x = 0.705, size = 22) +
     draw_label(expression(5 %*% 10^-10 * " – " * 9 %*% 10^-6), y = 0.14, x = 0.84, size = 22)
      
    ggsave(glue(here(this_dir, "SI_plots/extra_plots/economic_rarity_mean_prop_delta_plot.png")),
                plot = full_plot_si, 
                width = 180, height = 180, bg = "white", units = "mm")
        
```


Make map showing actual values
 - SI
 
```{r}
allocation_types <- c("mass", "economic", "ge")

 for(allocation in allocation_types){
   #allocation = "mass"
   # make publication ready mean extinction risk plots with all feed scenarios, economic allocation
 filepath <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/{allocation}_rarity_mean_prop.qs"))

all_df <-  qread(filepath) %>%
  filter(fcr_type == "regular")

 plot_list <- list()
  
 diets <- unique(all_df$diet)
 
 for(diet_type in diets){

      # diet_type = "plant-dominant"
      
  loop_loop_df <- all_df %>%
    filter(diet == diet_type)
  
quantile_value <- quantile(loop_loop_df$value, probs = 0.99, na.rm = TRUE)
  
  loop_loop_df <- loop_loop_df %>% 
    mutate(value_new = ifelse(value > quantile_value, quantile_value, value))
  
  loop_plot <-  ggplot() + 
      geom_tile(data = loop_loop_df, aes(x = long, y = lat, fill = value_new, color = value_new)) +
     geom_tile(fill = ifelse(loop_loop_df$value_new < quantile_value, "#B90000", NA)) + 
            geom_sf(data = countries_shp, fill = NA, colour = "grey") + 
       scale_fill_gradientn(colors = palette_red, na.value = "white", labels = scales::label_scientific()) +
            scale_color_gradientn(colors = palette_red, na.value = "white") +
            geom_sf(data = globe_border,
            fill = NA, color = 'grey70', 
            size = .1) +
    scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0)) + 
    theme_void() +
      theme(plot.title = element_text(hjust = 0.5), 
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.margin = margin(r = 10)) + # add a margin to legend text so it isn't cut off when saving 
  labs(fill = "Average rarity\nweighted proportion of\nhabitat\nimpacted") + 
     guides(color = "none") + 
    facet_wrap(~diet)
  
        
      plot_list[[length(plot_list) + 1]] <- loop_plot

 }
 
ggsave(glue(here(this_dir, "SI_plots/extra_plots/{allocation}_rarity_mean_prop.png")), plot = plot_grid(plotlist = plot_list, nrow = 1), width = 10, height = 4, bg = "white")
 }

```
