---
title: "Summary plots: maps"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Summary

The code here produces Figure 1:  
> (A) Map of mean proportion of habitat area impacted. Impacts are split by quantile group and aquafeed formulation. (B) Map of differences in mean impact (%) between the feed formulation scenarios. The magnitude of the difference is split into quantile groups, where darker shades indicate larger differences.

Additionally, we save some supplementary information plots of the values in histograms and maps of the actual impact values (rather than split by quantile groups).

## Setup

Load libraries

```{r setup, include=FALSE}

library(tidyverse)
library(here)
library(rnaturalearth)
library(strex)
library(sf)
library(patchwork)
library(cowplot)
library(glue)
library(RColorBrewer)
library(ggpubr)
library(smoothr)
library(ggtext)
library(ggnewscale)
library(qs)
library(showtext)
showtext_auto()

source(here("src/directories.R"))

source(here("src/spatial.R"))

setwd(dirname(rstudioapi::getSourceEditorContext()$path)) # set working directory to where this script is located
this_dir <- getwd()

biodiv_dir <- file.path(rdsi_dir, "biodiversity_impacts")

impact_maps_dir <- file.path(biodiv_dir, "output")

globe_bbox <- rbind(c(-180, -90), c(-180, 90), 
                      c(180, 90), c(180, -90), c(-180, -90)) 

globe_border <- st_polygon(list(globe_bbox)) %>%
  st_sfc(crs = 4326) %>%
  st_sf(data.frame(rgn = 'globe', geom = .)) %>%
  smoothr::densify(max_distance = 0.5) %>%
  st_transform(crs = crs(moll_template))


```

```{r establish color palettes and themes}
## establish color palette from food footprint project
d_brown <- "#515256"
m_brown <- "#B27B54"
l_brown <- "#BC995F"
green <- "#A8B072"
yellow<- "#EFCE71"
l_yellow <- "#F7F6C1"

jv_pal <- c(l_yellow, yellow, green, l_brown, m_brown, d_brown)

continuous_pal <-  colorRampPalette(jv_pal, space="Lab", bias = 3.5)(10000)
light_gray <- "#F8F9FA"
final_palette <- c(light_gray, continuous_pal)
palette_red <- c(final_palette, "#B90000")

countries_shp <- ne_countries(scale = 110, returnclass = "sf") |> 
  st_transform(crs(moll_template)) # read in countries shapefile

theme_ohara <- function(base_size = 9) {
  theme(axis.ticks = element_blank(),
        text             = element_text(family = 'Helvetica',
                                        color = 'gray30', size = base_size),
        plot.title       = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
        panel.background = element_blank(),
        legend.position  = 'right',
        panel.border     = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(colour = 'grey90', size = .25),
        legend.key       = element_rect(colour = NA, fill = NA),
        axis.line        = element_blank())
}


```

Make map split by quantiles in each cell
 - regular quantiles up to 95th 
 - Then 95-99th
 - then >= 99th
 
Additionally, make a delta map, and color either fish dominant or plant dominant. Have the legend colors be the same as the green and blue we use for the boxplot, but split by quantile of delta (same as regular impact quantiles above).
 
 - For main text
 
```{r}
# make publication ready mean extinction risk plots with all feed scenarios, economic allocation
 filepath <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/economic_mean_prop.qs"))

quantile_pal <- c(
    "0-24" = "#D4F7F0", 
    "25-49" = "#538AC6", 
    "50-74" = "#1E438F", 
    "75-94" = "#F8D422", 
    "95-99" = "#E95D08", 
    "≥99" = "#781211"
)

all_df <-  qread(filepath) %>%
  filter(fcr_type == "regular") %>%
    filter(!is.na(value)) %>%
    mutate(diet_cap = str_to_sentence(diet))

q25 <- quantile(all_df$value, 0.25)
q50 <- quantile(all_df$value, 0.50)
q75 <- quantile(all_df$value, 0.75)
q95 <- quantile(all_df$value, 0.95)
q99 <- quantile(all_df$value, 0.99)


# quantile_labels <- c(
#   trimws(as.character(round(q25[[1]], 10)), "both", "0"),
#   trimws(as.character(round(q50[[1]], 8)), "both", "0"),
#   trimws(as.character(round(q75[[1]], 7)), "both", "0"),
#   trimws(as.character(round(q95[[1]], 5)), "both", "0"),
#   trimws(as.character(round(q99[[1]], 4)), "both", "0"),
#   trimws(as.character(round(max(all_df$value), 2)), "both", "0")
# )
quantile_labels <- formatC(c(q25, q50, q75, q95, q99), format = "e", digits = 0)
formatC(max(all_df$value), digits = 1, format = "e")

all_df_quant <- all_df %>%
    mutate(category = case_when(
    value < q25 ~ "0-24",
    value >= q25 & value < q50 ~ "25-49",
    value >= q50 & value < q75 ~ "50-74",
    value >= q75 & value < q95 ~ "75-94",
    value >= q95 & value < q99 ~ "95-99",
    value >= q99 ~ "≥99"
  )) 

# %>%
#   mutate(new_category = case_when(
#       category == "0-24" ~ "0-\n0.00000004",
#       category == "25-49" ~ "0.00000004-\n0.000002",
#       category == "50-74" ~ "0.000001-\n0.00004",
#       category == "75-94" ~ "0.00004-\n0.006",
#       category == "95-99" ~ "0.006-\n0.06", 
#       category == "≥99" ~ "0.06-\n11"
#     ))

land_eez_rgns <- read.csv(file.path(biodiv_dir, "spatial/rgns_xy_all.csv"))

large_impact_cells <- all_df_quant %>%
  filter(category %in% c("95-99", "≥99")) %>%
  left_join(moll_template_xy, by = c("long" = "x", "lat" = "y")) %>%
  left_join(land_eez_rgns, by = c("long" = "x", "lat" = "y")) %>%
  filter(iso3c %in% c("NOR", "CAN", "ARG", "CHL", "AUS")) %>%
  distinct(cell_id) %>%
  pull() # ok down to 40k cells maybe thats small enough

saveRDS(large_impact_cells, here("analysis/int/large_impact_cells.rds"))

#   new_pal <- c(
#     "0-\n0.00000004" = "#D4F7F0", 
#     "0.00000004-\n0.000002" = "#538AC6", 
#     "0.000001-\n0.00004" = "#1E438F", 
#     "0.00004-\n0.006" = "#F8D422", 
#     "0.006-\n0.06" = "#E95D08", 
#     "0.06-\n11" = "#781211"
# )

all_df_quant$category <- factor(all_df_quant$category, levels = c("0-24", "25-49", "50-74", "75-94", "95-99", "≥99"))
# all_df_quant$new_category <- factor(all_df_quant$new_category, levels = c("0-\n0.00000004", "0.00000004-\n0.000002", "0.000001-\n0.00004", "0.00004-\n0.006", "0.006-\n0.06", "0.06-\n11"))

  quantile_plot <-  ggplot() + 
      geom_tile(data = all_df_quant, aes(x = long, y = lat, fill = category, color = category)) +
                 geom_sf(data = countries_shp, fill = NA, colour = "grey", alpha = 0.2)  +
                 geom_sf(data = globe_border, fill = NA, colour = "grey") + 
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_minimal() + 
  theme(
        axis.title.y = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.margin = margin(b = 0, t = 0), 
        legend.position = "bottom",
        text = element_text(family = "sans", colour = "black"),
        legend.title = element_text(size = 22, margin = margin(b = 3, t = 3)),
        legend.key.spacing.x = unit(-0.1, "pt"),
        strip.text.x = element_text(size = 22, family = "sans", colour = "black"),
        legend.key.width = unit(26, "mm"),
        legend.key.height = unit(0.3, "cm"),
        legend.box = "vertical",  # Stack legends vertically
        legend.box.just = "top",
        legend.spacing.y = unit(-1, "pt"),
        legend.text = element_text(margin = margin(t = 6))
      ) +
  guides(
    color = "none",
    fill = guide_legend(
      nrow = 1, 
      byrow = TRUE,
      label = FALSE
    )
    ) + 
    facet_wrap(~diet_cap) + 
    scale_fill_manual(values = quantile_pal) +
    scale_colour_manual(values = quantile_pal) + 
        labs(color = "",
             fill = "",
             x = "")
  
   quantile_prop_si <- ggdraw(quantile_plot) +
     draw_label("Average proportion of habitat impacted", y = 0.3, x = 0.50, size = 22, fontfamily = "sans") +
     draw_label("Quantiles", y = 0.36, x = 0.50, size =22, fontfamily = "sans") + 
     draw_label(expression('0 – ' * 4 %*% 10^-10), y = 0.32, x = 0.185, size =22) +
     draw_label(expression(4 %*% 10^-10 * " – " * 2 %*% 10^-8), y = 0.32, x = 0.31, size =22) +
     draw_label(expression(2 %*% 10^-8 * " – " * 4 %*% 10^-7), y = 0.32, x = 0.44, size =22) +
     draw_label(expression(4 %*% 10^-7 * " – " * 6 %*% 10^-5), y = 0.32, x = 0.57, size =22) +
     draw_label(expression(6 %*% 10^-5 * " – " * 6 %*% 10^-4), y = 0.32, x = 0.7, size =22) +
     draw_label(expression(6 %*% 10^-04 * " – " * 0.11), y = 0.32, x = 0.83, size =22) +
     draw_label("0–24", y = 0.35, x = 0.185, size =22) +
     draw_label("25–49", y = 0.35, x = 0.31, size =22) +
     draw_label("50–74", y = 0.35, x = 0.44, size =22) +
     draw_label("75–94", y = 0.35, x = 0.565, size =22) +
     draw_label("95–99", y = 0.35, x = 0.69, size =22) +
     draw_label("≥99", y = 0.35, x = 0.82, size =22) 
   

  ggsave(glue(here(this_dir, "SI_plots/extra_plots/economic_mean_prop_quantiles.png")), plot = quantile_prop_si, width = 8, height = 8, bg = "white")
    
    
## Delta plot 
filepath <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/delta/economic_mean_prop.qs"))

delta_df <-  qread(filepath) %>%
  filter(fcr_type == "regular", 
         delta_cat != "No difference") # there are only a few pixels where there is no difference... so just filter out.

delta_df_quant <- delta_df %>%
  mutate(delta_abs = abs(delta)) %>%
  mutate(q25 = quantile(delta_abs, 0.25),
            q50 = quantile(delta_abs, 0.50),
            q75 = quantile(delta_abs, 0.75),
            q95 = quantile(delta_abs, 0.95),
            q99 = quantile(delta_abs, 0.99)) %>%
  ungroup() %>%
  mutate(category = case_when(
    abs(delta) < abs(q25) ~ "0-24",
    abs(delta) >= abs(q25) & abs(delta) < abs(q50) ~ "25-49",
    abs(delta) >= abs(q50) & abs(delta) < abs(q75) ~ "50-74",
    abs(delta) >= abs(q75) & abs(delta) < abs(q95) ~ "75-94",
    abs(delta) >= abs(q95) & abs(delta) < abs(q99) ~ "95-99",
    abs(delta) >= abs(q99) ~ "≥99"
  ))

# establish palettes for delta map 
plant_pal <- c(
      "0-24" = "#DBF1D5", 
    "25-49" = "#ADDEA7", 
    "50-74" = "#74C476", 
    "75-94" = "#37A055", 
    "95-99" = "#0B7734", 
    "≥99" = "#00441B"
)

fish_pal <- c(
      "0-24" = "#D6E5F4", 
    "25-49" = "#ABCFE5", 
    "50-74" = "#6BAED6", 
    "75-94" = "#3787C0", 
    "95-99" = "#105BA4", 
    "≥99" = "#08306B"
)

delta_df_quant_fis <- delta_df_quant %>% filter(delta_cat == "Fish-dominant")
delta_df_quant_fis$category <- factor(delta_df_quant_fis$category, levels = c("0-24", "25-49", "50-74", "75-94", "95-99", "≥99"))


delta_df_quant_plant <- delta_df_quant %>% filter(delta_cat == "Plant-dominant")
delta_df_quant_plant$category <- factor(delta_df_quant_plant$category, levels = c("0-24", "25-49", "50-74", "75-94", "95-99", "≥99"))


delta_plot <-  ggplot() + 
      geom_tile(data = delta_df_quant_plant, aes(x = long, y = lat, fill = category, color = category)) +
        scale_fill_manual(values = plant_pal) +
    scale_colour_manual(values = plant_pal) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
     theme_ohara() + 
  guides(color = "none") +
          theme(legend.position = "none", 
                text = element_text(family = "Arial", colour = "black")) + 
     new_scale_color() +
     new_scale_fill() + 
           geom_tile(data = delta_df_quant_fis, aes(x = long, y = lat, fill = category, color = category)) +
             scale_fill_manual(values = fish_pal) +
                 geom_sf(data = countries_shp, fill = NA, colour = "grey", alpha = 0.2)  +
                 geom_sf(data = globe_border, fill = NA, colour = "grey") + 
    scale_colour_manual(values = fish_pal) +
       labs(x = "", y = "") +
  guides(color = "none") + 
     theme(legend.position = "none", 
           axis.text.y = element_blank(), 
           axis.text.x = element_blank(),
           plot.margin = margin(t = 10, r = 10, b = 70, l = 10, unit = "pt"))


  fish_legend <- get_legend(ggplot() + 
      geom_tile(data = delta_df_quant_fis, aes(x = long, y = lat, fill = category, color = category)) +
             scale_fill_manual(values = fish_pal) +
              scale_colour_manual(values = fish_pal) +
                  labs(fill = "Fish-dominant") +
 theme(
       legend.margin = margin(b = 0, t = 0), 
        legend.position = "bottom",
        text = element_text(family = "Arial", colour = "black"),
        legend.title = element_text(size = 22, margin = margin(b = 3, t = 3, r = 3)),
        legend.key.width = unit(26, "mm"),
        legend.key.height = unit(0.3, "cm"),
        legend.box = "vertical",  # Stack legends vertically
        legend.box.just = "top",
        legend.spacing.y = unit(0, "pt"),
        legend.text = element_text(margin = margin(t = 3)),
        legend.background = element_rect(fill = "transparent"),
       legend.key.spacing.x = unit(-1, "pt")
      ) +
  guides(
    color = "none",
    fill = guide_legend(
      title.position = "left",  
      title.hjust = 1, 
      title.vjust = -0.2,
      nrow = 1, 
      label.position = "top",  
      label.hjust = 0.5,
      label.vjust = 0.2,
     label.theme = element_text(size = 20),
      byrow = TRUE
    )
    ) 
   )
   
   quant_val_labels <- c(
     unique(delta_df_quant$q25),
     unique(delta_df_quant$q50),
      unique(delta_df_quant$q75),
      unique(delta_df_quant$q95),
      (unique(delta_df_quant$q99)),
            (max(delta_df_quant$delta_abs))
   )
   
   quant_val_labels <- formatC(quant_val_labels, digits = 0, format = "e")
   
   
plant_legend <- get_legend(ggplot() + 
      geom_tile(data = delta_df_quant_plant, aes(x = long, y = lat, fill = category, color = category)) +
             scale_fill_manual(values = plant_pal) +
              scale_colour_manual(values = plant_pal) +
              labs(fill = "Plant-dominant") +
 theme(
        legend.margin = margin(b = 0, t = 0), 
        legend.position = "bottom",
        text = element_text(family = "Arial", colour = "black"),
        legend.title = element_text(size = 22, margin = margin(b = 3, t = 3, r = 3)),
        legend.key.width = unit(26, "mm"),
        legend.key.height = unit(0.3, "cm"),
        legend.box = "vertical",  # Stack legends vertically
        legend.box.just = "top",
        legend.spacing.y = unit(0, "pt"),
        legend.text = element_text(margin = margin(t = 3)),
        legend.background = element_rect(fill = "transparent"),
        legend.key.spacing.x = unit(-1, "pt")
      ) +
  guides(
    color = "none",
    fill = guide_legend(
      title.position = "left",  
      title.hjust = 1, 
      title.vjust = 1,
      nrow = 1, 
      byrow = TRUE,
     label = FALSE
    )
    ) 
   )
     

  quantile_prop_ms <- ggdraw(quantile_plot) +
     draw_label("Average proportion of habitat impacted", y = 0.07, x = 0.50, size = 22, fontfamily = "sans") +
     draw_label("Quantiles", y = 0.21, x = 0.50, size = 22, fontfamily = "sans") + 
     draw_label(expression('0 – ' * 4 %*% 10^-10), y = 0.11, x = 0.17, size = 22) +
     draw_label(expression(4 %*% 10^-10 * " – " * 2 %*% 10^-8), y = 0.11, x = 0.295, size = 22) +
     draw_label(expression(2 %*% 10^-8 * " – " * 4 %*% 10^-7), y = 0.11, x = 0.43, size = 22) +
     draw_label(expression(4 %*% 10^-7 * " – " * 6 %*% 10^-5), y = 0.11, x = 0.57, size = 22) +
     draw_label(expression(6 %*% 10^-5 * " – " * 6 %*% 10^-4), y = 0.11, x = 0.71, size = 22) +
     draw_label(expression(6 %*% 10^-04 * " – " * 0.11), y = 0.11, x = 0.85, size = 22) +
     draw_label("0–24", y = 0.19, x = 0.175, size = 21) +
     draw_label("25–49", y = 0.19, x = 0.3, size = 21) +
     draw_label("50–74", y = 0.19, x = 0.43, size = 21) +
     draw_label("75–94", y = 0.19, x = 0.57, size = 21) +
     draw_label("95–99", y = 0.19, x = 0.71, size = 21) +
     draw_label("≥99", y = 0.19, x = 0.84, size = 21) 
   
   quant_val_labels
   full_plot_si <- ggdraw(delta_plot) +
                draw_plot(plant_legend, height = 0.235, width = 0.86) +
                draw_plot(fish_legend, height = 0.375, width = 0.868) +
     draw_label("Difference of average proportion of habitat impacted", y = 0.05, x = 0.50, size = 22, fontfamily = "Arial") +
     draw_label("Quantiles", y = 0.22, x = 0.50, size = 22, fontfamily = "Arial") +
     draw_label(expression('0 – ' * 4 %*% 10^-10), y = 0.09, x = 0.17, size = 22) +
     draw_label(expression(4 %*% 10^-10 * " – " * 2%*% 10^-8), y = 0.09, x = 0.3, size = 22) +
     draw_label(expression(2 %*% 10^-8 * " – " * 6 %*% 10^-7), y = 0.09, x = 0.43, size = 22) +
     draw_label(expression(6 %*% 10^-7 * " – " * 5 %*% 10^-5), y = 0.09, x = 0.57, size = 22) +
     draw_label(expression(5 %*% 10^-5 * " – " * 4 %*% 10^-4), y = 0.09, x = 0.705, size = 22) +
     draw_label(expression(4 %*% 10^-4 * " – " * 0.03), y = 0.09, x = 0.84, size = 22)
   
   
      full_plot_ms <- ggdraw(delta_plot) +
         draw_plot(plant_legend, height = 0.165, width = 0.727) +
         draw_plot(fish_legend, height = 0.38, width = 0.743) +
     draw_label("Difference of average proportion of habitat impacted", y = 0.04, x = 0.50, size = 22, fontfamily = "Arial") +
     draw_label("Quantiles", y = 0.22, x = 0.50, size = 22, fontfamily = "Arial") +
     draw_label(expression('0 – ' * 4 %*% 10^-10), y = 0.08, x = 0.17, size = 22) +
     draw_label(expression(4 %*% 10^-10 * " – " * 2%*% 10^-8), y = 0.08, x = 0.3, size = 22) +
     draw_label(expression(2 %*% 10^-8 * " – " * 6 %*% 10^-7), y = 0.08, x = 0.43, size = 22) +
     draw_label(expression(6 %*% 10^-7 * " – " * 5 %*% 10^-5), y = 0.08, x = 0.57, size = 22) +
     draw_label(expression(5 %*% 10^-5 * " – " * 4 %*% 10^-4), y = 0.08, x = 0.705, size = 22) +
     draw_label(expression(4 %*% 10^-4 * " – " * 0.03), y = 0.08, x = 0.84, size = 22)

   
   ## saving two separately because I can
   
        ggsave(glue(here(this_dir, "MS_plots/fig1_economic_mean_prop_quantiles_delta.png")),
           plot = plot_grid(quantile_prop_ms, full_plot_ms, nrow = 2, labels = c("A", "B"), label_fontfamily = "Arial", label_fontface = "bold", label_size = 22), 
                width = 190, height = 180, bg = "white", units = "mm")
              
        
```

Histogram of impact values by quantile group 
 - SI

```{r}
hist_plot_fish <- ggplot(data = all_df_quant %>% filter(diet == "fish-dominant")) + 
    geom_histogram(aes(value, fill = category)) +
    facet_wrap(~category, scales = "free", ncol = 2) +
    theme_ohara() +
    labs(x = "Average proportion of habitat impacted", y = "Number of cells", title = "Fish-dominant") +
    scale_fill_manual(values = quantile_pal) +
  theme(legend.position = "none")

hist_plot_plant <- ggplot(data = all_df_quant %>% filter(diet == "plant-dominant")) + 
    geom_histogram(aes(value, fill = category)) +
    facet_wrap(~category, scales = "free", ncol = 2) +
    theme_ohara() +
    labs(x = "Average proportion of habitat impacted", y = "Number of cells", title = "Plant-dominant") +
    scale_fill_manual(values = quantile_pal) + 
    theme(legend.position = "none")

  
  ggsave(glue(here(this_dir, "SI_plots/extra_plots/economic_mean_prop_quantiles_hist_plant.png")), plot = plot_grid(hist_plot_plant , nrow = 1), width = 8, height = 8, bg = "white") 
  
    ggsave(glue(here(this_dir, "SI_plots/extra_plots/economic_mean_prop_quantiles_hist_fish.png")), plot = plot_grid(hist_plot_fish, nrow = 1), width = 8, height = 8, bg = "white") 
    
```


Make map showing actual values
 - SI
 
```{r}
# make publication ready mean extinction risk plots with all feed scenarios, economic allocation
 filepath <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/economic_mean_prop.qs"))

all_df <-  qread(filepath) %>%
  filter(fcr_type == "regular")

 plot_list <- list()
  
 diets <- unique(all_df$diet)
 
 for(diet_type in diets){

      # diet_type = "plant-dominant"
      
  loop_loop_df <- all_df %>%
    filter(diet == diet_type)
  
quantile_value <- quantile(loop_loop_df$value, probs = 0.99, na.rm = TRUE)
  
  loop_loop_df <- loop_loop_df %>% 
    mutate(value_new = ifelse(value > quantile_value, quantile_value, value))
  
  loop_plot <-  ggplot() + 
      geom_tile(data = loop_loop_df, aes(x = long, y = lat, fill = value_new, color = value_new)) +
     geom_tile(fill = ifelse(loop_loop_df$value_new < quantile_value, "#B90000", NA)) + 
            geom_sf(data = countries_shp, fill = NA, colour = "grey") + 
       scale_fill_gradientn(colors = palette_red, na.value = "white") +
    # scale_fill_gradientn(breaks = c(round(quantile_value[[1]]/2, 4), round(quantile_value[[1]], 4), max(loop_loop_df$value)), colors = palette_red) + 
            scale_color_gradientn(colors = palette_red, na.value = "white") +
            geom_sf(data = globe_border,
            fill = NA, color = 'grey70', 
            size = .1) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    theme_void() +
      theme(plot.title = element_text(hjust = 0.5), 
          axis.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.margin = margin(r = 10)) + # add a margin to legend text so it isn't cut off when saving 
  labs(fill = "Proportion\nhabitat\nimpacted") + 
     guides(color = "none") + 
    facet_wrap(~diet)
  
        
      plot_list[[length(plot_list) + 1]] <- loop_plot

  } 

ggsave(glue(here(this_dir, "SI_plots/extra_plots/economic_mean_prop.png")), plot = plot_grid(plotlist = plot_list, nrow = 1), width = 10, height = 4, bg = "white")
```


Map stats!


Compare diet types and ncells impacted

```{r}
test <- all_df %>%
  group_by(diet, fcr_type) %>%
  summarise(prop_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

scenario_summary <- all_df %>% 
  filter(value > 0) %>%
  group_by(diet, fcr_type) %>%
  summarise(cells_impacted = n()) %>%
  ungroup()

moll_land <- readRDS(here("prep/03_prep_spp_habitats/data/spatial/moll_template_land_xy.rds"))

all_df_land <- all_df %>%
  left_join(moll_land, by = c("long" = "x", "lat" = "y")) %>%
  filter(!is.na(cell_id)) %>%
  filter(value > 0) %>%
    group_by(diet, fcr_type) %>%
  summarise(cells_impacted = n()) %>%
  ungroup() %>%
   mutate(land_cells_total = 2876998) %>%
  mutate(prop = cells_impacted/land_cells_total)

moll_ocean <- readRDS(here("prep/03_prep_spp_habitats/data/spatial/moll_template_ocean_xy.rds"))

all_df_ocean <- all_df %>%
  left_join(moll_ocean, by = c("long" = "x", "lat" = "y")) %>%
  filter(!is.na(cell_id)) %>%
    filter(value > 0) %>%
    group_by(diet, fcr_type) %>%
  summarise(cells_impacted = n()) %>%
  ungroup() %>%
  mutate(ocean_cells_total = 3684240) %>%
  mutate(prop = cells_impacted/ocean_cells_total)

# Comparing the fish-dominant and plant-dominant feed scenarios, under the regular FCR scenario, we see that the plant-dominant diet has 81,352 more pixels impacted (there are ~6.5 million pixels on a 10km by 10 km Mollweide grid). This difference is almost exclusively due to the difference in terrestrial crop impacts, with the plant-dominant diet having 80,789 more terrestrial pixels impacted than that of the fish-dominant diet. Overall, ocean pixels are impacted over a larger area: ~42% of ocean pixels are impacted, regardless of feed scenario. In comparison, 23%-26% of land pixels are impacted, depending on the scenario. 

```




