---
title: "Publication plots: non maps"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Summary
We are going to make a duplicate figure 3 using mass and energetic allocation. This is the best way to display the differences between allocation approaches and the impacts raw materials are responsible for. 

> Figure 3. Proportion of each taxa’s global habitat area which is impacted by raw material for each aquafeed scenario.


> SI Figure 10. Proportion of each taxa’s global habitat area which is impacted by raw material for each aquafeed scenario, using mass allocation. Related to Figure 3. 

> SI Figure 11. Proportion of each taxa’s global habitat area which is impacted by raw material for each aquafeed scenario, using energetic allocation. Related to Figure 3.


## Setup

Load libraries

```{r setup, include=FALSE}

library(tidyverse)
library(here)
library(data.table)
library(terra)
library(strex)
library(patchwork)
library(cowplot)
library(glue)
library(RColorBrewer)
library(ggpubr)
library(qs)
library(parallel)
library(doParallel)

source(here("src/directories.R"))

source(here("src/spatial.R"))

biodiv_dir <- file.path(rdsi_dir, "biodiversity_impacts")

impact_maps_dir <- file.path(biodiv_dir, "output")

this_dir <- here("analysis/")

```

```{r}
# Create a named vector with colors for each spp_type
color_scale <- readRDS(here("prep/03_prep_spp_habitats/int/color_scale.RDS"))

pal_mats <- c("Soybean" = "#328C41",
              "Trimmings fish" = "#8FBAED",
              "Sunflower" = "#4277A9",
              "Wheat" = "#FFE96D",
              "Rapeseed" = "#AC9900",
              "Pulses" = "#79E0DA",
              "Forage fish" = "#3DB1AB",
              "Maize" = "#0A6B66",
              "Other crops" = "#BC94C6")

theme_ohara <- function(base_size = 9) {
  theme(axis.ticks = element_blank(),
        text             = element_text(family = 'Helvetica',
                                        color = 'gray30', size = base_size),
        plot.title       = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
        panel.background = element_blank(),
        legend.position  = 'right',
        panel.border     = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(colour = 'grey90', size = .25),
        legend.key       = element_rect(colour = NA, fill = NA),
        axis.line        = element_blank())
}

```

Read in marine and terrestrial dfs and combine

```{r}
marine_overlap <- readRDS(here("prep/03_prep_spp_habitats/output/aoh_marine_impact.rds")) %>%
   filter(sensitivity_scenario == "original") %>%
    dplyr::select(sciname, spp_type, diet, fcr_type, allocation, ingredient, vulnerability, impact_total, total_hab_area)

terrestrial_overlap <- readRDS(here("prep/03_prep_spp_habitats/output/aoh_terrestrial_impact.rds")) %>% 
   filter(sensitivity_scenario == "original") %>%
      dplyr::select(sciname, spp_type, diet, fcr_type, allocation, ingredient, vulnerability, impact_total, total_hab_area)

all_overlap <- rbind(marine_overlap, terrestrial_overlap) %>%
    mutate(taxon = str_to_sentence(spp_type)) %>%
  mutate(taxon = case_when(
    taxon == "Terrestrial mammal" ~ "Terrestrial mammals",
    taxon == "Bird" ~ "Birds",
    taxon == "Fish" ~ "Finfish", 
    taxon == "Marine mammal" ~ "Marine mammals",
    taxon == "Marine plant" ~ "Marine plants",
    TRUE ~ taxon)) %>%
  mutate(ingredient = str_replace(ingredient, "CropsNES", "Other crops"))

```

```{r}
taxon_names <- all_overlap %>%
  distinct(spp_type, taxon)

spp_impacted <- all_overlap %>%
  filter(impact_total > 0) %>%
  group_by(taxon, diet, allocation, fcr_type) %>%
  summarise(spp_count_impacted = n_distinct(sciname)) %>%
  ungroup()

aquamaps_spp <- readRDS(file.path(rdsi_dir, "biodiversity_impacts/int/spp_vuln_depth_info.rds")) %>%
  distinct(spp_type = taxon, sciname = species)
  
eyres_spp <- readRDS(here("prep/03_prep_spp_habitats/data/iucn/eyres_iucn_spp.rds")) %>%
  dplyr::select(spp_type, sciname = scientific_name) %>%
  mutate(sciname = tolower(sciname)) %>%
  mutate(spp_type = case_when(
    spp_type == "mammals" ~ "Terrestrial mammal",
    spp_type == "amphibians" ~ "Amphibians", 
    spp_type == "reptiles" ~ "Reptiles", 
    spp_type == "birds" ~ "Bird"
  ))

spp_assessed <- rbind(aquamaps_spp, eyres_spp) %>%
  left_join(taxon_names) %>%
  group_by(taxon) %>%
  summarise(spp_count_assessed = n_distinct(sciname)) %>%
  ungroup() %>%
  filter(!is.na(taxon))

spp_assessed_impacted <- spp_impacted %>%
  left_join(spp_assessed) 

spp_impact_test <- spp_assessed_impacted %>%
  group_by(diet, allocation, fcr_type) %>%
  summarise(total_impact = sum(spp_count_impacted, na.rm = TRUE),
            total_assessed = sum(spp_count_assessed, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(prop = total_impact/total_assessed)

spp_total_hab <- all_overlap %>%
  distinct(sciname, spp_type, total_hab_area) %>%
  mutate(id = row_number())


```
 
```{r}
spp_assessed_plot_df <- spp_assessed %>%
  mutate(spp_count_str = glue("n = {spp_count_assessed}"))

all_overlap_df <- all_overlap %>%
  rename(fcr = fcr_type) %>%
        separate_wider_delim(ingredient, delim = "_", names = c("raw_material", "ingredient")) %>% 
  mutate(raw_material = str_to_sentence(raw_material), 
         ingredient = str_to_sentence(ingredient),
         allocation = str_to_sentence(allocation),
         diet = str_to_sentence(diet),
         fcr = str_to_sentence(fcr)) %>%
    filter(fcr == "Regular")

taxon_hab_area <- spp_total_hab %>%
  dplyr::select(-sciname, -id) %>%
  group_by(spp_type) %>%
  summarise(total_hab_area = sum(total_hab_area, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(taxon_names) %>%
  dplyr::select(-spp_type)

```
 
```{r}
impacted_spp <- all_overlap_df %>%
  filter(impact_total > 0) %>%
  pull(sciname) %>%
  unique()

taxon_hab_area <- spp_total_hab %>%
  filter(sciname %in% impacted_spp) %>%
  dplyr::select(-sciname, -id) %>%
  group_by(spp_type) %>%
  summarise(total_hab_area = sum(total_hab_area, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(taxon_names) %>%
  dplyr::select(-spp_type)

summary_df_barcharts <- all_overlap_df %>%
  group_by(diet, fcr, raw_material, allocation, taxon) %>%
  summarise(impact_total = sum(impact_total, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(impact_total > 0) %>%
  left_join(taxon_hab_area) %>%
  mutate(prop_impact = impact_total/total_hab_area) %>%
    mutate(prop_impact_mirror = ifelse(diet == "Fish-dominant", -prop_impact, prop_impact)) %>%
  mutate(allocation = ifelse(allocation == "Ge", "Energetic", allocation))

## switch the raw materials and species in that plot
allocations <- c("Energetic", "Mass")
for(allocation_type in allocations){
  
  # allocation_type = "Mass"
  
  summary_df_plots <- summary_df_barcharts %>%
    filter(allocation == allocation_type)

# Calculate the *total* mirrored impact per taxon (summed over raw materials)
sum_impacts <- summary_df_plots %>%
  group_by(diet, taxon) %>%
  summarize(total_impact = sum(prop_impact_mirror, na.rm = TRUE), .groups = "drop")

# Find the maximum absolute value
max_val <- max(abs(sum_impacts$total_impact), na.rm = TRUE)

summary_df_p <- summary_df_plots %>% filter(diet == "Plant-dominant")

summary_df_p$taxon <- with(summary_df_p, reorder(taxon, prop_impact, FUN = sum))

# 4. Plot
p_plot <- ggplot(summary_df_p, aes(x = prop_impact_mirror, 
                                 y = taxon, 
                                 fill = raw_material)) +
    geom_bar(stat = "identity") +
    facet_wrap(~diet, scales = "fixed") +  # No "scales = free_x" -- we keep same scale
    scale_fill_manual(values = pal_mats) +
    labs(x = NULL, 
         y = NULL, 
         fill = "Raw material",
         caption = glue("{allocation_type} allocation")) +
    theme_minimal() + 
    theme(
        panel.background = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 8),
        strip.text = element_text(size = 8),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        text = element_text(family = "arial"),
        axis.ticks.y = element_blank(),
        plot.margin = margin(0, 0, 15, 0)  # Remove any plot margins

    ) +
    scale_x_continuous(
        limits = c(0, max_val),  # Fix limits symmetrically
        expand = c(0, 0),
        labels = abs,  # Label all ticks as positive
        breaks = c(0.00001, 0.00002, 0.00003)
    )

summary_df_f <- summary_df_plots %>% filter(diet == "Fish-dominant")

summary_df_f$taxon <- factor(summary_df_f$taxon, levels = levels(summary_df_p$taxon))

# 4. Plot
f_plot <- ggplot(summary_df_f, aes(x = prop_impact_mirror, 
                                 y = taxon, 
                                 fill = raw_material)) +
    geom_bar(stat = "identity") +
    facet_wrap(~diet, scales = "fixed") +  # No "scales = free_x" -- we keep same scale
    scale_fill_manual(values = pal_mats) +
    labs(x = NULL, 
         y = "Taxa", 
         fill = NULL) +
    theme_minimal() + 
    theme(
        panel.background = element_blank(),
        axis.text.x = element_text(size = 8),
        axis.text.y = element_text(size = 8),
        strip.text = element_text(size = 8),
        axis.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        text = element_text(family = "arial"),
        plot.margin = margin(0, 0, 15, 0)  # Remove any plot margins

    ) +
    scale_x_continuous(
        limits = c(-max_val, 0),  # Fix limits symmetrically
        expand = c(0, 0),
        labels = abs,  # Label all ticks as positive
                breaks = c(-0.00001, -0.00002, -0.00003)

    ) + 
  guides(fill = FALSE)

(total_plot <-  f_plot + p_plot)

ggdraw(total_plot) +
       draw_label("0", y = 0.085, x = 0.49, size = 8, fontfamily = "arial", color = "grey30") +
  draw_label("Proportion of global habitat area impacted", y = 0.03, x = 0.5, size = 8, fontfamily = "arial", color = "black") 

if(allocation_type == "Energetic"){
ggsave(here(this_dir, glue("/SI_plots/SI_Fig_11_taxa_raw_mat_{allocation_type}_barchart.png")), width = 178.9, height =100.92 , units = "mm", bg = "white")
  
}else{
  ggsave(here(this_dir, glue("/SI_plots/SI_Fig_10_taxa_raw_mat_{allocation_type}_barchart.png")), width = 178.9, height =100.92 , units = "mm", bg = "white")
}

}

```




