---
title: "Summary plots: uncertainty and variability"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Setup

Load libraries

```{r setup, include=FALSE}

library(tidyverse)
library(here)
library(data.table)
library(terra)
library(rnaturalearth)
library(strex)
library(janitor)
library(tools)
library(khroma)
library(scales)
library(fields)
library(patchwork)
library(cowplot)
library(glue)
library(viridis)
library(RColorBrewer)
library(ggpubr)
library(smoothr)
library(tmap)

source(here("src/directories.R"))

source(here("src/spatial.R"))

select <- dplyr::select
setwd(dirname(rstudioapi::getSourceEditorContext()$path)) # set working directory to where this script is located
this_dir <- getwd()

biodiv_dir <- file.path(rdsi_dir, "biodiversity_impacts")

impact_maps_dir <- file.path(biodiv_dir, "output")

```

```{r}
## load in some data for categories and general exploration
marine_overlap <- readRDS(here("prep/03_prep_spp_habitats/output/aoh_marine_impact.rds")) %>%
    dplyr::select(sciname, spp_type, diet, fcr_type, allocation, ingredient, vulnerability, impact_total, total_hab_area)

terrestrial_overlap <- readRDS(here("prep/03_prep_spp_habitats/output/aoh_terrestrial_impact.rds")) %>% 
  dplyr::select(-suitability) %>%
      dplyr::select(sciname, spp_type, diet, fcr_type, allocation, ingredient, vulnerability, impact_total, total_hab_area)

all_overlap <- rbind(marine_overlap, terrestrial_overlap) %>%
    mutate(taxon = str_to_sentence(spp_type)) %>%
  mutate(taxon = case_when(
    taxon == "Terrestrial mammal" ~ "Terrestrial mammals",
    taxon == "Bird" ~ "Birds",
    taxon == "Fish" ~ "Finfish", 
    taxon == "Marine mammal" ~ "Marine mammals",
    taxon == "Marine plant" ~ "Marine plants",
    TRUE ~ taxon))

```

Lets see if the standard deviation generally scales with the mean; if it does, we should probably report coefficient of variation instead. 
 - these plots take a while to save - like 10 mins each.
```{r}
## read in mean rast for plant-dominant regular FCR

## read in sd rast for plant-dominant regular FCR

## calculate some sort of scaling plot

allocations <- c("ecomomic", "mass", "ge")

for(allocation_type in allocations){
  
# allocation_type = "economic"
  
# Read data
 filepath_mean <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/{allocation_type}_mean_prop.qs"))
 
  filepath_sd <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/{allocation_type}_sd_prop.qs"))

loop_df_mean <-  qs::qread(filepath_mean) %>%
  dplyr::select(-calc_type) %>%
  rename(mean_prop = value)

loop_df_sd <-  qs::qread(filepath_sd) %>%
    dplyr::select(-calc_type) %>%
  rename(sd_prop = value) 

loop_df_all <- loop_df_mean %>%
  left_join(loop_df_sd) %>%
  mutate(cv = sd_prop/mean_prop) %>%
  filter(!is.na(sd_prop))

loop_df_all$fcr_type = factor(loop_df_all$fcr_type, levels=c("regular", "efficient"))


p <- ggplot(data = loop_df_all, aes(x = mean_prop, y = sd_prop)) +
  geom_point() +
  geom_smooth(se = FALSE, formula = y~x) +
  geom_abline() +
  labs(caption = glue("{allocation_type}")) + 
  facet_grid(diet~fcr_type, scales = "free") 

ggsave(filename = glue(here("analysis/uncertainty/sd_explore/mean_sd_scaling_{allocation_type}.png")), plot = p)


}

```


Lets map how each scenario agrees (or not) with one another on high vs low impact areas:
  	  - most impacted cells (>90 quantile?) in 1, 2, 3, or 4 scenarios (different colors); least impacted cells (<25th quantile) in 1, 2, 3, or 4 scenarios (different colors); this will tell us the places that for sure have high or low impacts 
  	  - Fig 1 from [Stock et al.](https://conbio.onlinelibrary.wiley.com/doi/full/10.1111/cobi.13141?saml_referrer)
  	  - Maybe try with mean impacts for each allocation, diet, and fcr? So will be reading in 12 rasters (3 allocations X 2 diets X 2 FCRs)

```{r}

# read in each scenario, filter for >95th quantile, set all cells to 1; filter for <25th quantile, set all cells to 1. Stack all and count what cells match up 

allocation_types <- c("economic", "mass", "ge")
fcr_types <- c("regular", "efficient")
diet_types <- c("plant-dominant", "fish-dominant")



  for(diet_type in diet_types){
    for(fcr_type in fcr_types){
      
high_impact_rast <- moll_template
low_impact_rast <- moll_template

      # diet_type = "fish-dominant"
      # fcr_type = "efficient"

for(allocation_type in allocation_types){
      
      # allocation_type = "economic"
      
      imp_rast <- rast(glue(file.path(biodiv_dir, "output/impact_maps_across_taxon_ingredient/{diet_type}/{fcr_type}/{allocation_type}_mean_prop.tif")))
      
      quant_95 <- global(imp_rast, probs = 0.95, na.rm = TRUE)
      
      imp_rast_95 <- classify(imp_rast, cbind(quant_95$mean, Inf, 1), others = NA, right = FALSE)

      
      quant_25 <-  global(imp_rast, probs = 0.05, na.rm = TRUE)
      
      imp_rast_25 <- classify(imp_rast, cbind(0, quant_25$mean, 1), others = NA, right = TRUE)

      
      high_impact_rast <- c(high_impact_rast, imp_rast_95)
      
      low_impact_rast <- c(low_impact_rast, imp_rast_25)
      
    }
  


high_imp_sum <- sum(high_impact_rast, na.rm = TRUE)
low_imp_sum <- sum(low_impact_rast, na.rm = TRUE)

pal <- colorRampPalette(c("yellow","darkred"))


png(file = glue(here("analysis/uncertainty/{diet_type}_{fcr_type}/high_impact.png")), width = 1200, height = 700)
plot(high_imp_sum, col = pal(3), type = "classes") 
dev.off()

png(file = glue(here("analysis/uncertainty/{diet_type}_{fcr_type}/low_impact.png")), width = 1200, height = 700)
plot(low_imp_sum, col = pal(3), type = "classes") 
dev.off()

## look at scenario disagreement (sometimes scenarios has cells that are classified as both high impact and low impact)
all_imp_sum <- intersect(high_imp_sum, low_imp_sum)

r <- classify(all_imp_sum, cbind(FALSE, 0))

png(file = glue(here("analysis/uncertainty/{diet_type}_{fcr_type}/disagree.png")), width = 1200, height = 700)
plot(r)
dev.off()

    }
  }

```

```{r}
test <- 1 - rast(file.path(biodiv_dir, "output/impact_maps_across_taxon_ingredient/plant-dominant/regular/economic_mean_prop.tif"))
test[test == 0] <- NA
global(test, "mean", na.rm = TRUE) # 0.00001099836
plot(raster(test))
s <- raster::select(raster(test))
plot(s)
where.max(test)

test_rescale <- test/0.02474564

```


