---
title: "Summary plots: uncertainty and variability"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Summary

Here we map how well our maps agree across allocation approaches. 

We split impacts into high (>95th quantile) or low (<25th quantile), and see if each allocation approach (economic, ge, mass) agrees in those categories. So for example, we count how many of the allocation approaches agree in each cell. We also calculate cells in which there is disagreement (i.e., where one approach reports high impact cells but another reports low impact cells).

## Setup

Load libraries

```{r setup, include=FALSE}

library(tidyverse)
library(here)
library(data.table)
library(terra)
library(rnaturalearth)
library(strex)
library(janitor)
library(tools)
library(khroma)
library(scales)
library(fields)
library(patchwork)
library(cowplot)
library(glue)
library(viridis)
library(RColorBrewer)
library(ggpubr)
library(smoothr)
library(tmap)

source(here("src/directories.R"))

source(here("src/spatial.R"))

select <- dplyr::select
setwd(dirname(rstudioapi::getSourceEditorContext()$path)) # set working directory to where this script is located
this_dir <- getwd()

biodiv_dir <- file.path(rdsi_dir, "biodiversity_impacts")

impact_maps_dir <- file.path(biodiv_dir, "output")

```

```{r}
## load in some data for categories and general exploration
marine_overlap <- readRDS(here("prep/03_prep_spp_habitats/output/aoh_marine_impact.rds")) %>%
    dplyr::select(sciname, diet, fcr_type, allocation, ingredient, vulnerability, sensitivity_scenario, impact_total, total_hab_area)

terrestrial_overlap <- readRDS(here("prep/03_prep_spp_habitats/output/aoh_terrestrial_impact.rds")) %>% 
      dplyr::select(sciname, diet, fcr_type, allocation, ingredient, vulnerability, sensitivity_scenario, impact_total, total_hab_area)

all_overlap <- rbind(marine_overlap, terrestrial_overlap) 


test <- terrestrial_overlap %>% 
  filter(allocation == "economic") %>%
  group_by(diet, sensitivity_scenario) %>%
  summarise(total_impact = sum(impact_total, na.rm = TRUE)) %>%
  ungroup()

test <- marine_overlap %>% 
  filter(allocation == "economic") %>%
  group_by(diet, sensitivity_scenario) %>%
  summarise(total_impact = sum(impact_total, na.rm = TRUE)) %>%
  ungroup()

```

Lets take a look at our species sensitivity scenarios: 

Plots to make/things to explore: 
 - Violin or boxplot of total impact per species across sensitivity scenarios 
   - Facet by feed scenario (fish or plant) to show how uncertainty compares across diets 
 - For each species or other grouping
   - Compute difference in mean impact between two feed scenarios under original sensitivity
   - Compute the range (or sd) of impact across ±0.1 sensitivity scenarios 
   - Then make a signal-noise ratio where signal = mean_impact_plant - mean_impact_fish and noise = SD or range across sensitivity scenarios 
 - Are there any species where the rankings of the two feed scenarios flips depending on the sensitivity assumption? For example, does the fish based feed have higher impact under one sensitivity  but lower under another? 


```{r}
# read in data frames 
all_impacts <- all_overlap %>%
  filter(allocation == "economic") %>% # filter for economic
  mutate(diet = ifelse(diet == "plant-dominant", "plant", "fish")) # rename for easier column manipulation

all_impacts_grouped <- all_impacts %>%
  group_by(sciname, diet, sensitivity_scenario) %>%
  summarise(total_impact = sum(impact_total, na.rm = TRUE),
            mean_impact = mean(impact_total, na.rm = TRUE)) %>%
  ungroup()


## lets look at overall trends 
# Filter to economic allocation (or adjust as needed)
impacts_filtered <- all_impacts %>%
  filter(allocation == "economic") %>%
  mutate(sensitivity_scenario = str_remove(sensitivity_scenario, "-"))

# Summarize total and mean impact per diet and sensitivity
impacts_summary <- impacts_filtered %>%
  group_by(diet, sensitivity_scenario) %>%
  summarise(
    total_impact = sum(impact_total, na.rm = TRUE),
    mean_impact = mean(impact_total, na.rm = TRUE),
    .groups = "drop"
  )

# Reshape to wide to compare other scenarios to original
impacts_wide <- impacts_summary %>%
  pivot_wider(
    names_from = sensitivity_scenario,
    values_from = c(total_impact, mean_impact),
    names_glue = "{.value}_{sensitivity_scenario}"
  )

# Compute percent and absolute differences vs original
impacts_change <- impacts_wide %>%
  mutate(
    abs_diff_total_plus10 = total_impact_plus10 - total_impact_original,
    pct_diff_total_plus10 = 100 * abs_diff_total_plus10 / total_impact_original,

    abs_diff_total_minus10 = total_impact_minus10 - total_impact_original,
    pct_diff_total_minus10 = 100 * abs_diff_total_minus10 / total_impact_original,

    abs_diff_mean_plus10 = mean_impact_plus10 - mean_impact_original,
    pct_diff_mean_plus10 = 100 * abs_diff_mean_plus10 / mean_impact_original,

    abs_diff_mean_minus10 = mean_impact_minus10 - mean_impact_original,
    pct_diff_mean_minus10 = 100 * abs_diff_mean_minus10 / mean_impact_original
  ) %>%
  dplyr::select(diet, pct_diff_total_plus10, pct_diff_total_minus10, pct_diff_mean_plus10, pct_diff_mean_minus10)

## lets look at species level trends and see if any changes in imapct are unexpected 
# Filter to economic allocation (or adjust as needed)
impacts_filtered <- all_impacts %>%
  filter(allocation == "economic") %>%
  mutate(sensitivity_scenario = str_remove(sensitivity_scenario, "-")) 
#%>%
 # filter(impact_total > 0)

# Summarize total and mean impact per diet and sensitivity
impacts_summary <- impacts_filtered %>%
  group_by(diet, sciname, sensitivity_scenario) %>%
  summarise(
    total_impact = sum(impact_total, na.rm = TRUE),
    mean_impact = mean(impact_total, na.rm = TRUE),
    .groups = "drop"
  )

# Reshape to wide to compare other scenarios to original
impacts_wide <- impacts_summary %>%
  pivot_wider(
    names_from = sensitivity_scenario,
    values_from = c(total_impact, mean_impact),
    names_glue = "{.value}_{sensitivity_scenario}"
  )

# Compute percent and absolute differences vs original
impacts_change <- impacts_wide %>%
  mutate(
    abs_diff_total_plus10 = total_impact_plus10 - total_impact_original,
    pct_diff_total_plus10 = 100 * abs_diff_total_plus10 / total_impact_original,

    abs_diff_total_minus10 = total_impact_minus10 - total_impact_original,
    pct_diff_total_minus10 = 100 * abs_diff_total_minus10 / total_impact_original,

    abs_diff_mean_plus10 = mean_impact_plus10 - mean_impact_original,
    pct_diff_mean_plus10 = 100 * abs_diff_mean_plus10 / mean_impact_original,

    abs_diff_mean_minus10 = mean_impact_minus10 - mean_impact_original,
    pct_diff_mean_minus10 = 100 * abs_diff_mean_minus10 / mean_impact_original
  ) %>%
  dplyr::select(diet, sciname, pct_diff_total_plus10, pct_diff_total_minus10, pct_diff_mean_plus10, pct_diff_mean_minus10, total_impact_original, mean_impact_original, total_impact_plus10)

impacts_test <- impacts_change %>%
  filter(pct_diff_total_plus10 == "Inf") %>%
  filter(diet == "fish")

impacts_test_2 <- all_impacts %>% 
  filter(vulnerability == 0.1) %>%
  group_by(diet) %>%
  summarise(total_impact = sum(impact_total, na.rm = TRUE)) %>%
  ungroup()

# # A tibble: 2 × 2
#   diet  total_impact
#   <chr>        <dbl>
# 1 fish        31887.
# 2 plant       70914.

impacts_all <- impacts_filtered %>%
  group_by(diet, sensitivity_scenario) %>%
  summarise(
    total_impact = sum(impact_total, na.rm = TRUE),
    mean_impact = mean(impact_total, na.rm = TRUE),
    .groups = "drop"
  )
  
# total original plant diet impacts : 552924.0
70913.39/552924.0 # 0.1282516 - 13%
# total original fish diet impacts: 515164.5
31886.5/515164.5 # 6%

# compared to plus-10 scenario
31887/605338.5 # 5%
70914/650006.6 # 11%


# 6. Optional: identify reversals in scenario ranking across sensitivity
ranking_check <- all_impacts %>%
  filter(allocation == "economic") %>%
  group_by(sciname, diet, sensitivity_scenario) %>%
  summarise(total_impact = sum(impact_total, na.rm = TRUE)) %>%
  pivot_wider(names_from = diet, values_from = total_impact, names_prefix = "diet_") %>%
  mutate(winner = ifelse(diet_fish > diet_plant, "fish", "plant"))  ## need to look at how many of the species have a different "winner" in any of the three scenarios. When there is a different winner, is it always when we added a new species to be sensitive? 

ranking_pivot <- ranking_check %>%
  dplyr::select(sciname, sensitivity_scenario, winner) %>%
  pivot_wider(names_from = sensitivity_scenario, values_from = winner) 

inconsistent_rows <- ranking_pivot[
  !(ranking_pivot$`minus-10` == ranking_pivot$original & 
    ranking_pivot$original == ranking_pivot$`plus-10`), ] %>%
  filter(`minus-10` == original) # ok so for these species 2374 flip to fish domninant having more impacts under the plus 10 scenario. 

2374/54625 # no species flip from original to minus-10. 4% of species (n=2374) flip from plant-dominant to fish-dominant having more impacts going from original to plus-10. What % of impacts do these flipped species represent of the original impacts? 

flipped_spp <- inconsistent_rows %>%
  distinct(sciname) %>%
  pull(sciname)

impacts_test_2 <- all_impacts %>% 
  filter(sciname %in% flipped_spp) %>%
  group_by(diet, sensitivity_scenario) %>%
  summarise(total_impact = sum(impact_total, na.rm = TRUE)) %>%
  ungroup()

impacts_test_spp <- all_impacts %>% 
  filter(sciname %in% flipped_spp) %>%
  distinct(sciname, ingredient) %>%
  mutate(type = ifelse(ingredient %in% c("trimmings fish_fish meal, cut offs", "trimmings fish_fish oil, cut offs", "forage fish_fish meal", "forage fish_fish oil"), "marine", "terrestrial")) %>%
  distinct(sciname, type) # only 23 of the 2374 species which "flipped" the impacts are terrestrial, which explains why the fish-dominant is higher, because more marine species were sensitive and exposed under the new scenario.
 
impacts_test_spp <- all_impacts %>% 
  filter(vulnerability == 0.1) %>%
  pull(sciname) %>%
  unique() # so there were 7534 new species could be impacted

impacts_test_spp <- all_impacts %>% 
  filter(vulnerability == 0.1,
         impact_total > 0) %>%
  pull(sciname) %>%
  unique() # so there were 7202 new species with actual impacts 

impacts_test_spp <- all_impacts %>% 
  filter(sciname %in% flipped_spp) %>%
  distinct(sciname, ingredient) %>%
  mutate(type = ifelse(ingredient %in% c("trimmings fish_fish meal, cut offs", "trimmings fish_fish oil, cut offs", "forage fish_fish meal", "forage fish_fish oil"), "marine", "terrestrial")) %>%
  distinct(sciname, type) %>%
  pull(sciname) %>%
  unique()

impacts_test_flipped_spp_impact <- all_impacts %>% 
  filter(sciname %in% impacts_test_spp) %>% 
  group_by(diet, sensitivity_scenario) %>%
  summarise(total_impacts = sum(impact_total, na.rm = TRUE)) %>%
  ungroup()


# fish flipped impacts
4195.089/605338.5
# plant flipped spp impacts
1729.887/650006.6
```


```{r}
# 1. Calculate scenario difference under original sensitivity
impact_diff_by_species <- all_impacts %>%
  filter(sensitivity_scenario == "original") %>%
  group_by(sciname, diet) %>%
  summarise(total_impact = sum(impact_total, na.rm = TRUE)) %>%
  pivot_wider(names_from = diet, values_from = total_impact, names_prefix = "diet_") %>%
  mutate(diff_original = abs(diet_fish - diet_plant)) %>%
  ungroup()

# 2. Compute variability across sensitivity scenarios per species & diet
sensitivity_variation <- all_impacts %>%
  group_by(sciname, diet, sensitivity_scenario) %>%
  summarise(total_impact = sum(impact_total, na.rm = TRUE)) %>%
  group_by(sciname, diet) %>%
  summarise(range_impact = max(total_impact) - min(total_impact),
            sd_impact = sd(total_impact, na.rm = TRUE)) %>%
  ungroup()

# 3. Merge difference and uncertainty
sensitivity_analysis <- impact_diff_by_species %>%
  inner_join(sensitivity_variation, by = "sciname") %>%
  mutate(signal_to_range = diff_original / range_impact,
         uncertainty_dominates = diff_original < range_impact)

# 4. Plot: signal vs uncertainty
ggplot(sensitivity_analysis, aes(x = diff_original, y = range_impact, color = diet)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Impact difference between feed types (original)",
       y = "Impact range due to sensitivity",
       title = "Feed Scenario Signal vs Sensitivity Uncertainty") +
  theme_minimal()

plot_df <- all_impacts %>%
  group_by(sensitivity_scenario, diet) %>%
  summarise(impact_total = sum(impact_total, na.rm = TRUE)) %>%
  ungroup()

# 5. Violin plot of impact by sensitivity for one feed type (e.g., plant)
ggplot(plot_df,
       aes(x = sensitivity_scenario, y = impact_total, fill = diet)) +
  geom_col(position = "dodge") + 
  #geom_violin(fill = "#86b8e0", alpha = 0.7) +
 # facet_grid(~diet, scales = "free_y") +
  theme_minimal() +
  labs(
       x = "Sensitivity Scenario",
       y = "Total Impact")



```

Lets map how each scenario agrees (or not) with one another on high vs low impact areas:
  	  - most impacted cells (>95 quantile?) in 1, 2, 3 scenarios (different colors); least impacted cells (<25th quantile) in 1, 2, 3 scenarios (different colors); this will tell us the places that for sure have high or low impacts 
  	  - Fig 1 from [Stock et al.](https://conbio.onlinelibrary.wiley.com/doi/full/10.1111/cobi.13141?saml_referrer)
      - do this across allocation approaches 
      
```{r}

# read in each scenario, filter for >95th quantile, set all cells to 1; filter for <25th quantile, set all cells to 1. Stack all and count what cells match up 

allocation_types <- c("economic", "mass", "ge")
fcr_types <- c("regular")
diet_types <- c("plant-dominant", "fish-dominant")



  for(diet_type in diet_types){
    for(fcr_type in fcr_types){
      
high_impact_rast <- moll_template
low_impact_rast <- moll_template

      # diet_type = "fish-dominant"
      # fcr_type = "efficient"

for(allocation_type in allocation_types){
      
      # allocation_type = "economic"
      
      imp_rast <- rast(glue(file.path(biodiv_dir, "output/impact_maps_across_taxon_ingredient/{diet_type}/{fcr_type}/{allocation_type}_mean_prop.tif")))
      
      quant_95 <- global(imp_rast, probs = 0.95, na.rm = TRUE)
      
      imp_rast_95 <- classify(imp_rast, cbind(quant_95$mean, Inf, 1), others = NA, right = FALSE)

      
      quant_25 <-  global(imp_rast, probs = 0.05, na.rm = TRUE)
      
      imp_rast_25 <- classify(imp_rast, cbind(0, quant_25$mean, 1), others = NA, right = TRUE)

      
      high_impact_rast <- c(high_impact_rast, imp_rast_95)
      
      low_impact_rast <- c(low_impact_rast, imp_rast_25)
      
    }
  


high_imp_sum <- sum(high_impact_rast, na.rm = TRUE)
low_imp_sum <- sum(low_impact_rast, na.rm = TRUE)

pal <- colorRampPalette(c("yellow","darkred"))


png(file = glue(here("analysis/uncertainty/{diet_type}_{fcr_type}/high_impact.png")), width = 1200, height = 700)
plot(high_imp_sum, col = pal(3), type = "classes") 
dev.off()

png(file = glue(here("analysis/uncertainty/{diet_type}_{fcr_type}/low_impact.png")), width = 1200, height = 700)
plot(low_imp_sum, col = pal(3), type = "classes") 
dev.off()

## look at scenario disagreement (sometimes scenarios has cells that are classified as both high impact and low impact)
all_imp_sum <- intersect(high_imp_sum, low_imp_sum)

r <- classify(all_imp_sum, cbind(FALSE, 0))

png(file = glue(here("analysis/uncertainty/{diet_type}_{fcr_type}/disagree.png")), width = 1200, height = 700)
plot(r)
dev.off()

    }
  }

```

```{r}
test <- 1 - rast(file.path(biodiv_dir, "output/impact_maps_across_taxon_ingredient/plant-dominant/regular/economic_mean_prop.tif"))
test[test == 0] <- NA
global(test, "mean", na.rm = TRUE) # 0.9999617
plot(raster(test))
s <- raster::select(raster(test))
plot(s)
where.max(test)


```

Lets see if the standard deviation generally scales with the mean; if it does, we should probably report coefficient of variation instead. 
 - these plots take a while to save - like 10 mins each.
```{r}
## read in mean rast for plant-dominant regular FCR

## read in sd rast for plant-dominant regular FCR

## calculate some sort of scaling plot

allocations <- c("ecomomic", "mass", "ge")

for(allocation_type in allocations){
  
# allocation_type = "economic"
  
# Read data
 filepath_mean <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/{allocation_type}_mean_prop.qs"))
 
  filepath_sd <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/{allocation_type}_sd_prop.qs"))

loop_df_mean <-  qs::qread(filepath_mean) %>%
  dplyr::select(-calc_type) %>%
  rename(mean_prop = value)

loop_df_sd <-  qs::qread(filepath_sd) %>%
    dplyr::select(-calc_type) %>%
  rename(sd_prop = value) 

loop_df_all <- loop_df_mean %>%
  left_join(loop_df_sd) %>%
  mutate(cv = sd_prop/mean_prop) %>%
  filter(!is.na(sd_prop))

loop_df_all$fcr_type = factor(loop_df_all$fcr_type, levels=c("regular", "efficient"))


p <- ggplot(data = loop_df_all, aes(x = mean_prop, y = sd_prop)) +
  geom_point() +
  geom_smooth(se = FALSE, formula = y~x) +
  geom_abline() +
  labs(caption = glue("{allocation_type}")) + 
  facet_grid(diet~fcr_type, scales = "free") 

ggsave(filename = glue(here("analysis/uncertainty/sd_explore/mean_sd_scaling_{allocation_type}.png")), plot = p)


}

```
