---
title: "Case study for publication"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Summary

This script has code which produces our example case study figure, Figure 4. We use the rarity weighted impact metric for this version: 

> Figure 4. Impact hotspot map of central Chile, under the plant-dominant feed scenario. (A) Cells with average impacts greater than or equal to the 95th quantile. (B) The raw material which contributes most to the impact in each hotspot cell. (C) The proportion of impacted species in each cell which is at-risk of extinction (i.e., the IUCN Red List category is worse than “Least Concern”). (D) Reference map to show location, using the Open Street Map (https://www.openstreetmap.org/).

For reference, I had to combine each of the panels in an image editor (I used Inkscape to combine), because I could not get it to look good using R. 

## Setup

```{r setup, include=FALSE}

## load libraries
library(tidyverse)
library(here)
library(data.table)
library(terra)
library(strex)
library(patchwork)
library(cowplot)
library(glue)
library(RColorBrewer)
library(ggpubr)
library(qs)
library(parallel)
library(doParallel)
library(sf)
library(rnaturalearth)
library(ggspatial)

# call necessary functions
source(here("src/directories.R"))
source(here("src/spatial.R"))

# establish filepaths
biodiv_dir <- file.path(rdsi_dir, "biodiversity_impacts")
impact_maps_dir <- file.path(biodiv_dir, "output")
this_dir <- here("analysis/")

# read in shapefile with country borders
countries_shp <- ne_countries(scale = 10, returnclass = "sf") |> 
  st_transform(crs(moll_template)) # read in countries shapefile

globe_bbox <- rbind(c(-180, -90), c(-180, 90), 
                      c(180, 90), c(180, -90), c(-180, -90)) 

globe_border <- st_polygon(list(globe_bbox)) %>%
  st_sfc(crs = 4326) %>%
  st_sf(data.frame(rgn = 'globe', geom = .)) %>%
  smoothr::densify(max_distance = 0.5) %>%
  st_transform(crs = crs(moll_template))


# establish color palettes
pal_mats <- c("Soybean" = "#328C41",
              "Trimmings fish" = "#8FBAED",
              "Sunflower" = "#4277A9",
              "Wheat" = "#FFE96D",
              "Rapeseed" = "#AC9900",
              "Pulses" = "#79E0DA",
              "Forage fish" = "#3DB1AB",
              "Maize" = "#0A6B66",
              "Other crops" = "#BC94C6")

quantile_pal <- c(
    "0-24" = "#D4F7F0", 
    "25-49" = "#538AC6", 
    "50-74" = "#1E438F", 
    "75-94" = "#F8D422", 
    "95-99" = "#E95D08", 
    "≥99" = "#781211"
)

iucn_pal <- c("NE" = "#808080",
              "DD" = "#808080",
              "LC" = "#46A440",
              "NT" = "#BDC72D",
              "VU" = "#F8D422",
              "EN" = "#E95D08",
              "CR" = "#781211")

```


Ok let's dive a bit deeper by looking into larger areas of impact, rather than just one cell.
What species are impacted in largest impact cells:
 - grab any cells that fall into the 95-99 or >= 99 quantile categories; cell list saved in 02_publication_maps.Rmd
 - I've done this separately for marine and terrestrial, and chunked into each scenario, just incase they get too large and unwieldy.
 - additionally, I've only filtered for these countries, to cut down on the size of the data: "NOR", "CAN", "ARG", "CHL", "AUS"
 
 - ONLY NEEDS TO BE RUN ONE TIME 
 
```{r}

large_impact_cells <- readRDS(here("analysis/int/large_impact_cells_rarity.rds")) 

## lets only do this for regular and economic scenarios
diet_types <- c("fish-dominant", "plant-dominant")
fcr_types <- c("regular")
allocation_types <- c("economic")


max_impact_df <- data.frame(species_full = NA, cell_id = NA, cropland_suitability = NA, spp_type = NA, filepath = NA,
                            species_id = NA, harvest = NA, hab_area = NA, impact_km2 = NA, source_file = NA,
                            raw_material = NA, diet = NA, allocation = NA, fcr = NA)

for(diet in diet_types){
  for(fcr in fcr_types){
    for(allocation in allocation_types){

# allocation = "economic"
# diet = "plant-dominant"
# fcr = "regular"

all_files <- list.files(file.path(biodiv_dir, "output/impact_maps_by_spp_ingredient_lists"), pattern = glue(".*_{diet}_{fcr}_.*_{allocation}"), full.names = TRUE)

to_remove <- grep("fish oil|fish meal", all_files)
all_files <- all_files[-to_remove]


all_files_list <- lapply(all_files, qread)

## now filter for specific cell_id and see what is impacted...

max_cell_impacts <- all_files_list %>% 
  map(~ .x %>%
        map(~ filter(.x, cell_id %in% c(large_impact_cells))))  # filter for any cells that are in the cells that have an average impact in the 95-99 or >=99th quantiles, in the countries listed: "NOR", "CAN", "ARG", "CHL", "AUS". The large impact cells file was created in the "02_publication_maps_rarity.Rmd script"


for (i in seq_along(max_cell_impacts)) {
  for (j in seq_along(max_cell_impacts[[i]])) {
    max_cell_impacts[[i]][[j]]$source_file <- all_files[i]
  }
}

max_cell_impacts_fin <- max_cell_impacts %>%
  bind_rows() %>%
  mutate(raw_material = str_before_first(str_after_nth(str_after_last(source_file, "/"), "_", 3), "_")) %>%
  mutate(diet = diet, 
         allocation = allocation, 
         fcr = fcr)

qsave(max_cell_impacts_fin, glue(here("analysis/int/large_cell_impacts_rarity_terrestrial_{diet}_{fcr}_{allocation}.qs"))) # save as qs file to save space


    }
  }
}



```

Need to do the same thing for marine impacts. I am chunking this just in case the data gets too large and unwieldy. 

```{r}
diet_types <- c("plant-dominant", "fish-dominant")
fcr_types <- c("regular")
allocation_types <- c("economic")


for(diet in diet_types){
  for(fcr in fcr_types){
    for(allocation in allocation_types){
      
      # diet = "fish-dominant"
      # fcr = "regular"
      # allocation = "economic"
      
max_impact_df <- data.frame(species_full = NA, cell_id = NA, cropland_suitability = NA, spp_type = NA, filepath = NA,
                            species_id = NA, harvest = NA, hab_area = NA, impact_km2 = NA, source_file = NA,
                            raw_material = NA, diet = NA, allocation = NA, fcr = NA)      
      
for(chunk_id in 1:13){
        
       # chunk_id = 4
        
        files <- list.files(file.path(biodiv_dir, "int/marine_aggregation/sub_chunk_dfs/"), pattern = glue(".*_{diet}_{fcr}_.*_{allocation}_chunk_{chunk_id}.qs"), full.names = TRUE)
        
        impacts_chunk <- lapply(files, qread) %>%
          bind_rows() %>%
          filter(!is.na(cell_id)) %>%
          filter(cell_id %in% large_impact_cells) %>%
  as.data.frame() %>% 
  mutate(diet = diet, 
         allocation = allocation, 
         fcr = fcr)

        qsave(impacts_chunk, glue(here("analysis/int/chunked/large_cell_impacts_rarity_marine_{diet}_{fcr}_{allocation}_chunk_{chunk_id}.qs")))


      }
                 
    }
  }
} 

## seems that the chunking may have been unnecessary

fd_impacts_list <- list.files(here("analysis/int/chunked/"), full.names = TRUE, pattern = "large_cell_impacts_rarity_marine_fish-dominant")

fd_impacts_df <- lapply(fd_impacts_list, qread) %>%
  bind_rows()

qsave(fd_impacts_df, here("analysis/int/large_cell_impacts_rarity_marine_fish-dominant_regular_economic.qs"))

pd_impacts_list <- list.files(here("analysis/int/chunked/"), full.names = TRUE, pattern = "large_cell_impacts_rarity_marine_plant-dominant")

pd_impacts_df <- lapply(pd_impacts_list, qread) %>%
  bind_rows()

qsave(pd_impacts_df, here("analysis/int/large_cell_impacts_rarity_marine_plant-dominant_regular_economic.qs")) 


```


Ok, let's pull the "hotspot" areas for some examples:

 - Chile (coastal)
 - Norway
 - Argentina
 - Canada
 - Australia


```{r}
max_impact_df_terrestrial_pd <- qread(here(this_dir, "int/large_cell_impacts_rarity_terrestrial_plant-dominant_regular_economic.qs"))

max_impact_df_marine_pd <- qread(here(this_dir, "int/large_cell_impacts_rarity_marine_plant-dominant_regular_economic.qs"))

land_eez_rgns <- read.csv(file.path(biodiv_dir, "spatial/rgns_xy_all.csv")) %>%
  dplyr::select(x, y, iso3c) %>% 
  left_join(moll_template_xy)


max_impact_iso3c_terr <- max_impact_df_terrestrial_pd %>%
  left_join(land_eez_rgns) %>%
  dplyr::select(species = species_full, cell_id, impact_km2, taxon = spp_type, raw_material, allocation, fcr, diet, x, y, iso3c) %>%
  mutate(species = tolower(species)) %>%
    mutate(raw_material = case_when(raw_material == "CropsNES" ~ "Cropsnes",
                                    TRUE ~ raw_material))

unique(max_impact_iso3c_terr$raw_material) # [1] "Cropsnes"  "Maize"     "Pulses"    "Rapeseed"  "Soybean"   "Sunflower" "Wheat"  

max_impact_iso3c_marine <- max_impact_df_marine_pd %>%
  left_join(land_eez_rgns) %>%
    dplyr::select(species, cell_id, impact_km2, taxon, raw_material = fish_type, allocation, fcr, diet, x, y, iso3c) %>%
  mutate(raw_material = case_when(raw_material == "forage fish" ~ "Forage fish",
                                  raw_material == "trimmings fish" ~ "Trimmings fish",
       TRUE ~ raw_material))

unique(max_impact_iso3c_marine$raw_material) # [1] "Forage fish"    "Trimmings fish"

unique(max_impact_iso3c_terr$iso3c) # [1] "CAN" "NOR" "AUS" "ARG" "CHL" ok cool 
unique(max_impact_iso3c_marine$iso3c) # [1] "NOR" "AUS" "CHL" "ARG" "CAN" 
q95 <- 5 * 10^-11 # taken from the publication mapping script

```

**Figure 4. Impact hotspot map of central Chile, under the plant-dominant feed scenario.**
 > (A) Cells with average impacts greater than or equal to the 95th quantile. (B) The raw material which contributes most to the impact in each hotspot cell. (C) The proportion of impacted species in each cell which is at-risk of extinction (i.e., the IUCN Red List category is worse than “Least Concern”). (D) Reference map to show location, using the Open Street Map (https://www.openstreetmap.org/). 

```{r}
all_habs <- readRDS(here("analysis/int/all_spp_hab_areas.rds")) %>%
  dplyr::select(-spp_type) %>%
  mutate(sciname = tolower(sciname))

iucn_cats <- read.csv(here("prep/03_prep_spp_habitats/int/spp_iucn_cats.csv"))

all_richness <- rast(list.files(here("data/spatial/output"), pattern = "richness", full.names = TRUE)) %>% 
    app(., "sum", na.rm = TRUE) %>%
  as.data.frame(., xy = TRUE) %>%
  left_join(moll_template_xy) %>%
  dplyr::select(cell_id, sum_spp = sum)

# Now Chile

max_impact <- max_impact_iso3c_terr %>%
  filter(iso3c == "CHL") %>% # lets check out Chile
  rbind(max_impact_iso3c_marine %>% filter(iso3c == "CHL")) %>%
  filter(y > -4250000) %>%
  filter(y < -3900000) %>%
  filter(x< -6000000)  # filter for a small part of the coast that has both marine and land large impacts. kind of near Valparaíso I think?

impact_rast <- max_impact %>%
  # filter(raw_material == "forage fish") %>%
  group_by(x, y) %>%
  summarise(impact_km2 = sum(impact_km2, na.rm = TRUE)) %>%
  ungroup() %>%
  rast(., crs = moll_template) %>%
  project(., "EPSG:4326")

plot(impact_rast) # ok this is looking good now...


summary <- max_impact %>% 
  group_by(x, y, cell_id, species, raw_material) %>%
  summarise(impact_km2 = sum(impact_km2)) %>%
  ungroup() %>%
  group_by(x, y, cell_id, species) %>%
  mutate(total_spp_impact = sum(impact_km2, na.rm = TRUE)) %>%
  mutate(prop_total_spp_impact = total_spp_impact/100) %>%
  ungroup() %>%
  left_join(all_habs, by = c("species" = "sciname")) %>%
  mutate(rarity_spp_impact = prop_total_spp_impact * (1/total_hab_area)) %>%
  mutate(prop_raw_material_impact = impact_km2/total_spp_impact) %>%
  group_by(x, y, cell_id) %>%
  mutate(avg_impact = mean(prop_total_spp_impact, na.rm = TRUE),
         avg_rarity_impact = mean(rarity_spp_impact, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(iucn_cats, by = c("species" = "sciname"))

raw_mat_summary <- summary %>% 
  group_by(x, y, cell_id) %>%
  filter(prop_raw_material_impact == max(prop_raw_material_impact)) %>%
  dplyr::select(x, y, cell_id, dom_raw_material = raw_material) %>%
  ungroup() %>%
  distinct(x, y, cell_id, dom_raw_material)


q99 <- 6 * 10^-10 

summary_fin <- max_impact %>%
  group_by(x, y, cell_id, species) %>%
  summarise(impact_km2 = sum(impact_km2, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(prop_impact = impact_km2/100) %>%
    left_join(all_habs, by = c("species" = "sciname")) %>%
  mutate(rarity_impact = prop_impact * (1/total_hab_area)) %>%
  group_by(x, y, cell_id) %>%
  summarise(prop_impact_mean = mean(prop_impact, na.rm = TRUE),
            rarity_impact_mean = mean(rarity_impact, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(raw_mat_summary) %>%
   mutate(imp_category = case_when(
   rarity_impact_mean < q99 ~ "95-99",
    rarity_impact_mean >= q99 ~ "≥99"
  )) %>%
  filter(rarity_impact_mean >= q95) # make sure we are above 95th quantile



summary_iucn <- summary %>% 
  filter(cell_id %in% unique(summary_fin$cell_id))
  
iucn <- summary %>% 
  filter(cell_id %in% unique(summary_fin$cell_id)) %>%
  filter(!(category %in% c("NE", "DD", "LC")))

perc_threatened_impacts <- round((sum(iucn$impact_km2)/sum(summary_iucn$impact_km2))*100,1) # 6.6% of impacts shown are on threatened species? 

limits <- summary_fin %>%
  dplyr::select(x, y, prop_impact_mean) %>%
  rast(., crs = moll_template)

## now lets try plotting this:
country_sf <- countries_shp %>%
                          filter(sov_a3 == "CHL") %>%
  #   st_transform(crs(impact_rast)) %>%
  # st_crop(., st_bbox(impact_rast))
   st_crop(., st_bbox(limits))

summary_fin <- summary_fin %>%
  mutate(dom_raw_material = ifelse(dom_raw_material == "Cropsnes", "Other crops", dom_raw_material))

summary_fin$dom_raw_material <- factor(summary_fin$dom_raw_material, levels = c("Rapeseed", "Forage fish", "Wheat", "Pulses", "Other crops"))



ggplot() +
  geom_tile(data = summary_fin, aes(x = x, y = y, fill = dom_raw_material)) + 
  #   geom_sf(data = summary_fin_sf, aes(fill = dom_raw_material)) + 
       geom_sf(data = country_sf, fill = NA, colour = "black", lwd = 0.3)  +
  theme_minimal() +
      theme(legend.position = "bottom", ## put legend on bottom
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        legend.key = element_rect(fill = NA),
        axis.text.x = element_text(angle = 60, hjust = 1),
        text=element_text(family="Arial"),
              #  legend.text = element_text(angle = 90, hjust = 1),  # Tilt legend labels
panel.border = element_rect(colour = "grey69", fill=NA, size=3, linetype = "longdash")) +
  scale_fill_manual(values = pal_mats, labels = c("Rapeseed", "Forage\nfish", "Wheat", "Pulses", "Other\ncrops")) + 
    labs(fill = "Highest impact raw material", 
         x = "", y = "",
                  tag = bquote(bold("B"))) +
    guides(fill = guide_legend(
    title.position = "top", 
    title.hjust = 0.5, 
    barheight = 1,
    label.position = "bottom",
    label.hjust = 0.5,
    label.vjust = 1,
    keywidth = unit(1.5, "cm")
  )) 


ggsave(here("analysis/MS_plots/fig4_raw_material_rarity.png"), dpi = 300, bg = "white",  height = 16, width = 16, units = "cm")


summary_fin$imp_category <- factor(summary_fin$imp_category, levels = c("0-24", "25-49", "50-74", "75-94", "95-99", "≥99"))

ggplot() +
  geom_tile(data = summary_fin, aes(x = x, y = y, fill = imp_category)) + 
       geom_sf(data = country_sf, fill = NA,  colour = "black", lwd = 0.3)  +
  theme_minimal() +
      theme(legend.position = "bottom", ## put legend on bottom
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        legend.key = element_rect(fill = NA),
        axis.text.x = element_text(angle = 60, hjust = 1),
        text=element_text(family="Arial"),
panel.border = element_rect(colour = "grey69", fill=NA, size=3, linetype = "longdash")) +
  scale_fill_manual(values = quantile_pal) + 
    labs(fill = "Proportion of habitat impacted \nQuantiles",
         x = "", y = "",
                  tag = bquote(bold("A"))) + 
      guides(fill = guide_legend(
    title.position = "top", 
    title.hjust = 0.5, 
    barheight = 1,
    label.position = "bottom",
    label.hjust = 0.5,
    label.vjust = 1,
    keywidth = unit(1.5, "cm")
  ))

ggsave(here("analysis/MS_plots/fig4_hotspot_quantile_rarity.png"), dpi = 300, bg = "white", height = 16, width = 16, units = "cm")



## it would be nice to see the proportion of species impacted in the worst category out of the total species in that cell: so number of impacted species in the worst category in the cell / total number of species assessed - so in this case we don't care about the category of the non-impacted species, right? Lets try that out..

iucn_plot_df <- summary_iucn %>%
  mutate(iucn_rank = case_when(category %in% c("NE") ~ 0,
                               category == "DD" ~ 1, 
                               category == "LC" ~ 2,
                               category == "NT" ~ 3,
                               category == "VU" ~ 4,
                               category == 'EN' ~ 5,
                               category == "CR" ~ 6)) %>%
  mutate(category_large = case_when(category %in% c("NE", "DD", "LC") ~ "Not threatened or data deficient",
                                    TRUE ~ "At-risk"),
         new_rank = case_when(category %in% c("NE", "DD", "LC") ~ 1,
                                    TRUE ~ 2)) %>%
  left_join(all_richness, by = "cell_id") %>%
  rename(sum_spp_assessed = sum_spp) %>%
  group_by(x, y, cell_id, category_large) %>%
  mutate(nspp_cats = n_distinct(species)) %>%
  ungroup() %>%
  group_by(x, y, cell_id) %>%
  mutate(total_spp_impacted = n_distinct(species),
         max_iucn_rank = max(new_rank, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(prop_spp_worst_iucn = nspp_cats/total_spp_impacted,
             prop_spp_worst_iucn_assessed = nspp_cats/sum_spp_assessed) %>%
  group_by(x, y, cell_id) %>%
  filter(new_rank == max_iucn_rank) %>%
  summarise(max_iucn_rank = max(new_rank, na.rm = TRUE), 
            prop_at_risk_impacted = mean(prop_spp_worst_iucn, na.rm = TRUE),
            prop_at_risk_assessed = mean(prop_spp_worst_iucn_assessed, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(max_iucn_cat = case_when(max_iucn_rank == 1 ~ "Not threatened or data deficient",
                                    TRUE ~ "At-risk")) %>%
  mutate(prop_at_risk_impacted = ifelse(max_iucn_rank == 1, 0, prop_at_risk_impacted))


## this shows the worst IUCN rank of the IMPACTED species, with the shading showing the proportion of species which are impacted in that worst category out of the total species impacted

## so for example, one of the cells is CR with a value of 0.14. This means that 14% of the species in the cell are impacted and CR, while the remainder are not impacted

ggplot() +
     #annotation_map_tile(type = "cartolight", zoom = 8) +
  geom_tile(data = iucn_plot_df, aes(x = x, y = y, fill = prop_at_risk_impacted)) + 
  geom_sf(data = country_sf, fill = NA,  colour = "black", lwd = 0.3)  +
  theme_minimal() +
      theme(legend.position = "bottom", ## put legend on bottom
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        legend.key = element_rect(fill = NA),
        axis.text.x = element_text(angle = 60, hjust = 1),
        text=element_text(family="Arial"),
panel.border = element_rect(colour = "grey69", fill=NA, size=3, linetype = "longdash")) +
     scale_fill_gradientn(
    colors = c("#46A440", "#F8D422", "#781211"),
    values = c(0, 0.04, 1),  # Adjusting the values to handle the 0 case
    na.value = "#46A440"  # In case of any NAs, but your data summary doesn't show NAs
  ) +
    labs(fill = "Proportion of impacted species at-risk", 
         x = "", y = "", 
         tag = bquote(bold("C"))) +
  guides(fill = guide_colorbar(
    title.position = "top", 
    title.hjust = 0.5,
    barwidth = unit(6.244166667, "cm"),
    barheight = 1
  ))

ggsave(here("analysis/MS_plots/fig4_iucn_new_rarity.png"), dpi = 300, bg = "white",  height = 16, width = 16, units = "cm")


## Now save reference plot: 

map_plot <- ggplot() +
      geom_sf(data = country_sf, fill = NA,  colour = "black", lwd = 0.3) +
   annotation_map_tile(type = "cartolight", zoom = 8) +
        geom_sf(data = country_sf, fill = NA,  colour = "black", lwd = 0.3) +
   theme_minimal() +
   theme(
      legend.position = "left",
      legend.title = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      legend.key = element_rect(fill = NA),
      axis.text.x = element_text(angle = 60, hjust = 1),
      text = element_text(family = "Arial"),
      panel.border = element_rect(colour = "grey69", fill = NA, size = 3, linetype = "longdash")
   ) +
   labs(
      x = "", y = "",
      tag = bquote(bold("D"))
   )      +
     theme(
      plot.margin = margin(t = 0, r = 0, b = 70, l = 0)  # Adjust margins to align with the map
   )


ggsave(here("analysis/MS_plots/fig4_base_map.png"), plot = map_plot, dpi = 300, bg = "white", height = 16, width = 16, units = "cm")

```


```{r}
#### Calculate some stats about this area

test <- iucn_plot_df %>% 
  filter(max_iucn_rank == 1)
218/373 # 0.5844504 of cells are not threatened
test <- iucn_plot_df %>% 
  filter(max_iucn_rank == 2, prop_at_risk_impacted == 1)
8/373

summary_distinct <- distinct(summary, cell_id, avg_impact)

sum(summary_distinct$avg_impact) # 1.722756 km2

## see how much of impacts this area takes up of chiles total imapcts 
 filepath <- glue(file.path(impact_maps_dir, "csvs/impact_maps_across_taxon_ingredient/economic_total_km2.qs"))
total_km2_df_chl <- qread(filepath) %>%
  left_join(land_eez_rgns, by = c("long" = "x", "lat" = "y")) %>%
  filter(iso3c == "CHL", diet == "plant-dominant")

sum(total_km2_df_chl$value) # 6694.324

test <- total_km2_df_chl %>%
  group_by(long, lat) %>%
  summarise(sum = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  rast(., type = "xyz")

# chl_marine <- max_impact_iso3c_marine %>%
#   filter(iso3c == "CHL", diet == "plant-dominant")
# sum(chl_marine$impact_km2) # 759.4346
# 
# chl_terr <- max_impact_iso3c_terr %>%
#   filter(iso3c == "CHL", diet == "plant-dominant")
# sum(chl_terr$impact_km2) # 5261.169

max_impact <- max_impact_iso3c_terr %>%
  filter(iso3c == "CHL") %>% # lets check out Chile
  rbind(max_impact_iso3c_marine %>% filter(iso3c == "CHL")) %>%
  filter(y > -4250000) %>%
  filter(y < -3900000) %>%
  filter(x< -6000000) 

sum(max_impact$impact_km2)/sum(total_km2_df_chl$value) # 0.3516732 - this hotspot represents 35% of all aquafeed impacts in Chile's land area and EEZ. 

373*100 # 37300 km2; mainland Chile land and eez area is 2743128 km2 (wikipedia)
# 37300/2743128 # 1.35
# OR our data says chile land and EEZ is 40027*100 = 4002700 km2
# 37300/4002700 # 0.00931871 ~ 1%
# chat gpt says 4,438,091 km2 for land and EEZ in Chile... however, this is because they also include Chilean territories. Ours also includes territories, so we'll include that (it just makes our point stronger)


raw_mat_totals <- raw_mat_summary %>%
  group_by(dom_raw_material) %>% 
  summarise(n_cells = n()) %>%
  ungroup()

145/373 # 0.3887399 of the cells are dominated by rapeseed impacts

test <- iucn_plot_df %>% 
  filter(max_iucn_rank == 2) %>%
  left_join(raw_mat_summary) %>%
  filter(dom_raw_material == "Forage fish")

mean(test$prop_at_risk_impacted, na.rm = TRUE) # 0.05550568 forage fish - of the forage fish cells, only ~5.5% of species in the cells are at-risk

test <- iucn_plot_df %>% 
  filter(max_iucn_rank == 2) 


test <- iucn_plot_df %>% 
  filter(max_iucn_rank == 2) %>%
  left_join(raw_mat_summary) %>%
  group_by(dom_raw_material) %>%
  summarise(n()) %>%
  mutate(prop = `n()`/153)

```

Do the same thing, but for Canada, and save it as a supplementary plot

```{r}
# filter for CAN

max_impact <- max_impact_iso3c_terr %>%
  filter(iso3c == "CAN") %>% # lets check out Canada cool only 100k cells
  rbind(max_impact_iso3c_marine %>% filter(iso3c == "CAN")) %>%
  filter(x < -7250000) %>%
  filter(x > -8800000) %>%
  filter(y <6400000) %>%
  filter(y >= 5775048)

rast <- max_impact %>%
  group_by(x, y) %>%
  summarise(impact_km2 = sum(impact_km2, na.rm = TRUE)) %>%
  ungroup() %>%
  rast(., crs = moll_template)

plot(rast) # ok this is looking good now...


summary <- max_impact %>% 
  group_by(x, y, cell_id, species, raw_material) %>%
  summarise(impact_km2 = sum(impact_km2)) %>%
  ungroup() %>%
  group_by(x, y, cell_id, species) %>%
  mutate(total_spp_impact = sum(impact_km2, na.rm = TRUE)) %>%
  mutate(prop_total_spp_impact = total_spp_impact/100) %>%
  ungroup() %>%
  left_join(all_habs, by = c("species" = "sciname")) %>%
  mutate(rarity_spp_impact = prop_total_spp_impact * (1/total_hab_area)) %>%
  mutate(prop_raw_material_impact = impact_km2/total_spp_impact) %>%
  group_by(x, y, cell_id) %>%
  mutate(avg_impact = mean(prop_total_spp_impact, na.rm = TRUE),
         avg_rarity_impact = mean(rarity_spp_impact, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(iucn_cats, by = c("species" = "sciname"))

raw_mat_summary <- summary %>% 
  group_by(x, y, cell_id) %>%
  filter(prop_raw_material_impact == max(prop_raw_material_impact)) %>%
  dplyr::select(x, y, cell_id, dom_raw_material = raw_material) %>%
  ungroup() %>%
  distinct(x, y, cell_id, dom_raw_material)


q99 <- 6 * 10^-10 

summary_fin <- max_impact %>%
  group_by(x, y, cell_id, species) %>%
  summarise(impact_km2 = sum(impact_km2, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(prop_impact = impact_km2/100) %>%
    left_join(all_habs, by = c("species" = "sciname")) %>%
  mutate(rarity_impact = prop_impact * (1/total_hab_area)) %>%
  group_by(x, y, cell_id) %>%
  summarise(prop_impact_mean = mean(prop_impact, na.rm = TRUE),
            rarity_impact_mean = mean(rarity_impact, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(raw_mat_summary) %>%
   mutate(imp_category = case_when(
   rarity_impact_mean < q99 ~ "95-99",
    rarity_impact_mean >= q99 ~ "≥99"
  )) %>%
  filter(rarity_impact_mean >= q95) # make sure we are above 95th quantile


summary_iucn <- summary %>% 
  filter(cell_id %in% unique(summary_fin$cell_id))
  
iucn <- summary %>% 
  filter(cell_id %in% unique(summary_fin$cell_id)) %>%
  filter(!(category %in% c("NE", "DD", "LC")))

perc_threatened_impacts <- round((sum(iucn$impact_km2)/sum(summary_iucn$impact_km2))*100,1) # 3.7% of impacts shown are on threatened species? 

limits <- summary_fin %>%
  dplyr::select(x, y, prop_impact_mean) %>%
  rast(., crs = moll_template)

## now lets try plotting this:
country_sf <- countries_shp %>%
                          filter(sov_a3 == "CAN") %>%
  #   st_transform(crs(impact_rast)) %>%
  # st_crop(., st_bbox(impact_rast))
   st_crop(., st_bbox(limits))

summary_fin <- summary_fin %>%
  mutate(dom_raw_material = ifelse(dom_raw_material == "Cropsnes", "Other crops", dom_raw_material))

summary_fin$dom_raw_material <- factor(summary_fin$dom_raw_material, levels = c("Other crops", "Pulses", "Rapeseed", "Soybean", "Wheat"))

ggplot() +
  geom_tile(data = summary_fin, aes(x = x, y = y, fill = dom_raw_material)) + 
       geom_sf(data = country_sf, fill = NA, colour = NA, lwd = 0.3)  +
  theme_minimal() +
      theme(legend.position = "bottom", ## put legend on bottom
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        legend.key = element_rect(fill = NA),
        axis.text.x = element_text(angle = 60, hjust = 1),
        text=element_text(family="Arial"),
panel.border = element_rect(colour = "grey69", fill=NA, size=3, linetype = "longdash")) +
  scale_fill_manual(values = pal_mats, labels = c("Other/ncrops", "Pulses", "Rapeseed", "Soybean", "Wheat")) + 
    labs(fill = "Highest impact raw material", 
         x = "", y = "",
                  tag = bquote(bold("B"))) +
    guides(fill = guide_legend(
    title.position = "top", 
    title.hjust = 0.5, 
    barheight = 1,
    label.position = "bottom",
    label.hjust = 0.5,
    label.vjust = 1,
    keywidth = unit(1.5, "cm")
  )) 


ggsave(here("analysis/SI_plots/case_studies/fig4_raw_material_rarity_CAN.png"), dpi = 300, bg = "white",  height = 16, width = 16, units = "cm")


summary_fin$imp_category <- factor(summary_fin$imp_category, levels = c("0-24", "25-49", "50-74", "75-94", "95-99", "≥99"))

ggplot() +
  geom_tile(data = summary_fin, aes(x = x, y = y, fill = imp_category)) + 
       geom_sf(data = country_sf, fill = NA,  colour = NA, lwd = 0.3)  +
  theme_minimal() +
      theme(legend.position = "bottom", ## put legend on bottom
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        legend.key = element_rect(fill = NA),
        axis.text.x = element_text(angle = 60, hjust = 1),
        text=element_text(family="Arial"),
panel.border = element_rect(colour = "grey69", fill=NA, size=3, linetype = "longdash")) +
  scale_fill_manual(values = quantile_pal) + 
    labs(fill = "Proportion of habitat impacted \nQuantiles",
         x = "", y = "",
                  tag = bquote(bold("A"))) + 
      guides(fill = guide_legend(
    title.position = "top", 
    title.hjust = 0.5, 
    barheight = 1,
    label.position = "bottom",
    label.hjust = 0.5,
    label.vjust = 1,
    keywidth = unit(1.5, "cm")
  ))

ggsave(here("analysis/SI_plots/case_studies/fig4_hotspot_quantile_rarity_CAN.png"), dpi = 300, bg = "white", height = 16, width = 16, units = "cm")



## it would be nice to see the proportion of species impacted in the worst category out of the total species in that cell: so number of impacted species in the worst category in the cell / total number of species assessed - so in this case we don't care about the category of the non-impacted species, right? Lets try that out..

iucn_plot_df <- summary_iucn %>%
  mutate(iucn_rank = case_when(category %in% c("NE") ~ 0,
                               category == "DD" ~ 1, 
                               category == "LC" ~ 2,
                               category == "NT" ~ 3,
                               category == "VU" ~ 4,
                               category == 'EN' ~ 5,
                               category == "CR" ~ 6)) %>%
  mutate(category_large = case_when(category %in% c("NE", "DD", "LC") ~ "Not threatened or data deficient",
                                    TRUE ~ "At-risk"),
         new_rank = case_when(category %in% c("NE", "DD", "LC") ~ 1,
                                    TRUE ~ 2)) %>%
  left_join(all_richness, by = "cell_id") %>%
  rename(sum_spp_assessed = sum_spp) %>%
  group_by(x, y, cell_id, category_large) %>%
  mutate(nspp_cats = n_distinct(species)) %>%
  ungroup() %>%
  group_by(x, y, cell_id) %>%
  mutate(total_spp_impacted = n_distinct(species),
         max_iucn_rank = max(new_rank, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(prop_spp_worst_iucn = nspp_cats/total_spp_impacted,
             prop_spp_worst_iucn_assessed = nspp_cats/sum_spp_assessed) %>%
  group_by(x, y, cell_id) %>%
  filter(new_rank == max_iucn_rank) %>%
  summarise(max_iucn_rank = max(new_rank, na.rm = TRUE), 
            prop_at_risk_impacted = mean(prop_spp_worst_iucn, na.rm = TRUE),
            prop_at_risk_assessed = mean(prop_spp_worst_iucn_assessed, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(max_iucn_cat = case_when(max_iucn_rank == 1 ~ "Not threatened or data deficient",
                                    TRUE ~ "At-risk")) %>%
  mutate(prop_at_risk_impacted = ifelse(max_iucn_rank == 1, 0, prop_at_risk_impacted))


## this shows the worst IUCN rank of the IMPACTED species, with the shading showing the proportion of species which are impacted in that worst category out of the total species impacted

## so for example, one of the cells is CR with a value of 0.14. This means that 14% of the species in the cell are impacted and CR, while the remainder are not impacted

ggplot() +
     #annotation_map_tile(type = "cartolight", zoom = 8) +
  geom_tile(data = iucn_plot_df, aes(x = x, y = y, fill = prop_at_risk_impacted)) + 
  geom_sf(data = country_sf, fill = NA,  colour = NA, lwd = 0.3)  +
  theme_minimal() +
      theme(legend.position = "bottom", ## put legend on bottom
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        legend.key = element_rect(fill = NA),
        axis.text.x = element_text(angle = 60, hjust = 1),
        text=element_text(family="Arial"),
panel.border = element_rect(colour = "grey69", fill=NA, size=3, linetype = "longdash")) +
     scale_fill_gradientn(
    colors = c("#46A440", "#F8D422", "#781211"),
    values = c(0, 0.04, 1),  # Adjusting the values to handle the 0 case
    na.value = "#46A440"  # In case of any NAs, but your data summary doesn't show NAs
  ) +
    labs(fill = "Proportion of impacted species at-risk", 
         x = "", y = "", 
         tag = bquote(bold("C"))) +
  guides(fill = guide_colorbar(
    title.position = "top", 
    title.hjust = 0.5,
    barwidth = unit(6.244166667, "cm"),
    barheight = 1
  ))

ggsave(here("analysis/SI_plots/case_studies/fig4_iucn_new_rarity_CAN.png"), dpi = 300, bg = "white",  height = 16, width = 16, units = "cm")


## Now save reference plot: 

map_plot <- ggplot() +
      geom_sf(data = country_sf, fill = NA,  colour = "black", lwd = 0.3) +
   annotation_map_tile(type = "cartolight", zoom = 8) +
        geom_sf(data = country_sf, fill = NA,  colour = "black", lwd = 0.3) +
   theme_minimal() +
   theme(
      legend.position = "left",
      legend.title = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      legend.key = element_rect(fill = NA),
      axis.text.x = element_text(angle = 60, hjust = 1),
      text = element_text(family = "Arial"),
      panel.border = element_rect(colour = "grey69", fill = NA, size = 3, linetype = "longdash")
   ) +
   labs(
      x = "", y = "",
      tag = bquote(bold("D"))
   )      +
     theme(
      plot.margin = margin(t = 0, r = 0, b = 70, l = 0)  # Adjust margins to align with the map
   )


ggsave(here("analysis/SI_plots/case_studies/fig4_base_map_CAN.png"), plot = map_plot, dpi = 300, bg = "white", height = 16, width = 16, units = "cm")

```

Do the same thing, but for Australia, and save it as a supplementary plot

```{r}
# filter for CAN

max_impact <- max_impact_iso3c_terr %>%
  filter(iso3c == "AUS") %>% # lets check out Australia 
  rbind(max_impact_iso3c_marine %>% filter(iso3c == "AUS"))  %>%
  filter(x < 10800000) %>%
  filter(y > -4800000) %>%
  filter(y < -3400000)
# %>%
#   filter(x < 13500000) %>%
#   filter(x > 12000000) %>%
#   filter(y < -4000000) %>%
#   filter(y > -4700000)

rast <- max_impact %>%
  group_by(x, y) %>%
  summarise(impact_km2 = sum(impact_km2, na.rm = TRUE)) %>%
  ungroup() %>%
  rast(., crs = moll_template)

plot(rast) # ok this is looking good now...


summary <- max_impact %>% 
  group_by(x, y, cell_id, species, raw_material) %>%
  summarise(impact_km2 = sum(impact_km2)) %>%
  ungroup() %>%
  group_by(x, y, cell_id, species) %>%
  mutate(total_spp_impact = sum(impact_km2, na.rm = TRUE)) %>%
  mutate(prop_total_spp_impact = total_spp_impact/100) %>%
  ungroup() %>%
  left_join(all_habs, by = c("species" = "sciname")) %>%
  mutate(rarity_spp_impact = prop_total_spp_impact * (1/total_hab_area)) %>%
  mutate(prop_raw_material_impact = impact_km2/total_spp_impact) %>%
  group_by(x, y, cell_id) %>%
  mutate(avg_impact = mean(prop_total_spp_impact, na.rm = TRUE),
         avg_rarity_impact = mean(rarity_spp_impact, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(iucn_cats, by = c("species" = "sciname"))

raw_mat_summary <- summary %>% 
  group_by(x, y, cell_id) %>%
  filter(prop_raw_material_impact == max(prop_raw_material_impact)) %>%
  dplyr::select(x, y, cell_id, dom_raw_material = raw_material) %>%
  ungroup() %>%
  distinct(x, y, cell_id, dom_raw_material)


q99 <- 6 * 10^-10 

summary_fin <- max_impact %>%
  group_by(x, y, cell_id, species) %>%
  summarise(impact_km2 = sum(impact_km2, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(prop_impact = impact_km2/100) %>%
    left_join(all_habs, by = c("species" = "sciname")) %>%
  mutate(rarity_impact = prop_impact * (1/total_hab_area)) %>%
  group_by(x, y, cell_id) %>%
  summarise(prop_impact_mean = mean(prop_impact, na.rm = TRUE),
            rarity_impact_mean = mean(rarity_impact, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(raw_mat_summary) %>%
   mutate(imp_category = case_when(
   rarity_impact_mean < q99 ~ "95-99",
    rarity_impact_mean >= q99 ~ "≥99"
  )) %>%
  filter(rarity_impact_mean >= q95) # make sure we are above 95th quantile


summary_iucn <- summary %>% 
  filter(cell_id %in% unique(summary_fin$cell_id))
  
iucn <- summary %>% 
  filter(cell_id %in% unique(summary_fin$cell_id)) %>%
  filter(!(category %in% c("NE", "DD", "LC")))

perc_threatened_impacts <- round((sum(iucn$impact_km2)/sum(summary_iucn$impact_km2))*100,1) # 3.7% of impacts shown are on threatened species? 

limits <- summary_fin %>%
  dplyr::select(x, y, prop_impact_mean) %>%
  rast(., crs = moll_template)

## now lets try plotting this:
country_sf <- countries_shp %>%
                          filter(sov_a3 == "CAN") %>%
  #   st_transform(crs(impact_rast)) %>%
  # st_crop(., st_bbox(impact_rast))
   st_crop(., st_bbox(limits))

summary_fin <- summary_fin %>%
  mutate(dom_raw_material = ifelse(dom_raw_material == "Cropsnes", "Other crops", dom_raw_material))

 summary_fin$dom_raw_material <- factor(summary_fin$dom_raw_material, levels = c("Forage fish", "Pulses", "Rapeseed", "Wheat"))

ggplot() +
  geom_tile(data = summary_fin, aes(x = x, y = y, fill = dom_raw_material)) + 
       geom_sf(data = country_sf, fill = NA, colour = NA, lwd = 0.3)  +
  theme_minimal() +
      theme(legend.position = "bottom", ## put legend on bottom
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        legend.key = element_rect(fill = NA),
        axis.text.x = element_text(angle = 60, hjust = 1),
        text=element_text(family="Arial"),
panel.border = element_rect(colour = "grey69", fill=NA, size=3, linetype = "longdash")) +
  scale_fill_manual(values = pal_mats, labels = c("Forage fish", "Pulses", "Rapeseed", "Wheat")) + 
    labs(fill = "Highest impact raw material", 
         x = "", y = "",
                  tag = bquote(bold("B"))) +
    guides(fill = guide_legend(
    title.position = "top", 
    title.hjust = 0.5, 
    barheight = 1,
    label.position = "bottom",
    label.hjust = 0.5,
    label.vjust = 1,
    keywidth = unit(1.5, "cm")
  )) 


ggsave(here("analysis/SI_plots/case_studies/fig4_raw_material_rarity_AUS.png"), dpi = 300, bg = "white",  height = 16, width = 16, units = "cm")


summary_fin$imp_category <- factor(summary_fin$imp_category, levels = c("0-24", "25-49", "50-74", "75-94", "95-99", "≥99"))

ggplot() +
  geom_tile(data = summary_fin, aes(x = x, y = y, fill = imp_category)) + 
       geom_sf(data = country_sf, fill = NA,  colour = NA, lwd = 0.3)  +
  theme_minimal() +
      theme(legend.position = "bottom", ## put legend on bottom
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        legend.key = element_rect(fill = NA),
        axis.text.x = element_text(angle = 60, hjust = 1),
        text=element_text(family="Arial"),
panel.border = element_rect(colour = "grey69", fill=NA, size=3, linetype = "longdash")) +
  scale_fill_manual(values = quantile_pal) + 
    labs(fill = "Rarity weighted proportion of habitat impacted \nQuantiles",
         x = "", y = "",
                  tag = bquote(bold("A"))) + 
      guides(fill = guide_legend(
    title.position = "top", 
    title.hjust = 0.5, 
    barheight = 1,
    label.position = "bottom",
    label.hjust = 0.5,
    label.vjust = 1,
    keywidth = unit(1.5, "cm")
  ))

ggsave(here("analysis/SI_plots/case_studies/fig4_hotspot_quantile_rarity_AUS.png"), dpi = 300, bg = "white", height = 16, width = 16, units = "cm")



## it would be nice to see the proportion of species impacted in the worst category out of the total species in that cell: so number of impacted species in the worst category in the cell / total number of species assessed - so in this case we don't care about the category of the non-impacted species, right? Lets try that out..

iucn_plot_df <- summary_iucn %>%
  mutate(iucn_rank = case_when(category %in% c("NE") ~ 0,
                               category == "DD" ~ 1, 
                               category == "LC" ~ 2,
                               category == "NT" ~ 3,
                               category == "VU" ~ 4,
                               category == 'EN' ~ 5,
                               category == "CR" ~ 6)) %>%
  mutate(category_large = case_when(category %in% c("NE", "DD", "LC") ~ "Not threatened or data deficient",
                                    TRUE ~ "At-risk"),
         new_rank = case_when(category %in% c("NE", "DD", "LC") ~ 1,
                                    TRUE ~ 2)) %>%
  left_join(all_richness, by = "cell_id") %>%
  rename(sum_spp_assessed = sum_spp) %>%
  group_by(x, y, cell_id, category_large) %>%
  mutate(nspp_cats = n_distinct(species)) %>%
  ungroup() %>%
  group_by(x, y, cell_id) %>%
  mutate(total_spp_impacted = n_distinct(species),
         max_iucn_rank = max(new_rank, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(prop_spp_worst_iucn = nspp_cats/total_spp_impacted,
             prop_spp_worst_iucn_assessed = nspp_cats/sum_spp_assessed) %>%
  group_by(x, y, cell_id) %>%
  filter(new_rank == max_iucn_rank) %>%
  summarise(max_iucn_rank = max(new_rank, na.rm = TRUE), 
            prop_at_risk_impacted = mean(prop_spp_worst_iucn, na.rm = TRUE),
            prop_at_risk_assessed = mean(prop_spp_worst_iucn_assessed, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(max_iucn_cat = case_when(max_iucn_rank == 1 ~ "Not threatened or data deficient",
                                    TRUE ~ "At-risk")) %>%
  mutate(prop_at_risk_impacted = ifelse(max_iucn_rank == 1, 0, prop_at_risk_impacted))


## this shows the worst IUCN rank of the IMPACTED species, with the shading showing the proportion of species which are impacted in that worst category out of the total species impacted

## so for example, one of the cells is CR with a value of 0.14. This means that 14% of the species in the cell are impacted and CR, while the remainder are not impacted

ggplot() +
  geom_tile(data = iucn_plot_df, aes(x = x, y = y, fill = prop_at_risk_impacted)) + 
  geom_sf(data = country_sf, fill = NA,  colour = NA, lwd = 0.3)  +
  theme_minimal() +
      theme(legend.position = "bottom", ## put legend on bottom
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        legend.key = element_rect(fill = NA),
        axis.text.x = element_text(angle = 60, hjust = 1),
        text=element_text(family="Arial"),
panel.border = element_rect(colour = "grey69", fill=NA, size=3, linetype = "longdash")) +
     scale_fill_gradientn(
    colors = c("#46A440", "#F8D422", "#781211"),
    values = c(0, 0.04, 1),  # Adjusting the values to handle the 0 case
    na.value = "#46A440"  # In case of any NAs, but your data summary doesn't show NAs
  ) +
    labs(fill = "Proportion of impacted species at-risk", 
         x = "", y = "", 
         tag = bquote(bold("C"))) +
  guides(fill = guide_colorbar(
    title.position = "top", 
    title.hjust = 0.5,
    barwidth = unit(6.244166667, "cm"),
    barheight = 1
  ))

ggsave(here("analysis/SI_plots/case_studies/fig4_iucn_new_rarity_AUS.png"), dpi = 300, bg = "white",  height = 16, width = 16, units = "cm")


## Now save reference plot: 

# map_plot <- ggplot() +
#       geom_sf(data = country_sf, fill = NA,  colour = "black", lwd = 0.3) +
#    annotation_map_tile(type = "cartolight", zoom = 8) +
#         geom_sf(data = country_sf, fill = NA,  colour = "black", lwd = 0.3) +
#    theme_minimal() +
#    theme(
#       legend.position = "left",
#       legend.title = element_blank(),
#       axis.title.x = element_blank(),
#       axis.title.y = element_blank(),
#       legend.key = element_rect(fill = NA),
#       axis.text.x = element_text(angle = 60, hjust = 1),
#       text = element_text(family = "Arial"),
#       panel.border = element_rect(colour = "grey69", fill = NA, size = 3, linetype = "longdash")
#    ) +
#    labs(
#       x = "", y = "",
#       tag = bquote(bold("D"))
#    )      +
#      theme(
#       plot.margin = margin(t = 0, r = 0, b = 70, l = 0)  # Adjust margins to align with the map
#    )
# 
# 
# ggsave(here("analysis/SI_plots/case_studies/fig4_base_map_AUS.png"), plot = map_plot, dpi = 300, bg = "white", height = 16, width = 16, units = "cm")

```



Do the same thing, but for Norway, and save it as a supplementary plot

```{r}
# filter for NOR

max_impact_nor <- max_impact_iso3c_terr %>%
  filter(iso3c == "NOR") %>% # lets check out norway.. cool only 100k cells
  rbind(max_impact_iso3c_marine %>% filter(iso3c == "NOR")) # wow ok the land impacts are way more concentrate in norway than ocean impacts

nor_rast <- max_impact_nor %>%
  group_by(x, y) %>%
  summarise(impact_km2 = sum(impact_km2, na.rm = TRUE)) %>%
  ungroup() %>%
  rast(., crs = moll_template)


plot(nor_rast) # ok this is looking good now...


nor_summary <- max_impact_nor %>% 
  group_by(x, y, cell_id, species, raw_material) %>%
  summarise(impact_km2 = sum(impact_km2)) %>%
  ungroup() %>%
  group_by(x, y,cell_id, species) %>%
  mutate(total_spp_impact = sum(impact_km2, na.rm = TRUE)) %>%
  mutate(prop_total_spp_impact = total_spp_impact/100) %>%
  ungroup() %>%
  mutate(prop_raw_material_impact = impact_km2/total_spp_impact) %>%
  group_by(x, y, cell_id) %>%
  mutate(avg_impact = mean(prop_total_spp_impact, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(iucn_cats, by = c("species" = "sciname")) %>%
  mutate(raw_material = ifelse(raw_material == "Cropsnes", "Other crops", raw_material))

nor_raw_mat_summary <- nor_summary %>% 
  group_by(x, y, cell_id) %>%
  filter(prop_raw_material_impact == max(prop_raw_material_impact)) %>%
  dplyr::select(x, y, cell_id, dom_raw_material = raw_material) %>%
  ungroup() %>%
  distinct(x, y, cell_id, dom_raw_material)

nor_summary_fin <- max_impact_nor %>%
  group_by(x, y, cell_id, species) %>%
  summarise(impact_km2 = sum(impact_km2, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(prop_impact = impact_km2/100) %>%
  group_by(x, y, cell_id) %>%
  summarise(prop_impact_mean = mean(prop_impact, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(nor_raw_mat_summary) %>%
  filter(prop_impact_mean >= q95) %>% 
  filter(y <= 7295048) %>%
  mutate(imp_category = case_when(
   prop_impact_mean < 0.0005 ~ "95-99",
    prop_impact_mean >= 0.0005 ~ "≥99"
  ))


nor_limits <- nor_summary_fin %>%
  dplyr::select(x, y, prop_impact_mean) %>%
  rast(., crs = moll_template)

nor_rast <- nor_rast %>%
  crop(., nor_limits)


## now lets try plotting this:
country_sf <- countries_shp %>%
                          filter(sov_a3 == "NOR") %>%
  st_crop(., st_bbox(nor_rast))

ggplot() +
   geom_tile(data = nor_summary_fin, aes(x = x, y = y, fill = dom_raw_material)) + 
      geom_sf(data = country_sf, fill = NA,  colour = "black", lwd = 0.3) +
  theme_minimal() +
      theme(legend.position = "bottom", ## put legend on bottom
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        legend.key = element_rect(fill = NA),
        axis.text.x = element_text(angle = 60, hjust = 1),
        text=element_text(family="Arial"),
              #  legend.text = element_text(angle = 90, hjust = 1),  # Tilt legend labels
panel.border = element_rect(colour = "grey69", fill=NA, size=3, linetype = "longdash")) +
  scale_fill_manual(values = pal_mats, labels = c("Forage\nfish", "Other\ncrops", "Pulses", "Wheat")) + 
    labs(fill = "Highest impact raw material", 
         x = "", y = "",
                  tag = bquote(bold("B"))) +
    guides(fill = guide_legend(
    title.position = "top", 
    title.hjust = 0.5, 
    barheight = 1,
    label.position = "bottom",
    label.hjust = 0.5,
    label.vjust = 1,
    keywidth = unit(1.5, "cm")
  ))

ggsave(here("analysis/SI_plots/extra_plots/CS_NOR_raw_material.png"), height = 8, width = 8, units = "in", bg = "white")

nor_summary_fin$imp_category <- factor(nor_summary_fin$imp_category, levels = c("0-24", "25-49", "50-74", "75-94", "95-99", "≥99"))

ggplot() +
  geom_tile(data = nor_summary_fin, aes(x = x, y = y, fill = imp_category)) + 
      geom_sf(data = country_sf, fill = NA,  colour = "black", lwd = 0.3) +
  theme_minimal() +
      theme(legend.position = "bottom", ## put legend on bottom
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        legend.key = element_rect(fill = NA),
        axis.text.x = element_text(angle = 60, hjust = 1),
        text=element_text(family="Arial"),
panel.border = element_rect(colour = "grey69", fill=NA, size=3, linetype = "longdash")) +
  scale_fill_manual(values = quantile_pal) + 
    labs(fill = "Proportion of habitat impacted \nQuantiles",
         x = "", y = "",
                  tag = bquote(bold("A"))) + 
      guides(fill = guide_legend(
    title.position = "top", 
    title.hjust = 0.5, 
    barheight = 1,
    label.position = "bottom",
    label.hjust = 0.5,
    label.vjust = 1,
    keywidth = unit(1.5, "cm")
  ))

ggsave(here("analysis/SI_plots/extra_plots/CS_NOR_hotspot_quantile.png"), height = 8, width = 8, units = "in", bg = "white")



nor_summary_iucn <- nor_summary %>% 
  filter(cell_id %in% unique(nor_summary_fin$cell_id))
  
nor_iucn <- nor_summary %>% 
  filter(cell_id %in% unique(nor_summary_fin$cell_id)) %>%
  filter(!(category %in% c("NE", "DD", "LC")))

nor_summary_iucn <- nor_summary %>% 
  filter(cell_id %in% unique(nor_summary_fin$cell_id))

iucn_plot_df <- nor_summary_iucn %>%
  mutate(iucn_rank = case_when(category %in% c("NE") ~ 0,
                               category == "DD" ~ 1, 
                               category == "LC" ~ 2,
                               category == "NT" ~ 3,
                               category == "VU" ~ 4,
                               category == 'EN' ~ 5,
                               category == "CR" ~ 6)) %>%
  mutate(category_large = case_when(category %in% c("NE", "DD", "LC") ~ "Not threatened or data deficient",
                                    TRUE ~ "At-risk"),
         new_rank = case_when(category %in% c("NE", "DD", "LC") ~ 1,
                                    TRUE ~ 2)) %>%
  left_join(all_richness, by = "cell_id") %>%
  rename(sum_spp_assessed = sum_spp) %>%
  group_by(x, y, cell_id, category_large) %>%
  mutate(nspp_cats = n_distinct(species)) %>%
  ungroup() %>%
  group_by(x, y, cell_id) %>%
  mutate(total_spp_impacted = n_distinct(species),
         max_iucn_rank = max(new_rank, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(prop_spp_worst_iucn = nspp_cats/total_spp_impacted,
             prop_spp_worst_iucn_assessed = nspp_cats/sum_spp_assessed) %>%
  group_by(x, y, cell_id) %>%
  filter(new_rank == max_iucn_rank) %>%
  summarise(max_iucn_rank = max(new_rank, na.rm = TRUE), 
            prop_at_risk_impacted = mean(prop_spp_worst_iucn, na.rm = TRUE),
            prop_at_risk_assessed = mean(prop_spp_worst_iucn_assessed, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(max_iucn_cat = case_when(max_iucn_rank == 1 ~ "Not threatened or data deficient",
                                    TRUE ~ "At-risk")) %>%
  mutate(prop_at_risk_impacted = ifelse(max_iucn_rank == 1, 0, prop_at_risk_impacted))


ggplot() +
  geom_tile(data = iucn_plot_df, aes(x = x, y = y, fill = prop_at_risk_impacted)) + 
      geom_sf(data = country_sf, fill = NA,  colour = "black", lwd = 0.3) +
  theme_minimal() +
      theme(legend.position = "bottom", ## put legend on bottom
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        legend.key = element_rect(fill = NA),
        axis.text.x = element_text(angle = 60, hjust = 1),
        text=element_text(family="Arial"),
panel.border = element_rect(colour = "grey69", fill=NA, size=3, linetype = "longdash")) +
     scale_fill_gradientn(
    colors = c("#46A440", "#F8D422", "#781211"),
    values = c(0, 0.04, 1),  # Adjusting the values to handle the 0 case
    na.value = "#46A440"  # In case of any NAs, but your data summary doesn't show NAs
  ) +
    labs(fill = "Proportion of impacted species at-risk", 
         x = "", y = "", 
         tag = bquote(bold("C"))) +
  guides(fill = guide_colorbar(
    title.position = "top", 
    title.hjust = 0.5,
    barwidth = unit(6.244166667, "cm"),
    barheight = 1
  ))

ggsave(here("analysis/SI_plots/extra_plots/CS_NOR_iucn.png"), height = 8, width = 8, units = "in", bg = "white")



map_plot <- ggplot() +
      geom_sf(data = country_sf, fill = NA,  colour = "black", lwd = 0.3) +
    geom_tile(data = iucn_plot_df, aes(x = x, y = y), fill = NA) + # make sure extent matches
   annotation_map_tile(type = "cartolight", zoom = 8) +
        geom_sf(data = country_sf, fill = NA,  colour = "black", lwd = 0.3) +
   theme_minimal() +
   theme(
      legend.position = "left",
      legend.title = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      legend.key = element_rect(fill = NA),
      axis.text.x = element_text(angle = 60, hjust = 1),
      text = element_text(family = "Arial"),
      panel.border = element_rect(colour = "grey69", fill = NA, size = 3, linetype = "longdash")
   ) +
   scale_fill_manual(values = pal_mats) +
   labs(
      x = "", y = "",
      tag = bquote(bold("D"))
   )      +
     theme(
      plot.margin = margin(t = 0, r = 0, b = 70, l = 0)  # Adjust margins to align with the map
   )

ggsave(here("analysis/SI_plots/extra_plots/NOR_basemap.png"), plot = map_plot, dpi = 300, bg = "white", height = 16, width = 16, units = "cm")


```

